### 2022-03-29 Writing/draft_introduction.tex
Started writing an introduction section.

\section{Introduction}
In the field of Computer Vision (CV), Convolutional Neural Networks (CNNs) have played an important role for a long time [BRON]. Recently, however, a new type of neural network architecture, called the Vision Transformer (ViT), has gained state of the art performance on common learning benchmarks, including the ImageNet dataset [BRON].

Exciting as this may be, a plain ViT is not likely to become useful in circumstances with limited training data and/or only consumer-grade hardware available for learning. This is due to intrinsic limitations within the ViT architecture, that will be elaborated on below [subsectie]. The current paper investigates whether utilizing Transfer Learning (TL) capabilities of ViTs can help overcome these limitations, and if they allow ViTs to become the preferred architecture in the aforementioned circumstances as well.

This introduction section will first give high level descriptions of CNNs and ViTs, and how they compare to one another. What follows, is an explanation of TL, and why it might help ViTs in circumstances with limited data/compute available. Finally, a research question is proposed.

\subsection{Convolutional Neural Networks}
TODO

\subsection{Vision Transformers}
TODO

\paragraph{A limitation of Vision Transformers}
TODO

\subsection{Transfer Learning}
TODO

### 2022-03-31 Writing/draft_introduction.tex
mainly added some comments as a guide to what has to be worked out next.
Added a nice picture and some modifications to the introduction.

@@
 \section{Introduction}
-In the field of Computer Vision (CV), Convolutional Neural Networks (CNNs) have played an important role for a long time [BRON].
-Recently, however, a new type of neural network architecture, called the Vision Transformer (ViT), has gained state of the art performance on common learning benchmarks, including the ImageNet dataset [BRON].
+In the field of Computer Vision (CV), Convolutional Neural Networks (CNNs) have played an important role since the introduction of AlexNet, roughly a decade ago [BRON].
+Recently, however, a new type of neural network architecture, called Vision Transformer (ViT), has gained state of the art performance on common learning benchmarks, including the famous ImageNet dataset [BRON/VOETNOOT].
 
@@
 This is due to intrinsic limitations within the ViT architecture, that will be elaborated on below [subsectie].
-The current paper investigates whether utilizing Transfer Learning (TL) capabilities of ViTs can help overcome these limitations, and if they allow ViTs to become the preferred architecture in the aforementioned circumstances as well.
+The current paper investigates whether utilizing Transfer Learning (TL) capabilities of ViTs can help overcome these limitations, and if this allows ViTs to become the preferred architecture in the aforementioned circumstances as well.
 
@@
 What follows, is an explanation of TL, and why it might help ViTs in circumstances with limited data/compute available.
-Finally, a research question is proposed.
+Finally, background information on the topic of TL using ViTs is given, and a research question is proposed.
 
@@
 TODO
+% First type of reduction in number of weights (compared to MLP) is that it only looks at surrounding pixels, instead of all
+% Second type of reduction is that it re-uses the same weights for all pixels
 
@@
 TODO
+% While a CNN is restricted to only incorporate information from its surrounding pixels in a subsequent layer, a ViT
 
@@
 
+% CNNs have image specific inductive bias: the architecture itself is tailored to 2D images, and as such has structural information about how images work built-in.
+This is not the case for ViTs, which lack an image specific inductive bias, which makes them very data hungry.
+
 \subsection{Transfer Learning}
 TODO
+
+\subsection{The current study}
+TODO

### 2022-04-01 writing/draft_introduction.tex
some name changes, tweaks, and added paragraph headers to already outline the further structure of some introduction subsections.

--- 2022-03-31 Writing/draft_introduction.tex
@@
 \section{Introduction}
-In the field of Computer Vision (CV), Convolutional Neural Networks (CNNs) have played an important role since the introduction of AlexNet, roughly a decade ago [BRON].
-Recently, however, a new type of neural network architecture, called Vision Transformer (ViT), has gained state of the art performance on common learning benchmarks, including the famous ImageNet dataset [BRON/VOETNOOT].
+In the field of \textit{Computer Vision} (CV), \textit{Convolutional Neural Networks} (CNNs) have played an important role since the introduction of AlexNet, roughly a decade ago [BRON].
+Recently, however, a new type of neural network architecture, called \textit{Vision Transformer} (ViT), has gained state of the art performance on common learning benchmarks, including the famous ImageNet dataset [BRON/VOETNOOT].
 
@@
 This is due to intrinsic limitations within the ViT architecture, that will be elaborated on below [subsectie].
-The current paper investigates whether utilizing Transfer Learning (TL) capabilities of ViTs can help overcome these limitations, and if this allows ViTs to become the preferred architecture in the aforementioned circumstances as well.
+The current paper investigates whether utilizing \textit{Transfer Learning} (TL) capabilities of ViTs can help overcome these limitations, and if this allows ViTs to become the preferred architecture in the aforementioned circumstances as well.
 
@@
 \subsection{Convolutional Neural Networks}
+\paragraph{Architecture}
 TODO
+
+\paragraph{Behaviour}
+TODO
+
 % First type of reduction in number of weights (compared to MLP) is that it only looks at surrounding pixels, instead of all
@@
 \subsection{Vision Transformers}
+Where a CNN reduces the number of trainable weights by taking an image's topology into account, a ViT does this by learning itself what information it should incorporate in successive representations of the input.
+As such, it is not restricted to a local receptive field, but can potentially include information from all over a the input image.
+
+ViTs were first introduced in [BRON], and built around the encoder stack of a \textit{Transformer Neural Network} [BRON].
+
+\paragraph{Self-attention}
 TODO
+
+\paragraph{From Transformer to ViT}
+
 % While a CNN is restricted to only incorporate information from its surrounding pixels in a subsequent layer, a ViT
@@
 \paragraph{A limitation of Vision Transformers}
+While ViTs [GREAT STUFF]; the fact that an image's topology is not incorporated in the architecture, makes that they lack a so called image specific \textit{Inductive Bias}, which CNNs do have.
+Intuitively this means a ViT first has to learn what an image is, before it can move on to the learning task at hand.
+The result is that they require substantially more training data (and time) than CNNs do.
+% Bronnen waaruit dit blijkt?
 TODO

### 2022-04-05 writing/draft_introduction.tex
stuff

@@
 In the field of \textit{Computer Vision} (CV), \textit{Convolutional Neural Networks} (CNNs) have played an important role since the introduction of AlexNet, roughly a decade ago [BRON].
-Recently, however, a new type of neural network architecture, called \textit{Vision Transformer} (ViT), has gained state of the art performance on common learning benchmarks, including the famous ImageNet dataset [BRON/VOETNOOT].
+Recently, however, a new type of \textit{Artificial Neural Network} (ANN), called \textit{Vision Transformer} (ViT), has gained state of the art performance on common learning benchmarks, including the famous ImageNet dataset [BRON/VOETNOOT].
 
@@
 
-This introduction section will first give high level descriptions of CNNs and ViTs, and how they compare to one another.
+This introduction section will first give high level descriptions of CNNs and ViTs, and illustrate how they compare to one another.
 What follows, is an explanation of TL, and why it might help ViTs in circumstances with limited data/compute available.
@@
 \subsection{Convolutional Neural Networks}
+%SECOND DRAFT OF CNN EXPLANATION:
 \paragraph{Architecture}
-TODO
+The modern CNN architecture is often attributed to [BRON].
+It has asymptotically less trainable parameters per layer than the classical \textit{Multilayer Perceptron} (MLP) [BRON], and achieves this reduction in ways that make sense for the topology of an image.
+Firstly, connections are restricted to a neurons so called \textit{local receptive field}.
+This is a small patch of neurons in the previous layer, with a similar spatial location.
+The top right of figure [FIGREF] shows an example.
+Here, one neuron and its receptive field are highlighted.
+Also note how the hypothetical MLP implementation in the top left has many more connections overall.
 
-\paragraph{Behaviour}
-TODO
+Secondly, instead of using a different set of trainable weights to connect a neuron to its receptive field, the same set is shared.
+
+%FIRST DRAFT OF CNN EXPLANATION:
+%\paragraph{Architecture}
+% Het hele "This makes MLPs unfit" klinkt niet als een logisch gevolg. Meer uitleg nodig of schrappen >:-(
+%The modern CNN architecture is often attributed to [BRON], and can best be understood through comparison with the classical \textit{Multilayer Perceptron} (MLP) [BRON].
+In this ANN architecture, any neuron in layer $x$ is connected to every neuron in subsequent layer $y$ (see top left of figure [FIGREF]).
+This makes MLPs unfit for CV applications, as an unreasonable amount of training data is required.
+Moreover, training data would need to account for the lack of built-in invariance to distortions/translations.
+
+%CNNs reduce the number of weights in ways that make sense for images.
+Firstly, connections to a neuron are limited to its so called \textit{local receptive field}, which is a small neighbourhood of neurons in the previous layer.
+Secondly, instead of using a different set of weights for every receptive field, the same set (called a \textit{convolution kernel}) is used for all.
+Figure [FIGREF] illustrates this in the top right.
+The 4 accentuated connections map the highlighted receptive field to a single neuron.
+Note that none of the other connections add to the number of trainable parameters, as the same weights are used to map any receptive field to its corresponding neuron.
+The 2D grid of neurons this results in, is called a \textit{feature map}.
+In reality, multiple kernels and feature maps exist per layer.
+
+%Besides convolutional layers, CNNs contain so called \textit{pooling layers}, which reduce the spatial resolution of a feature map by aggregating patches of neurons into one -- using $max(\cdot)$, or $mean(\cdot)$, for example.
+
+%\paragraph{Properties}
+%Weights sharing between receptive fields makes CNNs largely invariant to translations, as some pattern in the input would be processed by a convolution kernel the same, regardless of whether it occurs in the top right, or bottom left.
+In addition, pooling layers make the architecture invariant to slight distortions.
+
+%More important, is that CNNs can be seen as automatic feature extractors.
+It is well documented (see, for example, [REFS]) that the first layers of a CNN detect low level features, such as edges and corners.
+These are combined in subsequent layers to detect higher level features.
+The final layer then (hopefully) represents the input in an abstract way, such that, for example, an MLP layer can use it to perform the task at hand.
 
@@
 Where a CNN reduces the number of trainable weights by taking an image's topology into account, a ViT does this by learning itself what information it should incorporate in successive representations of the input.
-As such, it is not restricted to a local receptive field, but can potentially include information from all over a the input image.
+As such, it is not restricted to a local receptive field, but can potentially include information from all over the input image.
 
@@
 \paragraph{A limitation of Vision Transformers}
-While ViTs [GREAT STUFF]; the fact that an image's topology is not incorporated in the architecture, makes that they lack a so called image specific \textit{Inductive Bias}, which CNNs do have.
+While ViTs are promising; the fact that an image's topology is not incorporated in the architecture, makes that they lack a so called image specific \textit{Inductive Bias}, which CNNs do have.
 Intuitively this means a ViT first has to learn what an image is, before it can move on to the learning task at hand.

### 2022-04-06 writing/draft_introduction.tex
Reduced the size of the CNN explanation.

@@
 It has asymptotically less trainable parameters per layer than the classical \textit{Multilayer Perceptron} (MLP) [BRON], and achieves this reduction in ways that make sense for the topology of an image.
-Firstly, connections are restricted to a neurons so called \textit{local receptive field}.
-This is a small patch of neurons in the previous layer, with a similar spatial location.
+Firstly, connections are restricted to a neuron its so called \textit{local receptive field}.
+This is a small patch of neurons in the previous layer, with similar spatial location.
 The top right of figure [FIGREF] shows an example.
 Here, one neuron and its receptive field are highlighted.
-Also note how the hypothetical MLP implementation in the top left has many more connections overall.
+The top left shows a hypothetical MLP implementation.
+Note how it has many more connections overall.
 
-Secondly, instead of using a different set of trainable weights to connect a neuron to its receptive field, the same set is shared.
+Secondly, instead of using a different set of trainable weights to connect a neuron to its receptive field, the same set, called a \textit{convolution kernel}, is shared.
+In figure [FIGREF] the 4 accentuated connections correspond to one such kernel.
+The other shown connections don't add to the number of trainable weights, as they use this kernel as well.
+Essentially, a one-layered mini-MLP is slided over the input grid to produce the output grid -- which is called a \textit{feature map}.
+In reality, multiple kernels and feature maps exist per layer.
+
+Finally, to reduce the size even further, so called \textit{pooling layers} are added.
+These reduce the spatial resolution of a feature map by aggregating patches of neurons into one, using $max(\cdot)$, or $mean(\cdot)$, for example.
+
+\paragraph{Properties}
+The architecture as described, is largely invariant to shifts and small distortions of the input image.
+More important, is that CNNs can be seen as automatic feature extractors.
+It is well documented (see, for example, [REFS]) that the first layers of a CNN detect low level features, such as edges and corners.
+These are combined in subsequent layers to detect higher level features.
+The final layer then (hopefully) represents the input in an abstract manner, such that, for example, an MLP-type layer can extract all the information it needs from it to perform the task at hand.
 
@@
 \paragraph{From Transformer to ViT}
+TODO
 

### 2022-04-09 writing/draft_introduction.tex
Some stuff. Don't even remember what.

@@
 \subsection{Convolutional Neural Networks}
-%SECOND DRAFT OF CNN EXPLANATION:
 \paragraph{Architecture}
@@
 
-%FIRST DRAFT OF CNN EXPLANATION:
-%\paragraph{Architecture}
-% Het hele "This makes MLPs unfit" klinkt niet als een logisch gevolg. Meer uitleg nodig of schrappen >:-(
-%The modern CNN architecture is often attributed to [BRON], and can best be understood through comparison with the classical \textit{Multilayer Perceptron} (MLP) [BRON].
-In this ANN architecture, any neuron in layer $x$ is connected to every neuron in subsequent layer $y$ (see top left of figure [FIGREF]).
-This makes MLPs unfit for CV applications, as an unreasonable amount of training data is required.
-Moreover, training data would need to account for the lack of built-in invariance to distortions/translations.
+\subsection{Vision Transformers}
+\paragraph{Architecture}
+Where a CNN reduces the number of trainable weights by taking an image's topology into account, a ViT does this by learning itself what information it should incorporate in successive representations of the input.
+As such, it is not restricted to a local receptive field, but can potentially include information from all over the previous layer.
 
-%CNNs reduce the number of weights in ways that make sense for images.
-Firstly, connections to a neuron are limited to its so called \textit{local receptive field}, which is a small neighbourhood of neurons in the previous layer.
-Secondly, instead of using a different set of weights for every receptive field, the same set (called a \textit{convolution kernel}) is used for all.
-Figure [FIGREF] illustrates this in the top right.
-The 4 accentuated connections map the highlighted receptive field to a single neuron.
-Note that none of the other connections add to the number of trainable parameters, as the same weights are used to map any receptive field to its corresponding neuron.
-The 2D grid of neurons this results in, is called a \textit{feature map}.
-In reality, multiple kernels and feature maps exist per layer.
+ViTs were first introduced in [BRON], and are based on the \textit{Transformer} architecture [BRON].
+This type of ANN was originally designed to solve problems in \textit{natural language processing} (a field it now dominates), and takes a variable length sequence of \textit{token embeddings}\footnote{The model has some vocabulary of tokens it knows, for example all words in the English language.
+For each it learns a vector representation, called a token embedding.} as input.
+While Transformers are made up of an \textit{encoder} and \textit{decoder}, the latter won't be discussed here, as it is discarded in the ViT architecture.
 
-%Besides convolutional layers, CNNs contain so called \textit{pooling layers}, which reduce the spatial resolution of a feature map by aggregating patches of neurons into one -- using $max(\cdot)$, or $mean(\cdot)$, for example.
+Central to the encoder is the \textit{self-attention mechanism}. The \textit{attention heads} that implement it, start with 3 learned projection matrices.
 
-%\paragraph{Properties}
-%Weights sharing between receptive fields makes CNNs largely invariant to translations, as some pattern in the input would be processed by a convolution kernel the same, regardless of whether it occurs in the top right, or bottom left.
-In addition, pooling layers make the architecture invariant to slight distortions.
+%The architecture consists of an \textit{encoder}, and \textit{decoder}, but since the latter is discarded in ViTs, it won't be discussed here.
 
-%More important, is that CNNs can be seen as automatic feature extractors.
-It is well documented (see, for example, [REFS]) that the first layers of a CNN detect low level features, such as edges and corners.
-These are combined in subsequent layers to detect higher level features.
-The final layer then (hopefully) represents the input in an abstract way, such that, for example, an MLP layer can use it to perform the task at hand.
+%Before entering the encoder, each token is converted into a corresponding vector representation, called a \textit{token embedding}.
+This is done using a learned mapping from tokens to vectors.
+To make the network aware of a token's location within the sequence, positional encodings are added to the embeddings as well\footnote{Trigonometric functions are used for this, instead of learned encodings.}.
 
-% First type of reduction in number of weights (compared to MLP) is that it only looks at surrounding pixels, instead of all
-% Second type of reduction is that it re-uses the same weights for all pixels
+%The encoder itself is composed of a stack of \textit{encoder blocks}.
+The first block gets the embeddings as input; the rest the output of its predecessor.
+Each block contains two components, namely a \textit{multiheaded self-attention} layer, which will be discussed below, and a simple MLP that processes all positions (i.e.
+vectors) of the input sequence individually.
+Residual connections add a component's output to its input, and this sum is then normalized; i.e.
+$Normalize(x + Component(x))$\footnote{Note that this means the dimensions of a component's output are identical to that of its input.}.
 
-\subsection{Vision Transformers}
-Where a CNN reduces the number of trainable weights by taking an image's topology into account, a ViT does this by learning itself what information it should incorporate in successive representations of the input.
-As such, it is not restricted to a local receptive field, but can potentially include information from all over the input image.
-
-ViTs were first introduced in [BRON], and built around the encoder stack of a \textit{Transformer Neural Network} [BRON].
+%Checklist:
+% - constant number of operations to relate tokens, so regardless of distance
+% - weight sharing because same learned operations are applied to each (each token goes through the same pipeline)
 
@@
 
-\paragraph{A limitation of Vision Transformers}
+\subsection{A limitation of Vision Transformers}
 While ViTs are promising; the fact that an image's topology is not incorporated in the architecture, makes that they lack a so called image specific \textit{Inductive Bias}, which CNNs do have.

### 2022-04-17 writing/report.tex
adding all components needed to compile the whole text

\documentclass[10pt,a4paper,twocolumn]{article}
\usepackage{bachelorproject}

\title{\vspace{-0.5in}{\bfseries\scshape 
    Transfer Learning Using Vision Transformers Project (No Title Yet)
    } \\ 
    \vspace{1ex} 
    \normalsize{Bachelor's Project Thesis}\vspace{-2ex}
}

\author{
    \normalsize{Vincent Tonkes, s3617157, v.tonkes@student.rug.nl,} \\ 
    \normalsize{Supervisor: Dr M. Sabatelli}
}

\date{\vspace{-5ex}}

\begin{document}

\twocolumn[
  \maketitle
  \begin{@twocolumnfalse}
%   \input{./tex/abstract.tex}
  \end{@twocolumnfalse}
]

\thispagestyle{firststyle}

\input{tex/draft_introduction.tex}

\bibliographystyle{plainnat}
\bibliography{literature}

\clearpage

%\begin{appendices}
%    \input{tex/appendixA.tex}
%    \newpage
%    \input{tex/appendixB.tex}
%    \newpage
%\end{appendices}

\end{document}

### 2022-04-19 writing/tex/draft_introduction.tex
Just went further with writing the introduction. Randomly pushing it now because haven't done that for some time..

\section{Introduction}
Since the introduction of AlexNet, roughly a decade ago, \textit{Convolutional Neural Networks} (CNNs) have played an important role in \textit{Computer Vision} (CV) \citep{krizhevsky2012imagened}. Recently, however, a new type of \textit{Artificial Neural Network} (ANN), called \textit{Vision Transformer} (ViT), gained state of the art performance on common learning benchmarks, including the ImageNet dataset\footnote{At the time of writing TODO is the best performing model, and the top TODO are all Transformer-based: (website)}.

Exciting as this may be, a plain ViT is not likely to become useful in circumstances with limited training data and/or only consumer-grade hardware available for learning. This is due to a limitation that is inherent to the ViT architecture, which will be elaborated on in section \ref{vits}. Consequently, the current paper investigates whether utilizing \textit{Transfer Learning} (TL) capabilities of ViTs can help overcome these limitations, and if this allows ViTs to become the preferred architecture in the aforementioned circumstances as well.

Before going into more detail, however, some background information is needed. The remainder of this section will first briefly describe CNNs and ViTs, and compare them to one another. What follows, is an explanation of TL, and why it might help ViTs in circumstances with limited data/compute available. Finally, related work is discussed, and a research question is proposed.

%\begin{figure}[!tbp]
%    \centering
%    \includegraphics[width=7.7cm]{img/vit_vs_cnn.png}
%    \caption{Intuition behind CNNs and ViTs. Note that many details are left out for the purpose of illustration. The top half compares MLP layers (left) to those of CNNs (right). A single neuron and its receptive field are highlighted. Connections are only allowed such as the accentuated 4, which reduces the total number of connections greatly. Moreover, the non-accentuated connections don't add to the number of trainable weights, as the same 4 are shared between all neuron-receptive-field-pairs.\\The bottom half draws a similar comparison between MLPs (left) and ViTs (right). A single query vector is combined with all key vectors to get a scalar value for each position (blue boxes). These are multiplied with corresponding value vectors (in red), and then added to get a weighted sum (green). This sum is fed through one or more MLP-type layers, to produce a single output position (opaque). The same steps are taken for all other positions (transparent) as well, using their corresponding query vectors.}
%    \label{vitVsCnn}
%\end{figure}

\subsection{Convolutional Neural Networks}
The modern CNN architecture is often attributed to \citeauthor{lecun1998gradient} (\citeyear{lecun1998gradient}), which describes how the number of trainable weights per layer can be reduced in manners that are sensible for an image's topology (i.e. by restricting connections to a neuron's local receptive field, and by sharing weights).

%which describes how the topology of an image can be taken into account to sensibly reduce the number of weights per layer (i.e. by restricting connections to a neuron's so called local receptive field, and by sharing weights between ).
%The top half of figure \ref{vitVsCnn} attempts to give an intuitive understanding of this, by drawing a comparison to the classical \textit{Multilayer Perceptron} (MLP).

CNNs are largely invariant to shifts and small distortions of the input image. More important, is that they can be seen as automatic feature extractors. It is well established (e.g. \citeauthor{lecun1998gradient}, \citeyear{lecun1998gradient}) that the first layers of a CNN detect low level features, such as edges and corners. These are combined in subsequent layers to detect higher order features. The final layer should then give a representative summary of the input, such that, for example, an MLP-type layer can extract all the information it needs from it to perform the task at hand.

%It has asymptotically less trainable parameters per layer than the classical \textit{Multilayer Perceptron} (MLP) [BRON], and achieves this reduction in ways that make sense for the topology of an image. Firstly, connections are restricted to a neuron its so called \textit{local receptive field}. This is a small patch of neurons in the previous layer, with similar spatial location. The top right of figure [FIGREF] shows an example. Here, one neuron and its receptive field are highlighted. The top left shows a hypothetical MLP implementation. Note how it has many more connections overall.

%Secondly, instead of using a different set of trainable weights to connect a neuron to its receptive field, the same set, called a \textit{convolution kernel}, is shared. In figure [FIGREF] the 4 accentuated connections correspond to one such kernel. The other shown connections don't add to the number of trainable weights, as they use this kernel as well. Essentially, a one-layered mini-MLP is slided over the input grid to produce the output grid -- which is called a \textit{feature map}. In reality, multiple kernels and feature maps exist per layer.

%Finally, to reduce the size even further, so called \textit{pooling layers} are added. These reduce the spatial resolution of a feature map by aggregating patches of neurons into one, using $max(\cdot)$, or $mean(\cdot)$, for example.

%\paragraph{Properties}
%The architecture as described, is largely invariant to shifts and small distortions of the input image. More important, is that CNNs can be seen as automatic feature extractors. It is well documented (see, for example, [REFS]) that the first layers of a CNN detect low level features, such as edges and corners. These are combined in subsequent layers to detect higher level features. The final layer then (hopefully) represents the input in an abstract manner, such that, for example, an MLP-type layer can extract all the information it needs from it to perform the task at hand.

\subsection{Vision Transformers} \label{vits}
\paragraph{Architecture}
Where CNNs reduce the number of trainable weights by taking an image's topology into account, ViTs do this by learning themselves what information they should incorporate in successive representations of the input. As such, they are not restricted to a local receptive field, but can potentially include information from all over the previous layer.

ViTs were first introduced in \citeauthor{dosovitskiy2020image} (\citeyear{dosovitskiy2020image}), and are based on the \textit{encoder stack} of the \textit{Transformer} architecture \citep{vaswani2017attention}. This stack takes a sequence of vectors as input, and is composed of alternating \textit{multi-headed self-attention} layers and position-wise MLP-type layers\footnote{position-wise meaning the MLP can take only a single position (constituent vector) as input at a time, and is shared between all positions.}.

The self-attention mechanism plays a central role in ViTs, and is worth elaborating on. Each head of a self-attention layer contains 3 learned projection matrices, which are used to produce $key$, $value$, and $query$ vectors for every position in the input sequence. The output for a single position is a weighted sum of all $value$ vectors. Weights are obtained by combining this position's $query$ with all $key$ vectors\footnote{In principle by taking softmax over all query-key dot products.}. As such, an attention head decides itself how much of each position should be incorporated in subsequent representations.

The original Transformer architecture was designed to solve problems in \textit{natural language processing} (a field it now dominates). To make it work for CV applications, ViTs divide an image into square patches, which get treated as tokens (words). Learned embeddings convert these into a sequence of vectors, to which 1D positional encodings are added\footnote{Do mind that ViTs have no built-in knowledge of an image's 2D topology.}. A learned $[class]$ vector gets prepended to the sequence. Its representation at the end of the encoder stack is the final output, which can then be used for classification, segmentation, etc.

%These are converted to learned embeddings, to which 1D position encodings are added. A learned $[class]$ embedding gets prepended to the sequence.

% TODO Using these it produces weighted sums of value vectors per position. The weights for a position are obtained by combining its $query$ vector with all $key$ vectors\footnote{sofmax TODO}.

%Combining a single $query$ with all keys\footnote{softmax over all dot products. Additionally, all dot products are first divided by the square root of the size of a $key$ vector, as to prevent very small gradients from occurring.}, gives an array of scalars that tells for each $value$ vector how much of it 
% TODO Weight sharing is used in the form of all positions going through the same pipeline

%ViTs were first introduced in \citeauthor{dosovitskiy2020image} (\citeyear{dosovitskiy2020image}), and are based on the \textit{Transformer} architecture \citep{vaswani2017attention}. This type of ANN was originally designed to solve problems in \textit{natural language processing} (a field it now dominates). It takes a series of tokens as input, and uses a learned mapping to convert them into a series of vectors, called \textit{token embeddings}. These embeddings are then fed through an encoder and decoder, but since the latter is discarded for ViTs, it won't be discussed here.

%, and takes a variable length sequence of \textit{token embeddings}\footnote{The model has some vocabulary of tokens it knows, for example all words in the English language. For each it learns a vector representation, called a token embedding.} as input. While Transformers are made up of an \textit{encoder} and \textit{decoder}, the latter won't be discussed here, as it is discarded in the ViT architecture.

%Central to the encoder is the \textit{self-attention mechanism}. The \textit{attention heads} that implement it, start with 3 learned projection matrices.

%The architecture consists of an \textit{encoder}, and \textit{decoder}, but since the latter is discarded in ViTs, it won't be discussed here.

%Before entering the encoder, each token is converted into a corresponding vector representation, called a \textit{token embedding}. This is done using a learned mapping from tokens to vectors. To make the network aware of a token's location within the sequence, positional encodings are added to the embeddings as well\footnote{Trigonometric functions are used for this, instead of learned encodings.}.

%The encoder itself is composed of a stack of \textit{encoder blocks}. The first block gets the embeddings as input; the rest the output of its predecessor. Each block contains two components, namely a \textit{multiheaded self-attention} layer, which will be discussed below, and a simple MLP that processes all positions (i.e. vectors) of the input sequence individually. Residual connections add a component's output to its input, and this sum is then normalized; i.e. $Normalize(x + Component(x))$\footnote{Note that this means the dimensions of a component's output are identical to that of its input.}.

%Checklist:
% - constant number of operations to relate tokens, so regardless of distance
% - weight sharing because same learned operations are applied to each (each token goes through the same pipeline)

% While a CNN is restricted to only incorporate information from its surrounding pixels in a subsequent layer, a ViT

\paragraph{A limitation of Vision Transformers}
While ViTs are promising; the fact that they are not tailored to CV tasks, makes that they lack a so called image specific \textit{Inductive Bias}, which CNNs do have. Intuitively this means a ViT first has to learn what an image is, before it can move on to the learning task at hand. The result is that they require substantially more training data (and time) than CNNs do.

It is for this reason the paper that first introduced ViTs \citep{dosovitskiy2020image} also mentions using pre-trained models rather than randomly initialised ones, as it allows general knowledge about images to be carried over to the new task.

% CNNs have image specific inductive bias: the architecture itself is tailored to 2D images, and as such has structural information about how images work built-in. This is not the case for ViTs, which lack an image specific inductive bias, which makes them very data hungry.

\subsection{Transfer Learning}
TODO

\subsection{The current study and related work}
TODO

### 2022-04-21 writing/tex/draft_introduction.tex
At the point where all preliminaries are explained (CNNs, ViTs, TL). Up next: related work and description of current research

@@
 It is well established (e.g.
-\citeauthor{lecun1998gradient}, \citeyear{lecun1998gradient}) that the first layers of a CNN detect low level features, such as edges and corners.
+\citeauthor{lecun1998gradient}, \citeyear{yosinski2014transferable}; \citeauthor{yosinski2014transferable}, \citeyear{lecun1998gradient}) that the first layers of a CNN detect low level features, such as edges and corners.
 These are combined in subsequent layers to detect higher order features.
@@
 
-\paragraph{A limitation of Vision Transformers}
+\paragraph{A limitation of ViTs}
 While ViTs are promising; the fact that they are not tailored to CV tasks, makes that they lack a so called image specific \textit{Inductive Bias}, which CNNs do have.
@@
 
-It is for this reason the paper that first introduced ViTs \citep{dosovitskiy2020image} also mentions using pre-trained models rather than randomly initialised ones, as it allows general knowledge about images to be carried over to the new task.
+It is for this reason the paper that first introduced ViTs \citep{dosovitskiy2020image} mentions using pre-trained models rather than randomly initialised ones, as it allows general knowledge about images to be carried over to the new task.
+This technique falls within the category of TL, which will be cover below.
 
@@
 \subsection{Transfer Learning}
-TODO
+Given two \textit{Machine Learning} (ML) problems, of which one is called the source, and the other the target task, Transfer Learning can be defined as \textit{exploiting knowledge implied in the source task to improve performance of an ML model on the target task}.
+A formal definition is found on p.
+56 of \citeauthor{sabatelli2022contributions} (\citeyear{sabatelli2022contributions}), but for the purposes of this paper, the one above suffices.
+
+In practice, many TL techniques exist.
+The example given in section \ref{vits} can be described as \textit{fine-tuning} TL.
+Here, a pre-trained model is used as a starting point, and trained fully on the target task -- possibly with different hyper-parameters, and often with a new final layer that fits the task.
+\citeauthor{yosinski2014transferable} (\citeyear{yosinski2014transferable}) studied transferability properties of CNNs pre-trained on ImageNet, and suggested that \textit{fine-tuning} results in improved performance and generalizability compared to training from scratch.
+Moreover, while transferability decreased as source and target task became more dissimilar, utilizing TL still seemed to be preferred.
+
+% FINE-TUNING: "yosinski2014transferable" CNNs
+% 1) Lower layers contain general information that can easily be transferred (deze doe ik naar volgende stukje misschien)
+% 2) Distance between source and target task increases -> transferability decreases
+%       - BUT: still better than randomly initialized
+% 3) Increases generalizability (This one is important for me ;-) )
+
+A different TL technique is called \textit{off-the-shelf} (OTS) TL.
+It is similar to \textit{fine-tuning}, except that all but the final layer is kept frozen during training.
+\citeauthor{sharif2014cnn} (\citeyear{sharif2014cnn}) found that features extracted by a CNN pre-trained on ImageNet were general enough for OTS TL to work.
+More specifically, these OTS CNNs showed better performance than the (fully trained non-CNN) state of the art at the time.
+Besides reducing computational cost of training, a benefit of OTS TL is that it can prevent overfitting in cases where \textit{fine-tuning} would not \citep{yosinski2014transferable}.
+
+% OFF-THE-SHELF:
+% 1) "yosinski2014transferable": OTS can prevent overfitting where fine-tuning would overfit
+% 2) "yosinski2014transferable": first layers general information
 

### 2022-04-23 writing/tex/draft_introduction.tex
Almost there :-)

@@
 The output for a single position is a weighted sum of all $value$ vectors.
-Weights are obtained by combining this position's $query$ with all $key$ vectors\footnote{In principle by taking softmax over all query-key dot products.}.
+Weights are obtained by matching this position's $query$ with all $key$ vectors\footnote{In principle by taking softmax over all query-key inner products.}.
 As such, an attention head decides itself how much of each position should be incorporated in subsequent representations.
@@
 The example given in section \ref{vits} can be described as \textit{fine-tuning} TL.
-Here, a pre-trained model is used as a starting point, and trained fully on the target task -- possibly with different hyper-parameters, and often with a new final layer that fits the task.
-\citeauthor{yosinski2014transferable} (\citeyear{yosinski2014transferable}) studied transferability properties of CNNs pre-trained on ImageNet, and suggested that \textit{fine-tuning} results in improved performance and generalizability compared to training from scratch.
+Here, a pre-trained model is used as a starting point, and then trained on the target task -- possibly with different hyper-parameters, and often with a new final layer that fits the task.
+\citeauthor{yosinski2014transferable} (\citeyear{yosinski2014transferable}) studied transferability properties of CNNs pre-trained on ImageNet, and suggested that fine-tuning results in improved performance and generalizability compared to training from scratch.
 Moreover, while transferability decreased as source and target task became more dissimilar, utilizing TL still seemed to be preferred.
@@
 A different TL technique is called \textit{off-the-shelf} (OTS) TL.
-It is similar to \textit{fine-tuning}, except that all but the final layer is kept frozen during training.
+It is similar to fine-tuning, except that all but the final layer is kept frozen during training.
 \citeauthor{sharif2014cnn} (\citeyear{sharif2014cnn}) found that features extracted by a CNN pre-trained on ImageNet were general enough for OTS TL to work.
 More specifically, these OTS CNNs showed better performance than the (fully trained non-CNN) state of the art at the time.
-Besides reducing computational cost of training, a benefit of OTS TL is that it can prevent overfitting in cases where \textit{fine-tuning} would not \citep{yosinski2014transferable}.
+Besides reducing computational cost of training, a benefit of OTS TL is that it can prevent overfitting in cases where fine-tuning would not \citep{yosinski2014transferable}.
 
@@
 
-\subsection{The current study and related work}
-TODO
+\subsection{Related work and the current study}
+As mentioned, the paper that first introduced ViTs already demonstrated they benefit from TL \citep{dosovitskiy2020image}, and more has been discovered about ViTs' TL properties since then.
+\citeauthor{matsoukas2021time} (\citeyear{matsoukas2021time}) investigated whether TL allowed ViTs to replace CNNs in Medical Imaging.
+This domain is characterized by small datasets, so to overcome the inductive bias problem described earlier, TL is imperative.
+It was found that ViTs benefit more from fine-tuning TL than CNNs do, and that ViTs pre-trained on ImageNet perform on par CNNs.
+
+What \citeauthor{matsoukas2021time} did not investigate, is OTS performance of ViTs.
+This was, however, done in \citeauthor{zhou2021convnets} (\citeyear{zhou2021convnets}), which found that ViTs pre-trained on ImageNet had worse OTS performance than their CNN counterparts.
+An exception was the \textit{Swin-B} architecture, which was the best performing model on all datasets, and uses \textit{(shifted-)window-based} self-attention methods, to make Transformers better suited for CV tasks \citep{liu2021swin}.
+% Not into detail about ots and other techniques
+
+Another ``shortcomming'' of \citeauthor{matsoukas2021time} is that it merely examined whether ViTs can use TL to outperform CNNs in a domain that happens to be characterized by limited-sized datasets, but that it did not actively focus on the extent to which ViTs can be made useful in limited data/compute environments.
+More related to this is \citeauthor{touvron2021training} (\citeyear{touvron2021training}), which proposed transformer-specific distillation-based learning techniques to make ViTs applicable for smaller datasets.
+This is not specifically related to TL as defined in this paper, however.
+% matsoukas2021time not specifically focussed on seeing to what degree TL allows ViTs to dominate CNNs in limited compute/data environments

### 2022-04-24 writing/tex/draft_introduction.tex
Seems like the first draft of the introduction is almost done..

@@
 It is well established (e.g.
-\citeauthor{lecun1998gradient}, \citeyear{yosinski2014transferable}; \citeauthor{yosinski2014transferable}, \citeyear{lecun1998gradient}) that the first layers of a CNN detect low level features, such as edges and corners.
+\citeauthor{lecun1998gradient}, \citeyear{lecun1998gradient}; \citeauthor{yosinski2014transferable}, \citeyear{yosinski2014transferable}) that the first layers of a CNN detect low level features, such as edges and corners.
 These are combined in subsequent layers to detect higher order features.
@@
 \paragraph{Architecture}
-Where CNNs reduce the number of trainable weights by taking an image's topology into account, ViTs do this by learning themselves what information they should incorporate in successive representations of the input.
-As such, they are not restricted to a local receptive field, but can potentially include information from all over the previous layer.
-
 ViTs were first introduced in \citeauthor{dosovitskiy2020image} (\citeyear{dosovitskiy2020image}), and are based on the \textit{encoder stack} of the \textit{Transformer} architecture \citep{vaswani2017attention}.
-This stack takes a sequence of vectors as input, and is composed of alternating \textit{multi-headed self-attention} layers and position-wise MLP-type layers\footnote{position-wise meaning the MLP can take only a single position (constituent vector) as input at a time, and is shared between all positions.}.
+This stack takes a sequence of vectors as input, and is composed of alternating \textit{multi-headed self-attention} layers and position-wise MLP-type layers -- position-wise meaning the MLP can take only a single position (constituent vector) as input at a time, and is shared between all positions.
 
@@
 Learned embeddings convert these into a sequence of vectors, to which 1D positional encodings are added\footnote{Do mind that ViTs have no built-in knowledge of an image's 2D topology.}.
-A learned $[class]$ vector gets prepended to the sequence.
-Its representation at the end of the encoder stack is the final output, which can then be used for classification, segmentation, etc.
+A learned $[class]$ vector gets prepended to the sequence, and its representation at the end of the encoder stack is the final output, which can then be used for classification, segmentation, etc.
 
@@
 % While a CNN is restricted to only incorporate information from its surrounding pixels in a subsequent layer, a ViT
+
+\paragraph{CNNs versus ViTs}
+Where CNNs reduce the number of trainable weights by taking an image's topology into account, ViTs do this by learning themselves what information they should incorporate in successive representations of the input.
+As such, they are not restricted to a local receptive field, but can potentially include information from all over the previous layer.
+
+Moreover, weight sharing applies as much to ViTs as it does to CNNs.
+In ViTs, each position goes through the same computational pipelines of being multiplied by 3 projection matrices, followed by a weighted sum calculation, followed by being fed through an MLP-type layer\footnote{Neglecting skip connections and normalization, as this is a comparison on an intuitive level.}.
+In CNNs these pipelines take the form of convolution kernels.
 
@@
 More related to this is \citeauthor{touvron2021training} (\citeyear{touvron2021training}), which proposed transformer-specific distillation-based learning techniques to make ViTs applicable for smaller datasets.
-This is not specifically related to TL as defined in this paper, however.
+This is not necessarily related to TL as defined in this paper, however.
 % matsoukas2021time not specifically focussed on seeing to what degree TL allows ViTs to dominate CNNs in limited compute/data environments
+\\\\
+The current study limits itself to the task of identifying which materials were used in artistic content.
+For this, it will examine to what extent TL techniques (both fine-tuning and OTS) allow ViTs to become the preferred architecture when compute resources are limited, and attempt to find a methodology that is preferred in circumstances as described.
+It will do this by means of several experiments, and both qualitative and quantitative analysis of the results.

### 2022-04-25 writing/tex/draft_introduction.tex
Still have to give it a final read, but seems like my first draft is done now..

@@
 It is for this reason the paper that first introduced ViTs \citep{dosovitskiy2020image} mentions using pre-trained models rather than randomly initialised ones, as it allows general knowledge about images to be carried over to the new task.
-This technique falls within the category of TL, which will be cover below.
+This technique falls within the category of TL, which will be covered next.
 
@@
 \\\\
-The current study limits itself to the task of identifying which materials were used in artistic content.
+The current study will examine to what extent TL techniques (both fine-tuning and OTS) allow ViTs to become the preferred architecture when compute resources are limited, and attempt to find a methodology that is recommended in these circumstances.
+It will, however, restrict itself to a single CV task, namely material classification of artworks.
+Several experiments will be performed, of which results will be analysed both quantitatively as well as qualitatively.
+
+%The current study limits itself to the task of identifying which materials were used in artworks.
 For this, it will examine to what extent TL techniques (both fine-tuning and OTS) allow ViTs to become the preferred architecture when compute resources are limited, and attempt to find a methodology that is preferred in circumstances as described.
-It will do this by means of several experiments, and both qualitative and quantitative analysis of the results.
+It will do this by means of several experiments, and both quantitative and qualitative analysis of the results.
+
+More formally, the following research question is proposed: \quote{\textit{To what extent (and in what manners) does utilizing TL allow ViTs to perform on par with CNNs for material classification of artworks, when compute resources are limited to a single GPU/TPU?}}

### 2022-05-25 writing/tex/draft_introduction.tex
Ready to hand in my draft to Sabatelli.
bunch of things: merged some files; cleaned some stuff up and did preliminary experiments for fine tuning.

@@
 Since the introduction of AlexNet, roughly a decade ago, \textit{Convolutional Neural Networks} (CNNs) have played an important role in \textit{Computer Vision} (CV) \citep{krizhevsky2012imagened}.
-Recently, however, a new type of \textit{Artificial Neural Network} (ANN), called \textit{Vision Transformer} (ViT), gained state of the art performance on common learning benchmarks, including the ImageNet dataset\footnote{At the time of writing TODO is the best performing model, and the top TODO are all Transformer-based: (website)}.
+Recently, however, a new type of architecture, called \textit{Vision Transformer} (ViT), gained state of the art performance on common learning benchmarks, including the ImageNet dataset\footnote{At the time of writing TODO is the best performing model, and the top TODO are all Transformer-based: (website)}.
 
 Exciting as this may be, a plain ViT is not likely to become useful in circumstances with limited training data and/or only consumer-grade hardware available for learning.
-This is due to a limitation that is inherent to the ViT architecture, which will be elaborated on in section \ref{vits}.
+This is due to a limitation inherent to the ViT architecture, which will be elaborated on in section \ref{vits}.
 Consequently, the current paper investigates whether utilizing \textit{Transfer Learning} (TL) capabilities of ViTs can help overcome these limitations, and if this allows ViTs to become the preferred architecture in the aforementioned circumstances as well.
@@
 
+% TODO!!!
+Maybe mention TL is a solution when target dataset is much smaller, and maybe mention model TL technique: information passed between source and target in the form of a pre-trained model!!!
+TODO TODO TODO
 % OFF-THE-SHELF:

### 2022-05-25 writing/tex/draft_methods.tex
bunch of things: merged some files; cleaned some stuff up and did preliminary experiments for fine tuning.

\section{Methods}

% Experiments build upon Sabatelli et al.

\subsection{Dataset}
% Mention the challenges and show the table, of course. Also add max and min classes per category to the table maybe
% Don't forget to tell that images get scaled and cropped to 224 * 224 size and imagenet normalized
% Dataset described in more detail in Appendix A, including histograms and confusion matrices + balancing 'algorithm'

\subsection{Models}
% General information (4 per type, base/tiny etc)
% Paragraph/section about CNN models
%  - ResNet50 and VGG19 best fine-tuned and ots networks in Sabatelli
%  - ResNet50 I think also used in medical imaging paper; ResNet101 ots vit paper
%  * Convnext and EfficientNetV2 to have some more modern ones in the mix (to make comparison more fair)
% Paragraph/section about ViT models -- explain the idea behind each in 1 sentence maybe (also for CNNs)

\subsection{Hyperparameters}

\subsection{Hardware and software}

### 2022-05-25 ft_experiment.py

@@ -1,0 +1,53 @@
+import sys
+from torch import nn, optim
+import torch
+import rijks_torch.learning_problems as probs
+import rijks_torch.learning_problems.defaults as defs
+from rijks_torch.data_loading.rijksdataloaders import RijksDataloaders
+from rijks_torch.training import train, test
+
+
+def main():
+    """
+    Arguments given to script must be:\n
+    1) name of the experiment: will be prepended to all output files;\n
+    2) name of the model such that 'get_{name}_problem()' is defined;\n
+    3) dataset files. e.g. if /x/y/z is given, then
+        /x/y/z-train.csv,
+        /x/y/z-val.csv,
+        /x/y/z-test.csv,
+        /x/y/z-hist.csv,must exist;\n
+    4) name of directory containing all the jpg files (should be stored locally, so under /local)
+    """
+    
+    # If there is no gpu, we aren't actually going to run it..
+    assert torch.cuda.is_available(), "There was no GPU :-("
+
+    experiment_name = sys.argv[1]
+    model_name = sys.argv[2]
+    dataset_files = sys.argv[3]
+    dataset_jpg_dir = sys.argv[4]
+
+    # Creating the dataloaders from given arguments:
+    datloader = RijksDataloaders(
+        ds_name=dataset_files,
+        hist_path=dataset_files + "-hist.csv",
+        img_dir=dataset_jpg_dir,
+        transforms={"all": defs.buildTransform(imnet_norm=True)},
+        batch_size=64
+    )
+
+    # Get the model tailored to specification. Using getattr because function from cli args
+    model, dl = getattr(probs, f"get_{model_name}_problem")(off_the_shelf=False, dl=datloader)
+
+    loss = nn.CrossEntropyLoss()
+    optimizer = optim.Adam([param for param in model.parameters() if param.requires_grad == True], lr=5e-6)
+
+    # Training and validating (best model on val set returned):
+    model = train(model, dl, loss, optimizer, name=experiment_name)
+
+    # Testing model that performed best on validation set:
+    test(model, dl, loss, name=experiment_name)
+
+if __name__ == "__main__":
+    main()

### 2022-05-25 ft_experiment.sh

@@ -1,0 +1,24 @@
+#!/bin/bash
+
+#SBATCH --job-name=ft_tl_with_vits
+#SBATCH --mail-type=ALL
+#SBATCH --mail-user=v.tonkes@student.rug.nl
+#SBATCH --nodes=1
+#SBATCH --ntasks=1
+#SBATCH --cpus-per-task=8
+#SBATCH --time=02:00:00
+#SBATCH --partition=gpu
+#SBATCH --gres=gpu:v100:1
+#SBATCH --array=1-8
+
+module purge
+module load Python/3.9.6-GCCcore-11.2.0
+
+mkdir -p /local/tmp/dataset
+tar xzf /data/$USER/datasets/$2.tar.gz -C /local/tmp/dataset
+
+infile=ft_models.txt
+model=$(sed "${SLURM_ARRAY_TASK_ID}q;d" $infile)
+
+source /data/$USER/bachproj/bin/activate
+python ft_experiment.py $1_$model $model /local/tmp/dataset/fullsize /local/tmp/dataset

### 2022-05-25 rijks_torch/learning_problems/__init__.py

@@ -1,19 +1,20 @@
 # A bunch of functions stating learning problems:
 #    CNN based:
-from .vgg19 import get_vgg19_problem
-from .resnet50 import get_resnet50_problem
-from .resnet101 import get_resnet101_problem
-from .convnext_b import get_convnext_b_problem
-from .convnext_t import get_convnext_t_problem
-from .efficientnetv2_m import get_efficientnetv2_m_problem
-from .efficientnetv2_t import get_efficientnetv2_t_problem
+from .vgg import get_vgg19_problem
+from .resnet import get_resnet50_problem
+from .resnet import get_resnet101_problem
+from .convnext import get_convnext_b_problem
+from .convnext import get_convnext_s_problem
+from .convnext import get_convnext_t_problem
+from .efficientnetv2 import get_efficientnetv2_m_problem
+from .efficientnetv2 import get_efficientnetv2_t_problem
 
 #    ViT based:
-from .vit_b_32 import get_vit_b_32_problem
-from .vit_b_16 import get_vit_b_16_problem
-from .vit_t_16 import get_vit_t_16_problem
-from .swin_b import get_swin_b_problem
-from .swin_t import get_swin_t_problem
-from .beit_b_16 import get_beit_b_16_problem
-from .deit_b_16 import get_deit_b_16_problem
-from .deit_t_16 import get_deit_t_16_problem
+from .vit import get_vit_b_16_problem
+from .vit import get_vit_t_16_problem
+from .swin import get_swin_b_problem
+from .swin import get_swin_s_problem
+from .swin import get_swin_t_problem
+from .beit import get_beit_b_16_problem
+from .deit import get_deit_b_16_problem
+from .deit import get_deit_t_16_problem

### 2022-05-25 rijks_torch/learning_problems/convnext.py

@@ -1,0 +1,59 @@
+from .defaults import freezeLayers
+from torch import nn
+from torchvision import models
+
+def get_convnext_b_problem(off_the_shelf: bool, dl, pretrained: bool = True):
+    """
+    Returns the whole problem statement for training convnext_base on the Rijksdataset.
+    In other words: a pre-trained model (with the head replaced), and the dataloaders.\n
+    :off_the_shelf: says if it should freeze all but the new head for learning.\n
+    :dataloaders: allows user to specify custom dataset.\n
+    :pretrained: states if it should load a model pretrained om ImageNet.\n
+    """
+    model = models.convnext_base(pretrained=pretrained)
+    
+    # Prepare for off the shelf learning if needed:
+    freezeLayers(model, off_the_shelf)
+    
+    # Replace head with one that fits the task
+    model.classifier[2] = nn.Linear(1024, len(dl.materials))
+
+    return model, dl
+
+
+def get_convnext_s_problem(off_the_shelf: bool, dl, pretrained: bool = True):
+    """
+    Returns the whole problem statement for training convnext_small on the Rijksdataset.
+    In other words: a pre-trained model (with the head replaced), and the dataloaders.\n
+    :off_the_shelf: says if it should freeze all but the new head for learning.\n
+    :dataloaders: allows user to specify custom dataset.\n
+    :pretrained: states if it should load a model pretrained om ImageNet.\n
+    """
+    model = models.convnext_small(pretrained=pretrained)
+    
+    # Prepare for off the shelf learning if needed:
+    freezeLayers(model, off_the_shelf)
+    
+    # Replace head with one that fits the task
+    model.classifier[2] = nn.Linear(768, len(dl.materials))
+
+    return model, dl
+
+
+def get_convnext_t_problem(off_the_shelf: bool, dl, pretrained: bool = True):
+    """
+    Returns the whole problem statement for training convnext_tiny on the Rijksdataset.
+    In other words: a pre-trained model (with the head replaced), and the dataloaders.\n
+    :off_the_shelf: says if it should freeze all but the new head for learning.\n
+    :dataloaders: allows user to specify custom dataset.\n
+    :pretrained: states if it should load a model pretrained om ImageNet.\n
+    """
+    model = models.convnext_tiny(pretrained=pretrained)
+    
+    # Prepare for off the shelf learning if needed:
+    freezeLayers(model, off_the_shelf)
+    
+    # Replace head with one that fits the task
[…]

### 2022-05-25 rijks_torch/learning_problems/convnext_b.py

@@ -1,21 +1,0 @@
-from .defaults import freezeLayers
-from torch import nn
-from torchvision import models
-
-def get_convnext_b_problem(off_the_shelf: bool, dl, pretrained: bool = True):
-    """
-    Returns the whole problem statement for training convnext_base on the Rijksdataset.
-    In other words: a pre-trained model (with the head replaced), and the dataloaders.\n
-    :off_the_shelf: says if it should freeze all but the new head for learning.\n
-    :dataloaders: allows user to specify custom dataset.\n
-    :pretrained: states if it should load a model pretrained om ImageNet.\n
-    """
-    model = models.convnext_base(pretrained=pretrained)
-    
-    # Prepare for off the shelf learning if needed:
-    freezeLayers(model, off_the_shelf)
-    
-    # Replace head with one that fits the task
-    model.classifier[2] = nn.Linear(1024, len(dl.materials))
-
-    return model, dl

### 2022-05-25 rijks_torch/learning_problems/convnext_t.py

@@ -1,21 +1,0 @@
-from .defaults import freezeLayers
-from torch import nn
-from torchvision import models
-
-def get_convnext_t_problem(off_the_shelf: bool, dl, pretrained: bool = True):
-    """
-    Returns the whole problem statement for training convnext_tiny on the Rijksdataset.
-    In other words: a pre-trained model (with the head replaced), and the dataloaders.\n
-    :off_the_shelf: says if it should freeze all but the new head for learning.\n
-    :dataloaders: allows user to specify custom dataset.\n
-    :pretrained: states if it should load a model pretrained om ImageNet.\n
-    """
-    model = models.convnext_tiny(pretrained=pretrained)
-    
-    # Prepare for off the shelf learning if needed:
-    freezeLayers(model, off_the_shelf)
-    
-    # Replace head with one that fits the task
-    model.classifier[2] = nn.Linear(768, len(dl.materials))
-
-    return model, dl

### 2022-05-25 rijks_torch/learning_problems/defaults.py

@@ -29,6 +29,6 @@ def buildTransform(imnet_norm: bool, imsize: int = 224, extratransforms = None) 
         tfs += [transforms.Normalize([0.485, 0.456, 0.406], [0.229, 0.224, 0.225])]
     
     if extratransforms != None:
-        tfs += extratransforms
+        tfs = extratransforms + tfs
     
     return transforms.Compose(tfs)

### 2022-05-25 rijks_torch/learning_problems/deit.py

@@ -2,6 +2,25 @@ from .defaults import freezeLayers
 from torch import nn
 import timm
 
+def get_deit_b_16_problem(off_the_shelf: bool, dl, pretrained: bool = True):
+    """
+    Returns the whole problem statement for training deit_base_16 on the Rijksdataset.
+    In other words: a pre-trained model (with the head replaced), and the dataloaders.\n
+    :off_the_shelf: says if it should freeze all but the new head for learning.\n
+    :dataloaders: allows user to specify custom dataset.\n
+    :pretrained: states if it should load a model pretrained om ImageNet.\n
+    """
+    model = timm.create_model('deit_base_patch16_224', pretrained=pretrained)
+
+    # Prepare for off the shelf learning if needed:
+    freezeLayers(model, off_the_shelf)
+    
+    # Replace head with one that fits the task
+    model.head = nn.Linear(768, len(dl.materials))
+
+    return model, dl
+
+
 def get_deit_t_16_problem(off_the_shelf: bool, dl, pretrained: bool = True):
     """
     Returns the whole problem statement for training deit_tiny_16 on the Rijksdataset.

### 2022-05-25 rijks_torch/learning_problems/deit_b_16.py

@@ -1,21 +1,0 @@
-from .defaults import freezeLayers
-from torch import nn
-import timm
-
-def get_deit_b_16_problem(off_the_shelf: bool, dl, pretrained: bool = True):
-    """
-    Returns the whole problem statement for training deit_base_16 on the Rijksdataset.
-    In other words: a pre-trained model (with the head replaced), and the dataloaders.\n
-    :off_the_shelf: says if it should freeze all but the new head for learning.\n
-    :dataloaders: allows user to specify custom dataset.\n
-    :pretrained: states if it should load a model pretrained om ImageNet.\n
-    """
-    model = timm.create_model('deit_base_patch16_224', pretrained=pretrained)
-
-    # Prepare for off the shelf learning if needed:
-    freezeLayers(model, off_the_shelf)
-    
-    # Replace head with one that fits the task
-    model.head = nn.Linear(768, len(dl.materials))
-
-    return model, dl

### 2022-05-25 rijks_torch/learning_problems/efficientnetv2.py

@@ -2,6 +2,25 @@ from .defaults import freezeLayers
 from torch import nn
 import timm
 
+def get_efficientnetv2_m_problem(off_the_shelf: bool, dl, pretrained: bool = True):
+    """
+    Returns the whole problem statement for training efficientnetv2_m on the Rijksdataset.
+    In other words: a pre-trained model (with the head replaced), and the dataloaders.\n
+    :off_the_shelf: says if it should freeze all but the new head for learning.\n
+    :dataloaders: allows user to specify custom dataset.\n
+    :pretrained: states if it should load a model pretrained om ImageNet.\n
+    """
+    model = timm.create_model('efficientnetv2_rw_m', pretrained=pretrained)
+    
+    # Prepare for off the shelf learning if needed:
+    freezeLayers(model, off_the_shelf)
+    
+    # Replace head with one that fits the task
+    model.classifier = nn.Linear(2152, len(dl.materials))
+
+    return model, dl
+
+
 def get_efficientnetv2_t_problem(off_the_shelf: bool, dl, pretrained: bool = True):
     """
     Returns the whole problem statement for training efficientnetv2_t on the Rijksdataset.

### 2022-05-25 rijks_torch/learning_problems/efficientnetv2_m.py

@@ -1,21 +1,0 @@
-from .defaults import freezeLayers
-from torch import nn
-import timm
-
-def get_efficientnetv2_m_problem(off_the_shelf: bool, dl, pretrained: bool = True):
-    """
-    Returns the whole problem statement for training efficientnetv2_m on the Rijksdataset.
-    In other words: a pre-trained model (with the head replaced), and the dataloaders.\n
-    :off_the_shelf: says if it should freeze all but the new head for learning.\n
-    :dataloaders: allows user to specify custom dataset.\n
-    :pretrained: states if it should load a model pretrained om ImageNet.\n
-    """
-    model = timm.create_model('efficientnetv2_rw_m', pretrained=pretrained)
-    
-    # Prepare for off the shelf learning if needed:
-    freezeLayers(model, off_the_shelf)
-    
-    # Replace head with one that fits the task
-    model.classifier = nn.Linear(2152, len(dl.materials))
-
-    return model, dl

### 2022-05-25 rijks_torch/learning_problems/resnet.py

@@ -19,3 +19,22 @@ def get_resnet50_problem(off_the_shelf: bool, dl, pretrained: bool = True):
     model.fc = nn.Linear(2048, len(dl.materials))
 
     return model, dl
+
+
+def get_resnet101_problem(off_the_shelf: bool, dl, pretrained: bool = True):
+    """
+    Returns the whole problem statement for training resnet101 on the Rijksdataset.
+    In other words: a pre-trained model (with the head replaced), and the dataloaders.\n
+    :off_the_shelf: says if it should freeze all but the new head for learning.\n
+    :dataloaders: allows user to specify custom dataset.\n
+    :pretrained: states if it should load a model pretrained om ImageNet.\n
+    """
+    model = models.resnet101(pretrained=pretrained)
+    
+    # Prepare for off the shelf learning if needed:
+    freezeLayers(model, off_the_shelf)
+    
+    # Replace head with one that fits the task
+    model.fc = nn.Linear(2048, len(dl.materials))
+
+    return model, dl

### 2022-05-25 rijks_torch/learning_problems/resnet101.py

@@ -1,21 +1,0 @@
-from .defaults import freezeLayers
-from torch import nn
-from torchvision import models
-
-def get_resnet101_problem(off_the_shelf: bool, dl, pretrained: bool = True):
-    """
-    Returns the whole problem statement for training resnet101 on the Rijksdataset.
-    In other words: a pre-trained model (with the head replaced), and the dataloaders.\n
-    :off_the_shelf: says if it should freeze all but the new head for learning.\n
-    :dataloaders: allows user to specify custom dataset.\n
-    :pretrained: states if it should load a model pretrained om ImageNet.\n
-    """
-    model = models.resnet101(pretrained=pretrained)
-    
-    # Prepare for off the shelf learning if needed:
-    freezeLayers(model, off_the_shelf)
-    
-    # Replace head with one that fits the task
-    model.fc = nn.Linear(2048, len(dl.materials))
-
-    return model, dl

### 2022-05-25 rijks_torch/learning_problems/swin.py

@@ -1,0 +1,59 @@
+from .defaults import freezeLayers
+from torch import nn
+import timm
+
+def get_swin_b_problem(off_the_shelf: bool, dl, pretrained: bool = True):
+    """
+    Returns the whole problem statement for training swin_b on the Rijksdataset.
+    In other words: a pre-trained model (with the head replaced), and the dataloaders.\n
+    :off_the_shelf: says if it should freeze all but the new head for learning.\n
+    :dataloaders: allows user to specify custom dataset.\n
+    :pretrained: states if it should load a model pretrained om ImageNet.\n
+    """
+    model = timm.create_model('swin_base_patch4_window7_224', pretrained=pretrained)
+
+    # Prepare for off the shelf learning if needed:
+    freezeLayers(model, off_the_shelf)
+    
+    # Replace head with one that fits the task
+    model.head = nn.Linear(1024, len(dl.materials))
+
+    return model, dl
+
+
+def get_swin_s_problem(off_the_shelf: bool, dl, pretrained: bool = True):
+    """
+    Returns the whole problem statement for training swin_s on the Rijksdataset.
+    In other words: a pre-trained model (with the head replaced), and the dataloaders.\n
+    :off_the_shelf: says if it should freeze all but the new head for learning.\n
+    :dataloaders: allows user to specify custom dataset.\n
+    :pretrained: states if it should load a model pretrained om ImageNet.\n
+    """
+    model = timm.create_model('swin_small_patch4_window7_224', pretrained=pretrained)
+
+    # Prepare for off the shelf learning if needed:
+    freezeLayers(model, off_the_shelf)
+    
+    # Replace head with one that fits the task
+    model.head = nn.Linear(768, len(dl.materials))
+
+    return model, dl
+
+
+def get_swin_t_problem(off_the_shelf: bool, dl, pretrained: bool = True):
+    """
+    Returns the whole problem statement for training swin_t on the Rijksdataset.
+    In other words: a pre-trained model (with the head replaced), and the dataloaders.\n
+    :off_the_shelf: says if it should freeze all but the new head for learning.\n
+    :dataloaders: allows user to specify custom dataset.\n
+    :pretrained: states if it should load a model pretrained om ImageNet.\n
+    """
+    model = timm.create_model('swin_tiny_patch4_window7_224', pretrained=pretrained)
+
+    # Prepare for off the shelf learning if needed:
+    freezeLayers(model, off_the_shelf)
+    
[…]

### 2022-05-25 rijks_torch/learning_problems/swin_b.py

@@ -1,21 +1,0 @@
-from .defaults import freezeLayers
-from torch import nn
-import timm
-
-def get_swin_b_problem(off_the_shelf: bool, dl, pretrained: bool = True):
-    """
-    Returns the whole problem statement for training swin_b on the Rijksdataset.
-    In other words: a pre-trained model (with the head replaced), and the dataloaders.\n
-    :off_the_shelf: says if it should freeze all but the new head for learning.\n
-    :dataloaders: allows user to specify custom dataset.\n
-    :pretrained: states if it should load a model pretrained om ImageNet.\n
-    """
-    model = timm.create_model('swin_base_patch4_window7_224', pretrained=pretrained)
-
-    # Prepare for off the shelf learning if needed:
-    freezeLayers(model, off_the_shelf)
-    
-    # Replace head with one that fits the task
-    model.head = nn.Linear(1024, len(dl.materials))
-
-    return model, dl

### 2022-05-25 rijks_torch/learning_problems/swin_t.py

@@ -1,21 +1,0 @@
-from .defaults import freezeLayers
-from torch import nn
-import timm
-
-def get_swin_t_problem(off_the_shelf: bool, dl, pretrained: bool = True):
-    """
-    Returns the whole problem statement for training swin_t on the Rijksdataset.
-    In other words: a pre-trained model (with the head replaced), and the dataloaders.\n
-    :off_the_shelf: says if it should freeze all but the new head for learning.\n
-    :dataloaders: allows user to specify custom dataset.\n
-    :pretrained: states if it should load a model pretrained om ImageNet.\n
-    """
-    model = timm.create_model('swin_tiny_patch4_window7_224', pretrained=pretrained)
-
-    # Prepare for off the shelf learning if needed:
-    freezeLayers(model, off_the_shelf)
-    
-    # Replace head with one that fits the task
-    model.head = nn.Linear(768, len(dl.materials))
-
-    return model, dl

### 2022-05-25 rijks_torch/learning_problems/vit.py

@@ -1,6 +1,26 @@
 from .defaults import freezeLayers
 from torch import nn
+from torchvision import models
 import timm
+
+def get_vit_b_16_problem(off_the_shelf: bool, dl, pretrained: bool = True):
+    """
+    Returns the whole problem statement for training vit_b_16 on the Rijksdataset.
+    In other words: a pre-trained model (with the head replaced), and the dataloaders.\n
+    :off_the_shelf: says if it should freeze all but the new head for learning.\n
+    :dataloaders: allows user to specify custom dataset.\n
+    :pretrained: states if it should load a model pretrained om ImageNet.\n
+    """
+    model = models.vit_b_16(pretrained=pretrained)
+
+    # Prepare for off the shelf learning if needed:
+    freezeLayers(model, off_the_shelf)
+    
+    # Replace head with one that fits the task
+    model.heads.head = nn.Linear(768, len(dl.materials))
+
+    return model, dl
+
 
 def get_vit_t_16_problem(off_the_shelf: bool, dl, pretrained: bool = True):
     """

### 2022-05-25 rijks_torch/learning_problems/vit_b_16.py

@@ -1,21 +1,0 @@
-from .defaults import freezeLayers
-from torch import nn
-from torchvision import models
-
-def get_vit_b_16_problem(off_the_shelf: bool, dl, pretrained: bool = True):
-    """
-    Returns the whole problem statement for training vit_b_16 on the Rijksdataset.
-    In other words: a pre-trained model (with the head replaced), and the dataloaders.\n
-    :off_the_shelf: says if it should freeze all but the new head for learning.\n
-    :dataloaders: allows user to specify custom dataset.\n
-    :pretrained: states if it should load a model pretrained om ImageNet.\n
-    """
-    model = models.vit_b_16(pretrained=pretrained)
-
-    # Prepare for off the shelf learning if needed:
-    freezeLayers(model, off_the_shelf)
-    
-    # Replace head with one that fits the task
-    model.heads.head = nn.Linear(768, len(dl.materials))
-
-    return model, dl

### 2022-05-25 rijks_torch/learning_problems/vit_b_32.py

@@ -1,21 +1,0 @@
-from .defaults import freezeLayers
-from torch import nn
-from torchvision import models
-
-def get_vit_b_32_problem(off_the_shelf: bool, dl, pretrained: bool = True):
-    """
-    Returns the whole problem statement for training vit_b_32 on the Rijksdataset.
-    In other words: a pre-trained model (with the head replaced), and the dataloaders.\n
-    :off_the_shelf: says if it should freeze all but the new head for learning.\n
-    :dataloaders: allows user to specify custom dataset.\n
-    :pretrained: states if it should load a model pretrained om ImageNet.\n
-    """
-    model = models.vit_b_32(pretrained=pretrained)
-
-    # Prepare for off the shelf learning if needed:
-    freezeLayers(model, off_the_shelf)
-    
-    # Replace head with one that fits the task
-    model.heads.head = nn.Linear(768, len(dl.materials))
-
-    return model, dl

### 2022-05-25 rijks_torch/training/interactive_valplot.py

@@ -1,0 +1,29 @@
+import matplotlib.pyplot as plt
+from IPython import display
+from numpy import disp
+import pandas as pd
+
+def updatePlots(csv_file: str):
+    """"Show plots interactively: update after each validation run"""
+    display.clear_output()
+    df = pd.read_csv(csv_file)
+
+    epochs = [x + 1 for x in range(len(df))]
+    loss = df["mean_loss"].to_list()
+    acc = df["accuracy"].to_list()
+
+    # Give a horizontal line if only one datapoint available:
+    if len(epochs) == 1:
+        epochs += [1.1]
+        loss += loss
+        acc += acc
+
+    plt.plot(epochs, loss)
+    plt.xlabel("epochs")
+    plt.ylabel("loss")
+    plt.show()
+
+    plt.plot(epochs, acc)
+    plt.xlabel("epochs")
+    plt.ylabel("accuracy")
+    plt.show()

### 2022-05-26 ft_experiment.py
Made the code fine-tuning proof B-).. Added dropout layers to all tiny models (and added dropout version of non tiny ones if needed); added scheduler to the training loop; changed hyperparameters in the fine-tuning script.

@@ -1,5 +1,7 @@
+from sched import scheduler
 import sys
 from torch import nn, optim
+from torchvision import transforms
 import torch
 import rijks_torch.learning_problems as probs
 import rijks_torch.learning_problems.defaults as defs
@@ -33,18 +35,23 @@ def main():
         ds_name=dataset_files,
         hist_path=dataset_files + "-hist.csv",
         img_dir=dataset_jpg_dir,
-        transforms={"all": defs.buildTransform(imnet_norm=True)},
-        batch_size=64
+        transforms={"rest": defs.buildTransform(imnet_norm=True),
+                "train": defs.buildTransform(imnet_norm=True, extratransforms = [
+                    transforms.RandomRotation(10),
+                    transforms.RandomHorizontalFlip()
+                ])},
+        batch_size=32
     )
 
     # Get the model tailored to specification. Using getattr because function from cli args
     model, dl = getattr(probs, f"get_{model_name}_problem")(off_the_shelf=False, dl=datloader)
 
-    loss = nn.CrossEntropyLoss()
-    optimizer = optim.Adam([param for param in model.parameters() if param.requires_grad == True], lr=5e-6)
+    loss = nn.CrossEntropyLoss(label_smoothing=0.1)
+    optimizer = optim.Adam([param for param in model.parameters() if param.requires_grad == True], lr=1e-4)
+    scheduler = optim.lr_scheduler.ReduceLROnPlateau(optimizer, factor=0.5, patience=2, min_lr=5e-7)
 
     # Training and validating (best model on val set returned):
-    model = train(model, dl, loss, optimizer, name=experiment_name)
+    model = train(model, dl, loss, optimizer, scheduler=scheduler, early_stop=10, name=experiment_name)
 
     # Testing model that performed best on validation set:
     test(model, dl, loss, name=experiment_name)

### 2022-05-26 rijks_torch/learning_problems/__init__.py

@@ -1,8 +1,10 @@
 # A bunch of functions stating learning problems:
+
 #    CNN based:
 from .vgg import get_vgg19_problem
+from .vgg import get_vgg19_drop_problem
 from .resnet import get_resnet50_problem
-from .resnet import get_resnet101_problem
+from .resnet import get_resnet50_drop_problem
 from .convnext import get_convnext_b_problem
 from .convnext import get_convnext_s_problem
 from .convnext import get_convnext_t_problem
@@ -16,5 +18,6 @@ from .swin import get_swin_b_problem
 from .swin import get_swin_s_problem
 from .swin import get_swin_t_problem
 from .beit import get_beit_b_16_problem
+from .beit import get_beit_b_16_drop_problem
 from .deit import get_deit_b_16_problem
 from .deit import get_deit_t_16_problem

### 2022-05-26 rijks_torch/learning_problems/beit.py

@@ -19,3 +19,13 @@ def get_beit_b_16_problem(off_the_shelf: bool, dl, pretrained: bool = True):
     model.head = nn.Linear(768, len(dl.materials))
 
     return model, dl
+
+
+def get_beit_b_16_drop_problem(off_the_shelf: bool, dl, pretrained: bool = True):
+    """ Same but with a dropout layer. This version is used for fine tuning """
+    model, dl = get_beit_b_16_problem(off_the_shelf, dl, pretrained)
+    model.head = nn.Sequential(
+        nn.Dropout(p=0.2),
+        nn.Linear(768, len(dl.materials))
+    )
+    return model, dl

### 2022-05-26 rijks_torch/learning_problems/convnext.py

@@ -54,6 +54,9 @@ def get_convnext_t_problem(off_the_shelf: bool, dl, pretrained: bool = True):
     freezeLayers(model, off_the_shelf)
     
     # Replace head with one that fits the task
-    model.classifier[2] = nn.Linear(768, len(dl.materials))
+    model.classifier[2] = nn.Sequential(
+        nn.Dropout(p=0.2),
+        nn.Linear(768, len(dl.materials))
+    )
 
     return model, dl

### 2022-05-26 rijks_torch/learning_problems/defaults.py

@@ -15,7 +15,9 @@ def buildTransform(imnet_norm: bool, imsize: int = 224, extratransforms = None) 
     Builds a transform that modifies input images in a dataset.\n
     imnet_norm: if true, transforms.Normalize images with mean and std of ImageNet.\n
     imsize: required size for input images.\n
-    extratransforms gets appended to the returned transforms.Compose object.\n
+    extratransforms gets appended to the returned transforms.Compose object. Note that
+    these get applied last! (though for rotation it does not seem to matter that there
+    are black patches around).\n
     """
     if extratransforms != None and not isinstance(extratransforms, list):
         extratransforms = [extratransforms]
@@ -29,6 +31,7 @@ def buildTransform(imnet_norm: bool, imsize: int = 224, extratransforms = None) 
         tfs += [transforms.Normalize([0.485, 0.456, 0.406], [0.229, 0.224, 0.225])]
     
     if extratransforms != None:
-        tfs = extratransforms + tfs
+        # NOTE: these get applied last, so something like a rotation will show plack around it
+        tfs += extratransforms
     
     return transforms.Compose(tfs)

### 2022-05-26 rijks_torch/learning_problems/deit.py

@@ -35,6 +35,9 @@ def get_deit_t_16_problem(off_the_shelf: bool, dl, pretrained: bool = True):
     freezeLayers(model, off_the_shelf)
     
     # Replace head with one that fits the task
-    model.head = nn.Linear(192, len(dl.materials))
+    model.head = nn.Sequential(
+        nn.Dropout(p=0.2),
+        nn.Linear(192, len(dl.materials))
+    )
 
     return model, dl

### 2022-05-26 rijks_torch/learning_problems/efficientnetv2.py

@@ -35,6 +35,9 @@ def get_efficientnetv2_t_problem(off_the_shelf: bool, dl, pretrained: bool = Tru
     freezeLayers(model, off_the_shelf)
     
     # Replace head with one that fits the task
-    model.classifier = nn.Linear(1024, len(dl.materials))
+    model.classifier = nn.Sequential(
+        nn.Dropout(p=0.2),
+        nn.Linear(1024, len(dl.materials))
+    )
 
     return model, dl

### 2022-05-26 rijks_torch/learning_problems/resnet.py

@@ -21,20 +21,11 @@ def get_resnet50_problem(off_the_shelf: bool, dl, pretrained: bool = True):
     return model, dl
 
 
-def get_resnet101_problem(off_the_shelf: bool, dl, pretrained: bool = True):
-    """
-    Returns the whole problem statement for training resnet101 on the Rijksdataset.
-    In other words: a pre-trained model (with the head replaced), and the dataloaders.\n
-    :off_the_shelf: says if it should freeze all but the new head for learning.\n
-    :dataloaders: allows user to specify custom dataset.\n
-    :pretrained: states if it should load a model pretrained om ImageNet.\n
-    """
-    model = models.resnet101(pretrained=pretrained)
-    
-    # Prepare for off the shelf learning if needed:
-    freezeLayers(model, off_the_shelf)
-    
-    # Replace head with one that fits the task
-    model.fc = nn.Linear(2048, len(dl.materials))
-
+def get_resnet50_drop_problem(off_the_shelf: bool, dl, pretrained: bool = True):
+    """ Same but with a dropout layer. This version is used for fine tuning """
+    model, dl = get_resnet50_problem(off_the_shelf, dl, pretrained)
+    model.fc = nn.Sequential(
+        nn.Dropout(p=0.2),
+        nn.Linear(2048, len(dl.materials))
+    )
     return model, dl

### 2022-05-26 rijks_torch/learning_problems/swin.py

@@ -54,6 +54,9 @@ def get_swin_t_problem(off_the_shelf: bool, dl, pretrained: bool = True):
     freezeLayers(model, off_the_shelf)
     
     # Replace head with one that fits the task
-    model.head = nn.Linear(768, len(dl.materials))
+    model.head = nn.Sequential(
+        nn.Dropout(p=0.2),
+        nn.Linear(768, len(dl.materials))
+    )
 
     return model, dl

### 2022-05-26 rijks_torch/learning_problems/vgg.py

@@ -19,3 +19,13 @@ def get_vgg19_problem(off_the_shelf: bool, dl, pretrained: bool = True):
     model.classifier[6] = nn.Linear(4096, len(dl.materials))
 
     return model, dl
+
+
+def get_vgg19_drop_problem(off_the_shelf: bool, dl, pretrained: bool = True):
+    """ Same as 'get_vgg19_problem' but with a dropout layer. This version is used for fine tuning """
+    model, dl = get_vgg19_problem(off_the_shelf, dl, pretrained)
+    model.classifier[6] = nn.Sequential(
+        nn.Dropout(p=0.2),
+        nn.Linear(4096, len(dl.materials))
+    )
+    return model, dl

### 2022-05-26 rijks_torch/learning_problems/vit.py

@@ -36,6 +36,9 @@ def get_vit_t_16_problem(off_the_shelf: bool, dl, pretrained: bool = True):
     freezeLayers(model, off_the_shelf)
     
     # Replace head with one that fits the task
-    model.head = nn.Linear(192, len(dl.materials))
+    model.head = nn.Sequential(
+        nn.Dropout(p=0.2),
+        nn.Linear(192, len(dl.materials))
+    )
 
     return model, dl

### 2022-05-26 rijks_torch/training/train.py

@@ -1,12 +1,18 @@
 import torch
-from torch import nn
+from torch import nn, optim
 from torch.utils.data import DataLoader
 from ..data_loading import RijksDataloaders
 
 from time import time
 from copy import deepcopy
 
-def train(model: nn.Module, dataloaders: RijksDataloaders, lossfunc, optimizer, max_epochs: int = 200, early_stop: int = 7, name: str = "default"):
+def train(model: nn.Module,
+          dataloaders: RijksDataloaders,
+          lossfunc, optimizer, max_epochs:
+          int = 200,
+          early_stop: int = 7,
+          scheduler: optim.lr_scheduler.ReduceLROnPlateau = None,
+          name: str = "default"):
     """
     ## Function for training of a model
 
@@ -16,6 +22,7 @@ def train(model: nn.Module, dataloaders: RijksDataloaders, lossfunc, optimizer, 
     optimizer is the used optimizer.\n
     max_epochs maximum number of epochs if no early stopping occured\n
     early_stop if no new best is found after this many trials, training stops\n
+    scheduler dynamically changes the learning rate\n
     name is the name of the experiment. Gets prepended to output filenames\n
     ### Returns the model that scored best on the validation set!\n
     ### Also saves validation statistics to {name}-validation.csv
@@ -41,6 +48,10 @@ def train(model: nn.Module, dataloaders: RijksDataloaders, lossfunc, optimizer, 
     for epoch in range(max_epochs):
         train_loop(model, dataloaders.train, lossfunc, optimizer, device)
         loss = validation_loop(model, dataloaders.val, lossfunc, device, validation_log)
+        
+        # Scheduler will always be ReduceLROnPlateau, so needs loss as input
+        if scheduler != None:
+            scheduler.step(loss)
 
         # Save best model and early stop if no new best for too long:
         if loss < best_loss:

### 2022-06-01 ft_experiment.py
Working on qualitative analysis now

@@ -48,7 +48,7 @@ def main():
 
     loss = nn.CrossEntropyLoss(label_smoothing=0.1)
     optimizer = optim.Adam([param for param in model.parameters() if param.requires_grad == True], lr=1e-4)
-    scheduler = optim.lr_scheduler.ReduceLROnPlateau(optimizer, factor=0.5, patience=2, min_lr=5e-7)
+    scheduler = optim.lr_scheduler.ReduceLROnPlateau(optimizer, factor=0.1, patience=2, min_lr=5e-7)
 
     # Training and validating (best model on val set returned):
     model = train(model, dl, loss, optimizer, scheduler=scheduler, early_stop=10, name=experiment_name)

### 2022-06-01 rijks_torch/learning_problems/__init__.py

@@ -6,18 +6,23 @@ from .vgg import get_vgg19_drop_problem
 from .resnet import get_resnet50_problem
 from .resnet import get_resnet50_drop_problem
 from .convnext import get_convnext_b_problem
+from .convnext import get_convnext_b_drop_problem
 from .convnext import get_convnext_s_problem
 from .convnext import get_convnext_t_problem
 from .efficientnetv2 import get_efficientnetv2_m_problem
+from .efficientnetv2 import get_efficientnetv2_m_drop_problem
 from .efficientnetv2 import get_efficientnetv2_t_problem
 
 #    ViT based:
 from .vit import get_vit_b_16_problem
+from .vit import get_vit_b_16_drop_problem
 from .vit import get_vit_t_16_problem
 from .swin import get_swin_b_problem
+from .swin import get_swin_b_drop_problem
 from .swin import get_swin_s_problem
 from .swin import get_swin_t_problem
 from .beit import get_beit_b_16_problem
 from .beit import get_beit_b_16_drop_problem
 from .deit import get_deit_b_16_problem
+from .deit import get_deit_b_16_drop_problem
 from .deit import get_deit_t_16_problem

### 2022-06-01 rijks_torch/learning_problems/convnext.py

@@ -21,6 +21,16 @@ def get_convnext_b_problem(off_the_shelf: bool, dl, pretrained: bool = True):
     return model, dl
 
 
+def get_convnext_b_drop_problem(off_the_shelf: bool, dl, pretrained: bool = True):
+    """ Same but with a dropout layer. This version is used for fine tuning """
+    model, dl = get_convnext_b_problem(off_the_shelf, dl, pretrained)
+    model.classifier[2] = nn.Sequential(
+        nn.Dropout(p=0.2),
+        nn.Linear(1024, len(dl.materials))
+    )
+    return model, dl
+
+
 def get_convnext_s_problem(off_the_shelf: bool, dl, pretrained: bool = True):
     """
     Returns the whole problem statement for training convnext_small on the Rijksdataset.

### 2022-06-01 rijks_torch/learning_problems/deit.py

@@ -21,6 +21,16 @@ def get_deit_b_16_problem(off_the_shelf: bool, dl, pretrained: bool = True):
     return model, dl
 
 
+def get_deit_b_16_drop_problem(off_the_shelf: bool, dl, pretrained: bool = True):
+    """ Same but with a dropout layer. This version is used for fine tuning """
+    model, dl = get_deit_b_16_problem(off_the_shelf, dl, pretrained)
+    model.head = nn.Sequential(
+        nn.Dropout(p=0.2),
+        nn.Linear(768, len(dl.materials))
+    )
+    return model, dl
+
+
 def get_deit_t_16_problem(off_the_shelf: bool, dl, pretrained: bool = True):
     """
     Returns the whole problem statement for training deit_tiny_16 on the Rijksdataset.

### 2022-06-01 rijks_torch/learning_problems/efficientnetv2.py

@@ -21,6 +21,16 @@ def get_efficientnetv2_m_problem(off_the_shelf: bool, dl, pretrained: bool = Tru
     return model, dl
 
 
+def get_efficientnetv2_m_drop_problem(off_the_shelf: bool, dl, pretrained: bool = True):
+    """ Same but with a dropout layer. This version is used for fine tuning """
+    model, dl = get_efficientnetv2_m_problem(off_the_shelf, dl, pretrained)
+    model.classifier = nn.Sequential(
+        nn.Dropout(p=0.2),
+        nn.Linear(2152, len(dl.materials))
+    )
+    return model, dl
+
+
 def get_efficientnetv2_t_problem(off_the_shelf: bool, dl, pretrained: bool = True):
     """
     Returns the whole problem statement for training efficientnetv2_t on the Rijksdataset.

### 2022-06-01 rijks_torch/learning_problems/swin.py

@@ -21,6 +21,16 @@ def get_swin_b_problem(off_the_shelf: bool, dl, pretrained: bool = True):
     return model, dl
 
 
+def get_swin_b_drop_problem(off_the_shelf: bool, dl, pretrained: bool = True):
+    """ Same but with a dropout layer. This version is used for fine tuning """
+    model, dl = get_swin_b_problem(off_the_shelf, dl, pretrained)
+    model.head = nn.Sequential(
+        nn.Dropout(p=0.2),
+        nn.Linear(1024, len(dl.materials))
+    )
+    return model, dl
+
+
 def get_swin_s_problem(off_the_shelf: bool, dl, pretrained: bool = True):
     """
     Returns the whole problem statement for training swin_s on the Rijksdataset.

### 2022-06-01 rijks_torch/learning_problems/vit.py

@@ -22,6 +22,16 @@ def get_vit_b_16_problem(off_the_shelf: bool, dl, pretrained: bool = True):
     return model, dl
 
 
+def get_vit_b_16_drop_problem(off_the_shelf: bool, dl, pretrained: bool = True):
+    """ Same but with a dropout layer. This version is used for fine tuning """
+    model, dl = get_vit_b_16_problem(off_the_shelf, dl, pretrained)
+    model.heads.head = nn.Sequential(
+        nn.Dropout(p=0.2),
+        nn.Linear(768, len(dl.materials))
+    )
+    return model, dl
+
+
 def get_vit_t_16_problem(off_the_shelf: bool, dl, pretrained: bool = True):
     """
     Returns the whole problem statement for training vit_t_16 on the Rijksdataset.

### 2022-06-01 rijks_torch/training/train.py

@@ -1,6 +1,7 @@
 import torch
 from torch import nn, optim
 from torch.utils.data import DataLoader
+from torch.cuda.amp import autocast, GradScaler
 from ..data_loading import RijksDataloaders
 
 from time import time
@@ -73,18 +74,23 @@ def train(model: nn.Module,
 def train_loop(model: nn.Module, dataloader: DataLoader, lossfunc, optimizer, device):
     model.train()
 
+    # Scalar for mixed precision computations:
+    scalar = GradScaler()
+
     for batchnum, (x, y) in enumerate(dataloader):
         x = x.to(device)
         y = y.to(device)
 
         # Forward pass:
-        pred = model(x)
-        loss = lossfunc(pred, y)
+        with autocast(): # Taking advantage of mixed precision speedup (on v100)
+            pred = model(x)
+            loss = lossfunc(pred, y)
 
         # Backward pass:
-        optimizer.zero_grad()
-        loss.backward()
-        optimizer.step()
+        optimizer.zero_grad(set_to_none=True)
+        scalar.scale(loss).backward()
+        scalar.step(optimizer)
+        scalar.update()
 
 def validation_loop(model: nn.Module, dataloader: DataLoader, lossfunc, device, validation_csv):
     model.eval()

### 2022-06-06 result_analysis/qualitative/all_attention_layers.py
Further progress with qualitative analysis

@@ -1,0 +1,67 @@
+import pandas as pd
+import numpy as np
+from torchvision import io
+import cv2
+
+import os
+import sys
+sys.path.insert(1, "/home/vincent/Documenten/BachelorsProject/GitHub_Repo/")
+from rijks_torch.learning_problems.defaults import buildTransform
+
+models = [
+    (
+        "get_swin_attention_layers.py",
+        "/home/vincent/Documenten/BachelorsProject/best_models/ft_type1_swin_b_drop-best.pth",
+        "/home/vincent/Documenten/BachelorsProject/Rijksdata/type/1/fullsize",
+        False # Not off the shelf
+    ),
+    (
+        "get_swin_attention_layers.py",
+        "/home/vincent/Documenten/BachelorsProject/best_models/ots_type2_swin_b-best.pth",
+        "/home/vincent/Documenten/BachelorsProject/Rijksdata/type/2/fullsize",
+        True # Is off the shelf
+    ),
+    (
+        "get_deit_attention_layers.py",
+        "/home/vincent/Documenten/BachelorsProject/best_models/ft_type3_deit_b_16_drop-best.pth",
+        "/home/vincent/Documenten/BachelorsProject/Rijksdata/type/3/fullsize",
+        False # Not off the shelf
+    ),
+    (
+        "get_deit_attention_layers.py",
+        "/home/vincent/Documenten/BachelorsProject/best_models/ots_type3_deit_b_16-best.pth",
+        "/home/vincent/Documenten/BachelorsProject/Rijksdata/type/3/fullsize",
+        True # Is off the shelf
+    )
+]
+
+image_path = "/home/vincent/Documenten/BachelorsProject/Rijksdata/jpg/"
+
+images_csv = sys.argv[1]
+target_dir = sys.argv[2]
+
+df = pd.read_csv(images_csv)
+
+for idx, img in enumerate(df["jpg"]):
+    print(f"---Doing image {idx}" + 40 * "-")
+
+    # Saving both a colored and grayscale version of the input image
+    image = io.read_image(image_path + img, mode = io.ImageReadMode.RGB).float() / 255
+    image = buildTransform(imnet_norm=False)(image).unsqueeze(0)[0]
+    image = image.detach().numpy().transpose(1,2,0)
+    np.save(os.path.join(target_dir, f"img{idx:03}"), image)
+    image = cv2.cvtColor(image, cv2.COLOR_RGB2GRAY)
+    np.save(os.path.join(target_dir, f"img{idx:03}_gray"), image)
+
[…]

### 2022-06-06 result_analysis/qualitative/get_deit_attention_layers.py

@@ -1,0 +1,55 @@
+import sys
+sys.path.insert(1, "/home/vincent/Documenten/BachelorsProject/GitHub_Repo/")
+import rijks_torch.learning_problems as lp
+import rijks_torch.learning_problems.defaults as defs
+from rijks_torch.data_loading import RijksDataloaders
+
+import numpy as np
+import torch
+from torchvision import io
+import argparse
+
+
+arg_parser = argparse.ArgumentParser(description="Export all Swin attention layers for given model file (.pth) and image")
+arg_parser.add_argument("--state_dict", type=str, required=True)
+arg_parser.add_argument("--dataset", type=str, required=True)
+arg_parser.add_argument("--image", type=str, required=True)
+arg_parser.add_argument("--target", type=str, required=True)
+arg_parser.add_argument("--ots", action="store_true")
+args = arg_parser.parse_args()
+
+# Getting dataset and model:
+datloader = RijksDataloaders(
+    ds_name=args.dataset,
+    hist_path=args.dataset + "-hist.csv",
+    img_dir="/home/vincent/Documenten/BachelorsProject/Rijksdata/jpg/",
+    transforms={"all": defs.buildTransform(imnet_norm=True)},
+    batch_size=1
+)
+getmodel_fun = lp.get_deit_b_16_problem if args.ots else lp.get_deit_b_16_drop_problem
+model, dl = getmodel_fun(off_the_shelf=False, pretrained=False, dl=datloader)
+model.load_state_dict(torch.load(args.state_dict, map_location=torch.device('cpu')))
+model = model.eval()
+
+#---GETTING-ATTENTION-WEIGHTS-AND-SAVING-TO-FILE:----------------------------------
+f = open(args.target + "_layers.npy", "wb")
+def add_to_att_weights(module, inp, outp):
+    print(".", end="", flush=True)
+    np.save(f, outp.detach().numpy()[0].mean(axis=0))
+
+for name, module in model.named_modules():
+    if "attn_drop" in name: # A timm thing. This only works for timm models ;-)
+        module.register_forward_hook(add_to_att_weights)
+
+# Getting the image and passing it through the model:
+image = io.read_image(args.image, mode = io.ImageReadMode.RGB).float() / 255
+image = defs.buildTransform(imnet_norm=True)(image).unsqueeze(0)
+prediction = model(image)
+f.close()
+
+# Saving prediction to file
+prediction = dl.materials[torch.argmax(prediction)]
[…]

### 2022-06-06 result_analysis/qualitative/get_swin_attention_layers.py

@@ -1,0 +1,112 @@
+import sys
+sys.path.insert(1, "/home/vincent/Documenten/BachelorsProject/GitHub_Repo/")
+import rijks_torch.learning_problems as lp
+import rijks_torch.learning_problems.defaults as defs
+from rijks_torch.data_loading import RijksDataloaders
+
+import numpy as np
+import torch
+from torchvision import io
+import argparse
+
+
+arg_parser = argparse.ArgumentParser(description="Export all Swin attention layers for given model file (.pth) and image")
+arg_parser.add_argument("--state_dict", type=str, required=True)
+arg_parser.add_argument("--dataset", type=str, required=True)
+arg_parser.add_argument("--image", type=str, required=True)
+arg_parser.add_argument("--target", type=str, required=True)
+arg_parser.add_argument("--ots", action="store_true")
+args = arg_parser.parse_args()
+
+# Getting dataset and model:
+datloader = RijksDataloaders(
+    ds_name=args.dataset,
+    hist_path=args.dataset + "-hist.csv",
+    img_dir="/home/vincent/Documenten/BachelorsProject/Rijksdata/jpg/",
+    transforms={"all": defs.buildTransform(imnet_norm=True)},
+    batch_size=1
+)
+getmodel_fun = lp.get_swin_b_problem if args.ots else lp.get_swin_b_drop_problem
+model, dl = getmodel_fun(off_the_shelf=False, pretrained=False, dl=datloader)
+model.load_state_dict(torch.load(args.state_dict, map_location=torch.device('cpu')))
+model = model.eval()
+
+#---GETTING-ATTENTION-WEIGHTS:-----------------------------------------------------
+att_weighs = []
+def add_to_att_weights(module, inp, outp):
+    global att_weighs
+    # Taking the mean of all attention heads, but could also be max or min
+    att_weighs += [outp.detach().numpy()]
+
+for name, module in model.named_modules():
+    if "attn_drop" in name: # A timm thing. This only works for timm models ;-)
+        module.register_forward_hook(add_to_att_weights)
+
+# Getting the image and passing it through the model:
+image = io.read_image(args.image, mode = io.ImageReadMode.RGB).float() / 255
+image = defs.buildTransform(imnet_norm=True)(image).unsqueeze(0)
+prediction = model(image)
[…]

### 2022-06-08 writing/report.tex
Just went further with writing the introduction. Randomly pushing it now because haven't done that for some time..
At the point where all preliminaries are explained (CNNs, ViTs, TL). Up next: related work and description of current research
Still have to give it a final read, but seems like my first draft is done now..
Working on qualitative analysis now
Added all graphics I have currently to the paper. Qualitative analysis results still need to be included.

@@
 \title{\vspace{-0.5in}{\bfseries\scshape 
-    Transfer Learning Using Vision Transformers Project (No Title Yet)
+    Transfer Learning for computationally limited training of Vision Transformers on art classification tasks
     } \\ 
@@
 \author{
-    \normalsize{Vincent Tonkes, s3617157, v.tonkes@student.rug.nl,} \\ 
-    \normalsize{Supervisor: Dr M. Sabatelli}
+    \normalsize{V. Tonkes, s3617157, v.tonkes@student.rug.nl,} \\ 
+    \normalsize{Supervisor: Dr. M. Sabatelli}
 }
@@
 \input{tex/draft_introduction.tex}
+\input{tex/draft_methods.tex}
+\input{tex/draft_results.tex}
 
-\bibliographystyle{plainnat}
+\bibliographystyle{apacite}
 \bibliography{literature}

### 2022-06-08 writing/tex/draft_methods.tex
Added all graphics I have currently to the paper. Qualitative analysis results still need to be included.

@@
 % Experiments build upon Sabatelli et al.
+Most of the experiments and analysis discussed below build upon the work done in \citeauthor{sabatelli2018deep} (\citeyear{sabatelli2018deep}).
 
@@
 % Mention the challenges and show the table, of course. Also add max and min classes per category to the table maybe
-% Don't forget to tell that images get scaled and cropped to 224 * 224 size and imagenet normalized
+% Don't forget to tell that images get scaled and cropped to 224 * 224 size and imagenet normalised
 % Dataset described in more detail in Appendix A, including histograms and confusion matrices + balancing 'algorithm'
+
+\begin{table*}[tb]
+\centering
+\small
+\begin{tabular}{lllll}
+\hline
+\textbf{Task} & \textbf{\# Samples} & \textbf{\# Classes} & \textbf{Gini coefficient} & \textbf{Sample overlap} \\ \hline
+Type classification & 9607 (77628) & 30 (801) & 0.466 (0.974) & 0.686 \\
+Material classification & 7788 (96583) & 30 (136) & 0.563 (0.980) & 0.798 \\
+Artist classification & 6530 (38296) & 30 (8592) & 0.236 (0.676) & 1 \\
+Scaling experiment & 7926 (77628) & 15 (801) & 0.300 (0.974) & - \\ \hline
+\end{tabular}
+\caption{Overview of the used datasets.
+Values between brackets show the situation before balancing operations were performed.
+`Sample overlap' gives the average overlap between 2 of the 5 randomly generated sets per task ($i$ and $j$ where $i \neq j$).}
+\label{methods:datasets}
+\end{table*}
 

### 2022-06-08 writing/tex/draft_results.tex
Added all graphics I have currently to the paper. Qualitative analysis results still need to be included.

\section{Results}

\begin{figure*}
    \centering
    \begin{subfigure}{0.32\textwidth}
    \def\svgwidth{5.5cm}
    \input{img/ots_type_acc.pdf_tex}
    \caption{Type}
    \label{results:img:ots_type}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.32\textwidth}
    \def\svgwidth{5.5cm}
    \input{img/ots_mat_acc.pdf_tex}
    \caption{Material}
    \label{results:img:ots_mat}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.32\textwidth}
    \def\svgwidth{5.5cm}
    \input{img/ots_artist_acc.pdf_tex}
    \caption{Artist}
    \label{results:img:ots_artist}
    \end{subfigure}
    \caption{Validation accuracy using off the shelf models.}
\end{figure*}

\begin{table}[]
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lll}
\hline
\textbf{Model} & \textbf{Accuracy} & \textbf{Balanced accuracy} \\ \hline
\textbf{vit\_b\_16} & 86.06\% (${\pm 1.06\%}$) & 84.13\% (${\pm 1.57\%}$)  \\
\textbf{swin\_b} & \cellcolor{bestcol}89.43\% (${\pm 0.93\%}$) & \cellcolor{bestcol}87.47\% (${\pm 1.02\%}$)  \\
\textbf{beit\_b\_16} & \cellcolor{worstcol}82.26\% (${\pm 0.72\%}$) & \cellcolor{worstcol}77.75\% (${\pm 0.27\%}$)  \\
\textbf{deit\_b\_16} & 88.18\% (${\pm 0.66\%}$) & 85.36\% (${\pm 0.54\%}$)  \\\hdashline
\textbf{vgg19} & 83.93\% (${\pm 0.72\%}$) & 83.35\% (${\pm 0.81\%}$)  \\
\textbf{resnet50} & 85.51\% (${\pm 0.64\%}$) & 82.33\% (${\pm 1.85\%}$)  \\
\textbf{efficientnetv2\_m} & 83.41\% (${\pm 0.76\%}$) & 82.05\% (${\pm 1.25\%}$)  \\
\textbf{convnext\_b} & \cellcolor{secondbestcol}89.19\% (${\pm 0.64\%}$) & \cellcolor{secondbestcol}86.95\% (${\pm 1.38\%}$)  \\\hline
\end{tabular}
}
\caption{Type classification off the shelf}
\label{results:tab:ots_type}
\end{table}

\begin{table}[]
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lll}
\hline
\textbf{Model} & \textbf{Accuracy} & \textbf{Balanced accuracy} \\ \hline
\textbf{vit\_b\_16} & 81.78\% (${\pm 0.48\%}$) & 67.38\% (${\pm 1.37\%}$)  \\
\textbf{swin\_b} & \cellcolor{bestcol}85.87\% (${\pm 0.35\%}$) & \cellcolor{bestcol}71.19\% (${\pm 1.60\%}$)  \\
\textbf{beit\_b\_16} & 76.87\% (${\pm 0.96\%}$) & 60.16\% (${\pm 1.56\%}$)  \\
\textbf{deit\_b\_16} & 82.80\% (${\pm 1.12\%}$) & 66.46\% (${\pm 1.03\%}$)  \\\hdashline
\textbf{vgg19} & 76.87\% (${\pm 0.44\%}$) & 61.39\% (${\pm 1.47\%}$)  \\
\textbf{resnet50} & 80.99\% (${\pm 0.82\%}$) & 65.51\% (${\pm 0.93\%}$)  \\
\textbf{efficientnetv2\_m} & \cellcolor{worstcol}75.96\% (${\pm 1.24\%}$) & \cellcolor{worstcol}59.15\% (${\pm 1.24\%}$)  \\
\textbf{convnext\_b} & \cellcolor{secondbestcol}84.14\% (${\pm 0.92\%}$) & \cellcolor{secondbestcol}69.10\% (${\pm 1.05\%}$)  \\\hline
\end{tabular}
}
\caption{Material classification off the shelf}
\label{results:tab:ots_mat}
\end{table}

\begin{table}[]
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lll}
\hline
\textbf{Model} & \textbf{Accuracy} & \textbf{Balanced accuracy} \\ \hline
\textbf{vit\_b\_16} & 84.89\% (${\pm 0.46\%}$) & 81.42\% (${\pm 0.42\%}$)  \\
\textbf{swin\_b} & \cellcolor{bestcol}90.40\% (${\pm 0.65\%}$) & \cellcolor{bestcol}88.64\% (${\pm 0.78\%}$)  \\
\textbf{beit\_b\_16} & 79.70\% (${\pm 0.69\%}$) & 75.35\% (${\pm 1.09\%}$)  \\
\textbf{deit\_b\_16} & 88.13\% (${\pm 0.76\%}$) & 85.62\% (${\pm 0.87\%}$)  \\\hdashline
\textbf{vgg19} & 82.01\% (${\pm 0.66\%}$) & 78.10\% (${\pm 0.77\%}$)  \\
\textbf{resnet50} & 87.71\% (${\pm 1.06\%}$) & 85.12\% (${\pm 1.34\%}$)  \\
\textbf{efficientnetv2\_m} & \cellcolor{worstcol}78.62\% (${\pm 1.07\%}$) & \cellcolor{worstcol}73.92\% (${\pm 0.96\%}$)  \\
\textbf{convnext\_b} & \cellcolor{secondbestcol}90.13\% (${\pm 0.94\%}$) & \cellcolor{secondbestcol}87.84\% (${\pm 1.07\%}$)  \\\hline
\end{tabular}
}
\caption{Artist classification off the shelf}
\label{results:tab:ots_artist}
\end{table}

\begin{figure*}
    \centering
    \begin{subfigure}{0.32\textwidth}
    \def\svgwidth{5.5cm}
    \input{img/ft_type_acc.pdf_tex}
    \caption{Type}
    \label{results:img:ft_type}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.32\textwidth}
    \def\svgwidth{5.5cm}
    \input{img/ft_mat_acc.pdf_tex}
    \caption{Material}
    \label{results:img:ft_mat}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.32\textwidth}
    \def\svgwidth{5.5cm}
    \input{img/ft_artist_acc.pdf_tex}
    \caption{Artist}
    \label{results:img:ft_artist}
    \end{subfigure}
    \caption{Validation accuracy using fine-tuned models.}
\end{figure*}

\begin{table}[]
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lll}
\hline
\textbf{Model} & \textbf{Accuracy} & \textbf{Balanced accuracy} \\ \hline
\textbf{vit\_b\_16} & \cellcolor{worstcol}90.11\% (${\pm 0.35\%}$) & 87.40\% (${\pm 0.29\%}$)  \\
\textbf{swin\_b} & \cellcolor{bestcol}92.17\% (${\pm 0.98\%}$) & \cellcolor{secondbestcol}89.71\% (${\pm 1.03\%}$)  \\
\textbf{beit\_b\_16} & 90.81\% (${\pm 0.41\%}$) & 87.95\% (${\pm 0.69\%}$)  \\
\textbf{deit\_b\_16} & 91.78\% (${\pm 0.64\%}$) & 89.22\% (${\pm 0.90\%}$)  \\\hdashline
\textbf{vgg19} & 90.54\% (${\pm 0.37\%}$) & \cellcolor{worstcol}87.05\% (${\pm 1.03\%}$)  \\
\textbf{resnet50} & 91.78\% (${\pm 0.44\%}$) & 88.24\% (${\pm 0.59\%}$)  \\
\textbf{efficientnetv2\_m} & 90.87\% (${\pm 0.67\%}$) & 88.34\% (${\pm 1.37\%}$)  \\
\textbf{convnext\_b} & \cellcolor{secondbestcol}92.15\% (${\pm 0.40\%}$) & \cellcolor{bestcol}89.82\% (${\pm 1.18\%}$)  \\\hline
\end{tabular}
}
\caption{Type classification fine-tuned}
\label{results:tab:ft_type}
\end{table}

\begin{table}[]
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lll}
\hline
\textbf{Model} & \textbf{Accuracy} & \textbf{Balanced accuracy} \\ \hline
\textbf{vit\_b\_16} & 87.42\% (${\pm 0.45\%}$) & 73.46\% (${\pm 1.44\%}$)  \\
\textbf{swin\_b} & \cellcolor{bestcol}89.35\% (${\pm 0.68\%}$) & 77.16\% (${\pm 2.98\%}$)  \\
\textbf{beit\_b\_16} & \cellcolor{worstcol}85.74\% (${\pm 0.37\%}$) & \cellcolor{worstcol}72.12\% (${\pm 1.38\%}$)  \\
\textbf{deit\_b\_16} & 87.85\% (${\pm 1.12\%}$) & 74.42\% (${\pm 1.99\%}$)  \\\hdashline
\textbf{vgg19} & 85.74\% (${\pm 1.40\%}$) & 72.43\% (${\pm 3.03\%}$)  \\
\textbf{resnet50} & 88.69\% (${\pm 0.99\%}$) & \cellcolor{secondbestcol}77.97\% (${\pm 2.25\%}$)  \\
\textbf{efficientnetv2\_m} & 87.55\% (${\pm 1.15\%}$) & 75.31\% (${\pm 1.60\%}$)  \\
\textbf{convnext\_b} & \cellcolor{secondbestcol}88.79\% (${\pm 1.07\%}$) & \cellcolor{bestcol}78.40\% (${\pm 1.26\%}$)  \\\hline
\end{tabular}
}
\caption{Material classification fine-tuned}
\label{results:tab:ft_mat}
\end{table}

\begin{table}[]
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lll}
\hline
\textbf{Model} & \textbf{Accuracy} & \textbf{Balanced accuracy} \\ \hline
\textbf{vit\_b\_16} & 92.05\% (${\pm 0.44\%}$) & 89.77\% (${\pm 0.38\%}$)  \\
\textbf{swin\_b} & \cellcolor{bestcol}95.05\% (${\pm 0.47\%}$) & \cellcolor{bestcol}93.94\% (${\pm 0.81\%}$)  \\
\textbf{beit\_b\_16} & \cellcolor{worstcol}91.27\% (${\pm 1.13\%}$) & \cellcolor{worstcol}88.83\% (${\pm 1.63\%}$)  \\
\textbf{deit\_b\_16} & 93.37\% (${\pm 1.15\%}$) & 91.67\% (${\pm 1.55\%}$)  \\\hdashline
\textbf{vgg19} & 92.20\% (${\pm 0.49\%}$) & 90.18\% (${\pm 0.72\%}$)  \\
\textbf{resnet50} & \cellcolor{secondbestcol}94.72\% (${\pm 0.74\%}$) & \cellcolor{secondbestcol}93.41\% (${\pm 1.05\%}$)  \\
\textbf{efficientnetv2\_m} & 92.65\% (${\pm 0.54\%}$) & 90.84\% (${\pm 0.51\%}$)  \\
\textbf{convnext\_b} & 94.60\% (${\pm 0.54\%}$) & 93.13\% (${\pm 0.61\%}$)  \\\hline
\end{tabular}
}
\caption{Artist classification fine-tuned}
\label{results:tab:ft_artist}
\end{table}

\begin{figure*}
    \centering
    \begin{subfigure}{0.45\textwidth}
    \def\svgwidth{7.7cm}
    \input{img/ots_scale_acc.pdf_tex}
    \caption{Off the shelf}
    \label{results:img:ots_scale}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.45\textwidth}
    \def\svgwidth{7.7cm}
    \input{img/ft_scale_acc.pdf_tex}
    \caption{Fine tuning}
    \label{results:img:ft_scale}
    \end{subfigure}
    \caption{Testing accuracy as datasets become smaller.}
\end{figure*}

\begin{figure}
    \centering
    \def\svgwidth{7.7cm}
    \input{img/ots_vs_ft_type.pdf_tex}
    \caption{Time/accuracy trade-offs for ViTs and CNNs either off the shelf or fine-tuned.}
    \label{results:img:ots_vs_ft_type}
\end{figure}

### 2022-06-08 result_analysis/csv2latex.py

@@ -1,0 +1,54 @@
+# Reads a csv-file showing all accuracyes and prints a latex table from it
+
+import pandas as pd
+import sys
+
+csv_file = sys.argv[1]
+df = pd.read_csv(csv_file)
+
+# print(df.columns)
+# print("\n\n\n")
+
+argmax_acc = df["accuracy_mean"].argmax()
+argmax2_acc = df["accuracy_mean"].argsort()[len(df) - 2]
+argmin_acc = df["accuracy_mean"].argmin()
+
+b_argmax_acc = df["balanced_accuracy_mean"].argmax()
+b_argmax2_acc = df["balanced_accuracy_mean"].argsort()[len(df) - 2]
+b_argmin_acc = df["balanced_accuracy_mean"].argmin()
+
+
+print(r"""\begin{table}[]
+\resizebox{\columnwidth}{!}{%
+\begin{tabular}{lll}
+\hline
+\textbf{Model} & \textbf{Accuracy} & \textbf{Balanced accuracy} \\ \hline""")
+
+for idx, r in df.iterrows():
+    model = r['model'].replace('_', '\_')
+    
+    acc = f"{(100 * r['accuracy_mean']):.02f}\%"
+    acc_std = f"(${{\pm {(100 * r['accuracy_std']):.02f}\%}}$)"
+    if idx == argmax_acc:
+        acc = "\cellcolor{bestcol}" + acc
+    elif idx == argmax2_acc:
+        acc = "\cellcolor{secondbestcol}" + acc
+    elif idx == argmin_acc:
+        acc = "\cellcolor{worstcol}" + acc
+    
+    b_acc = f"{(100 * r['balanced_accuracy_mean']):.02f}\%"
+    b_acc_std = f"(${{\pm {(100 * r['balanced_accuracy_std']):.02f}\%}}$)"
+    if idx == b_argmax_acc:
+        b_acc = "\cellcolor{bestcol}" + b_acc
+    elif idx == b_argmax2_acc:
+        b_acc = "\cellcolor{secondbestcol}" + b_acc
+    elif idx == b_argmin_acc:
+        b_acc = "\cellcolor{worstcol}" + b_acc
+    
+    print(f"\\textbf{{{model}}} & {acc} {acc_std} & {b_acc} {b_acc_std}  \\\\" + ("\hdashline" if idx == 3 else "") + ("\hline" if idx == 7 else ""))
+
+print(r"""\end{tabular}
+}
[…]

### 2022-06-09 writing/tex/draft_introduction.tex
Still working on the methods. Sleepy and headache, so might take a break :-(

@@
 In practice, many TL techniques exist.
-The example given in section \ref{vits} can be described as \textit{fine-tuning} TL.
+The example given in section \ref{vits} can be described as \textit{fine-tuning} (FT) TL.
 Here, a pre-trained model is used as a starting point, and then trained on the target task -- possibly with different hyper-parameters, and often with a new final layer that fits the task.
-\citeauthor{yosinski2014transferable} (\citeyear{yosinski2014transferable}) studied transferability properties of CNNs pre-trained on ImageNet, and suggested that fine-tuning results in improved performance and generalizability compared to training from scratch.
+\citeauthor{yosinski2014transferable} (\citeyear{yosinski2014transferable}) studied transferability properties of CNNs pre-trained on ImageNet, and suggested that FT results in improved performance and generalizability compared to training from scratch.
 Moreover, while transferability decreased as source and target task became more dissimilar, utilizing TL still seemed to be preferred.
@@
 A different TL technique is called \textit{off-the-shelf} (OTS) TL.
-It is similar to fine-tuning, except that all but the final layer is kept frozen during training.
+It is similar to FT, except that all but the final layer is kept frozen during training.
 \citeauthor{sharif2014cnn} (\citeyear{sharif2014cnn}) found that features extracted by a CNN pre-trained on ImageNet were general enough for OTS TL to work.
 More specifically, these OTS CNNs showed better performance than the (fully trained non-CNN) state of the art at the time.
-Besides reducing computational cost of training, a benefit of OTS TL is that it can prevent overfitting in cases where fine-tuning would not \citep{yosinski2014transferable}.
+Besides reducing computational cost of training, a benefit of OTS TL is that it can prevent overfitting in cases where FT would not \citep{yosinski2014transferable}.
 
@@
 This domain is characterized by small datasets, so to overcome the inductive bias problem described earlier, TL is imperative.
-It was found that ViTs benefit more from fine-tuning TL than CNNs do, and that ViTs pre-trained on ImageNet perform on par CNNs.
+It was found that ViTs benefit more from FT TL than CNNs do, and that ViTs pre-trained on ImageNet perform on par CNNs.
 
@@
 \\\\
-The current study will examine to what extent TL techniques (both fine-tuning and OTS) allow ViTs to become the preferred architecture when compute resources are limited, and attempt to find a methodology that is recommended in these circumstances.
-It will, however, restrict itself to a single CV task, namely material classification of artworks.
+The current study will examine to what extent TL techniques (both FT and OTS) allow ViTs to become the preferred architecture when compute resources are limited, and attempts to find a methodology that is recommended in these circumstances.
+It will, however, restrict itself to art classification tasks.
 Several experiments will be performed, of which results will be analysed both quantitatively as well as qualitatively.
@@
 
-More formally, the following research question is proposed: \quote{\textit{To what extent (and in what manners) does utilizing TL allow ViTs to perform on par with CNNs for material classification of artworks, when compute resources are limited to a single GPU/TPU?}}
+More formally, the following research question is proposed: \begin{quote}\textit{To what extent (and in what manners) does utilizing TL allow ViTs to perform on par with CNNs on common art classification problems, when compute resources are limited to a single GPU/TPU?}\end{quote}

### 2022-06-09 writing/tex/draft_methods.tex
Still working on the methods. Sleepy and headache, so might take a break :-(

@@
 % Experiments build upon Sabatelli et al.
-Most of the experiments and analysis discussed below build upon the work done in \citeauthor{sabatelli2018deep} (\citeyear{sabatelli2018deep}).
+Much of the experiments and analysis discussed below, builds upon the work done in \citeauthor{sabatelli2018deep} (\citeyear{sabatelli2018deep}).
+This paper investigated TL properties of CNNs in the art classification domain.
+The main focus was on a comparison between CNN architectures pretrained on ImageNet1K when they were either used as OTS feature extractors, or fine-tuned to the new task.
+It was suggested that either approach was beneficial compared to training from scratch, but that FT resulted in substantially better performance than OTS TL.
 
 \subsection{Dataset}
-% Mention the challenges and show the table, of course. Also add max and min classes per category to the table maybe
-% Don't forget to tell that images get scaled and cropped to 224 * 224 size and imagenet normalised
-% Dataset described in more detail in Appendix A, including histograms and confusion matrices + balancing 'algorithm'
 
@@
 
+Similar to \citeauthor{sabatelli2018deep}, used models were pre-trained on ImageNet1K.
+Moreover, The Rijksmuseum Challenge dataset \citep{mensink14icmr} is used to extract the target tasks from.
+This dataset consists of a large collection of digitized artworks, together with xml-formatted metadata.
+The classification tasks distilled from it are: (1) \textit{Type classification}, containing labels as `painting', `sculpture', `drawing', etc.; (2) \textit{Material classification}, with labels as `paper', `porcelain', `silver', etc.; and finally, (3) `Artist classification', where the model has to predict who the creator is.
+
+The full dataset contains 112,039 images, and allows for multiple or zero labels per classification task.
+Different from \citeauthor{sabatelli2018deep}, only a fraction of the full set is used per task, as the research question concerns small datasets specifically.
+Samples are excluded from a task if they contain no labels for it.
+Samples containing more than one label are excluded as well, since analysis showed that in all cases this is true for only a small portion of the dataset.
+
+There are still many samples and classes left after doing the operations described above.
+Samples are also unevenly distributed among classes, with Gini-coefficients being as high as 0.98\footnote{The Gini-coefficient is a measure of balance, with 0 meaning all classes have the same number of samples, and 1 meaning all samples belong to a single class.
+A uniform distribution has coefficient $\frac{1}{3}$}.
+To counteract this, only the 30 most occurring classes are selected, and a cap of 1000 instances per class is enforced by taking a random sample when this limit is exceeded.
+Table \ref{methods:datasets} shows the result of these balancing operations.
+The situation prior is shown between brackets.
+
+For robustness, 5 subsets are randomly generated per classification task. These are then split up into 80\%, 10\%, 10\% training, validating and testing sets, respectively.
+
+Other than the tasks described above, table \ref{methods:datasets} shows an additional experiment called `Scaling'.
+The goal of this experiment is to see how the findings of this paper hold up as dataset sizes are shrunk further.
+It does this by taking the top 15 classes of type classification, and scaling it 4 times by a factor $\sqrt[4]{\frac{1}{10}} \approx 0.56$.
+The 5 datasets produced in this manner are then respectively 100\%, 56\%, 32\%, 18\% and 10\% the size of the dataset shown in table \ref{methods:datasets}.
+Steps are taken to ensure the distribution remains the same.
+Finally, 5 datasets are generated per scaling factor, leading to a total of 25 datasets.
+The same 80\%, 10\%, 10\% split is used for all.
+
+% Mention the challenges and show the table, of course. Also add max and min classes per category to the table maybe
+% Don't forget to tell that images get scaled and cropped to 224 * 224 size and imagenet normalised
+% Dataset described in more detail in Appendix A, including histograms and confusion matrices + balancing 'algorithm'
+
 \subsection{Models}
+In total, 8 different neural network architectures were selected for all experiments, of which 4 are CNN-based and 4 ViT-based.
+
 % General information (4 per type, base/tiny etc)
@@
 
+% Actually, mention resizing and imagenet scaling here. Its more of a hyper-param thing anyway :-p
+
 \subsection{Hardware and software}

### 2022-06-10 writing/tex/draft_introduction.tex
Sort of finished the first draft of the Methods section. Probably will still change some thing before I move on to results, however..

@@
 
-\subsection{Related work and the current study}
+\subsection{Related work and the current study} \label{related_work}
 As mentioned, the paper that first introduced ViTs already demonstrated they benefit from TL \citep{dosovitskiy2020image}, and more has been discovered about ViTs' TL properties since then.

### 2022-06-10 writing/tex/draft_methods.tex
Sort of finished the first draft of the Methods section. Probably will still change some thing before I move on to results, however..

@@
 
-Similar to \citeauthor{sabatelli2018deep}, used models were pre-trained on ImageNet1K.
-Moreover, The Rijksmuseum Challenge dataset \citep{mensink14icmr} is used to extract the target tasks from.
+Similar to \citeauthor{sabatelli2018deep}, used models are pre-trained on ImageNet1K.
+Moreover, the \textit{Rijksmuseum Challenge} dataset \citep{mensink14icmr} is used to extract the target tasks from.
 This dataset consists of a large collection of digitized artworks, together with xml-formatted metadata.
@@
 
-For robustness, 5 subsets are randomly generated per classification task. These are then split up into 80\%, 10\%, 10\% training, validating and testing sets, respectively.
+For robustness, five subsets are randomly generated per classification task. These are then split up into 80\%, 10\%, 10\% training, validating and testing sets, respectively.
 
@@
 \subsection{Models}
-In total, 8 different neural network architectures were selected for all experiments, of which 4 are CNN-based and 4 ViT-based.
+Eight different neural network architectures are used throughout this paper, of which four are CNN-based and four ViT-based.
+When multiple versions are available (i.e.
+`base', `small', `tiny'), `base'/`medium' ones are chosen.
+
+To start off with CNNs, ResNet50 \citep{he2016deep} and VGG19 \citep{simonyan2014very} are chosen, as these were the best performing models in \citeauthor{sabatelli2018deep} for FT an OTS TL, respectively.
+In addition, \citeauthor{matsoukas2021time} (\citeyear{matsoukas2021time}) used ResNet50 too, while \citeauthor{zhou2021convnets} (\citeyear{zhou2021convnets}) used ResNet101 and ResNet152.
+To also add some more recent CNNs, ConvNext \citep{liu2022convnet} -- a purely CNN-based model inspired by ViTs' recent success -- and EfficientNetV2 \citep{tan2021efficientnetv2} are included as well.
+
+For ViTs, versions with a 16x16 patch size are chosen.
+The first model is the original ViT \citep{dosovitskiy2020image}.
+Swin \citep{liu2021swin}, which was mentioned in section \ref{related_work}, is also selected, as it showed promising performance in \citeauthor{zhou2021convnets}.
+DeiT \citep{touvron2021training}, the distillation based architecture, was also mentioned in section \ref{related_work}.
+The small version of this architecture was used in \citeauthor{matsoukas2021time}, as it is similar in size to ResNet50, to which it was compared.
+Here, however, a base-sized version is chosen.
+Specifically, the version that does not include output for distillation learning.
+% TODO if you do end up using distilation with ConvNext as a teacher: don't forget to change this section!!!
+
+All ViT-based models have a bit over 86 million parameters.
+ConvNext is of similar size, with 88.6 million parameters.
+ResNet50 and EfficientNetV2 are smaller, with 25.6 and 13.6 million respectively, while VGG19 is the largest model, with 143.7 parameters.
+It should be mentioned that for both OTS TL as FT, preliminary experiments were performed with `small' and `tiny' models.
+The results of these suggested that similar conclusion could have been drawn if these were used instead.
 
@@
 
-\subsection{Hyperparameters}
+\subsection{Hyperparameters and data augmentation}
+All images are resized to a $224 \times 224$ resolution.
+This is achieved by first scaling them to the desired size along the shortest axis (retaining aspect ratio), and then taking a center crop along the longer axis.
+In addition, all images are normalized to the RGB mean and standard deviation of ImageNet1K.
 
-% Actually, mention resizing and imagenet scaling here. Its more of a hyper-param thing anyway :-p
+For all models, the linear classification layer is replaced as to fit the new task.
+Cross entropy loss uses the raw outputs of this layer (conform PyTorch documentation), together with one-hot encoded prediction targets.
+
+For OTS TL, the Adam optimizer is used with standard PyTorch parameters (\texttt{lr}=1e-3, $\beta_1=0.9$, $\beta_2=0.999$), and a batch size of 256.
+For FT, more is needed to achieve good convergence, especially for ViT-based architectures.
+Used hyperparameters, etc.
+are partially inspired by \citeauthor{matsoukas2021time} (\citeyear{matsoukas2021time}) and \citeauthor{zhou2021convnets} (\citeyear{zhou2021convnets}).
+The Adam learning rate now starts off at 1e-4, and is reduced by a factor 10 after 3 epochs without improvement (i.e.
+patience=2); the batch size is reduced to 32; label smoothing of 0.1 is added to the cross entropy loss, and a dropout layer with p=0.2 is inserted before the classification layer.
+Finally, input images are augmented with random horizontal flips and rotations in a $\pm 10 ^\circ$ range.
 
 \subsection{Hardware and software}
+All experiments are conducted on a single compute node containing one 32GB Nvidia V100 GPU.
+These nodes are constituents of the \textit{Peregrine high performance computing cluster}, belonging to the \textit{University of Groningen's Center for Information Technology}.
+The FT experiments take advantage of the V100's mixed precision acceleration.
+An exception is made for the type classification experiment, as this one is also used to compare OTS TL and FT in terms of time/accuracy trade-offs.
+Preliminary findings suggested that using mixed precision does not affect performance for the experiments where it is used.
+
+The PyTorch machine learning framework is used for all experiments, and many pre-trained models are taken from its Torchvision library.
+Exceptions are EfficientNetV2, Swin, DeiT and Beit, which are taken from the Timm library\footnote{\url{https://timm.fast.ai/}.}.
+
+Finally, all code and data is made available on GitHub\footnote{\url{https://github.com/IndoorAdventurer/ViTTransferLearningForArtClassification}.}.

### 2022-06-13 writing/tex/draft_methods.tex
Worked on the results section

@@
 The main focus was on a comparison between CNN architectures pretrained on ImageNet1K when they were either used as OTS feature extractors, or fine-tuned to the new task.
-It was suggested that either approach was beneficial compared to training from scratch, but that FT resulted in substantially better performance than OTS TL.
+It was suggested that FT resulted in substantially better performance than training from scratch or OTS TL.
 
@@
 
-\subsection{Models}
+\subsection{Models} \label{methods:models}
 Eight different neural network architectures are used throughout this paper, of which four are CNN-based and four ViT-based.
@@
 
-To start off with CNNs, ResNet50 \citep{he2016deep} and VGG19 \citep{simonyan2014very} are chosen, as these were the best performing models in \citeauthor{sabatelli2018deep} for FT an OTS TL, respectively.
+To start off with CNNs, ResNet50 \citep{he2016deep} and VGG19 \citep{simonyan2014very} are chosen, as these were the best performing models in \citeauthor{sabatelli2018deep} for respectively FT and OTS TL.
 In addition, \citeauthor{matsoukas2021time} (\citeyear{matsoukas2021time}) used ResNet50 too, while \citeauthor{zhou2021convnets} (\citeyear{zhou2021convnets}) used ResNet101 and ResNet152.
@@
 The small version of this architecture was used in \citeauthor{matsoukas2021time}, as it is similar in size to ResNet50, to which it was compared.
-Here, however, a base-sized version is chosen.
+Here, however, also a base-sized version is chosen.
 Specifically, the version that does not include output for distillation learning.
@@
 ConvNext is of similar size, with 88.6 million parameters.
-ResNet50 and EfficientNetV2 are smaller, with 25.6 and 13.6 million respectively, while VGG19 is the largest model, with 143.7 parameters.
-It should be mentioned that for both OTS TL as FT, preliminary experiments were performed with `small' and `tiny' models.
-The results of these suggested that similar conclusion could have been drawn if these were used instead.
+ResNet50 and EfficientNetV2 are smaller, with 25.6 and 13.6 million, respectively, while VGG19 is the largest model, with 143.7 million parameters.
+It should be mentioned that for both OTS TL and FT, preliminary experiments were performed with `small' and `tiny' models.
+The results of these suggested that similar conclusions could have been drawn if these were used instead.
 
@@
 Cross entropy loss uses the raw outputs of this layer (conform PyTorch documentation), together with one-hot encoded prediction targets.
+An early stop occurs after 10 epochs without improved loss on the validation set, and the model with the lowest loss is benchmarked on the testing set.
 
@@
 \subsection{Hardware and software}
-All experiments are conducted on a single compute node containing one 32GB Nvidia V100 GPU.
+All experiments are conducted on a single compute node containing one 32 GB Nvidia V100 GPU.
 These nodes are constituents of the \textit{Peregrine high performance computing cluster}, belonging to the \textit{University of Groningen's Center for Information Technology}.
@@
 An exception is made for the type classification experiment, as this one is also used to compare OTS TL and FT in terms of time/accuracy trade-offs.
-Preliminary findings suggested that using mixed precision does not affect performance for the experiments where it is used.
+Preliminary findings suggested that using mixed precision does not affect performance for the aforementioned experiments.
 

### 2022-06-13 writing/tex/draft_results.tex
Sort of finished the first draft of the Methods section. Probably will still change some thing before I move on to results, however..
Worked on the results section

@@
 \section{Results}
+
+This section presents results of the conducted experiments, mostly in the form of tables and line plots.
+Tables show average testing performance over the 5 trials per experiment, with standard deviation ($s$) between brackets.
+A green shaded cell marks best performance, while yellow and red depict second best and worst performance, respectively.
+
+Line plots show average validation accuracy (dashed line is CNN-based, continuous is ViT-based), with \textit{standard error of the mean} ($\pm s \div \sqrt{N}$, where $N=5$) depicted as a shaded region.
+Plots end when an early stop occurred for the first of 5 trials.
+
+% Tell that the appendix shows all other plots as well maybe. Only for the BP paper tho.
+
+\subsection{Off the shelf learning}
 
@@
     \caption{Validation accuracy using off the shelf models.}
+    \label{results:img:ots}
 \end{figure*}
@@
 
+Tables \ref{results:tab:ots_type} through \ref{results:tab:ots_artist} show performance of OTS trained models on the 3 classification tasks, while figure \ref{results:img:ots} shows corresponding validation accuracy per epoch.
+Performance on the testing set is comparable on all tasks in terms of ranking.
+Other observations are consistent as well, such as ConvNext taking the longest to converge.
+
+Swin is coming out on top in all cases, which is in line with results of \citeauthor{zhou2021convnets} (\citeyear{zhou2021convnets}).
+What slightly deviates from these result, is the observation that ViT performs better than a ResNet-based model on 2 out of 3 tasks.
+This comparison is, however, not very meaningful, as \citeauthor{zhou2021convnets} uses ResNet101 and 152 instead of ResNet50.
+
+It is interesting that VGG19 performs worse than ResNet50, as it was the best performing OTS model on all 3 tasks in \citeauthor{sabatelli2018deep} (\citeyear{sabatelli2018deep}).
+This discrepancy is likely attributable to the difference in dataset size, as VGG19 is by far the largest model (see section \ref{methods:models}), and likely too large for such small datasets.
+
+When comparing the tasks to one another, it can be observed that overall accuracies are lowest for material classification, and highest for artist classification.
+This seems counterintuitive, as one would think artist identification is much more challenging than determining if something is a painting, or if it is made out of paper instead of wood.
+An explanation can likely be found in table \ref{methods:datasets}, which shows that material and artist classification sets are the least and most balanced, respectively.
+In addition, tables \ref{results:tab:ots_type} through \ref{results:tab:ots_artist} show that balanced accuracies are up to 14\% lower than plain ones for material classification, while they are much closer for artist classification.
+This suggests much accuracy is lost by misclassifying instances of less occurring classes in favor of more occurring ones.
+Similar finding can be done for FT results in section \ref{results:ft}.
+% Maybe switch this to discussion tho ^
+
+%OTS TL:
+%    Comparisons with Sabatelli paper (ResNet50 vs VGG19)
+%    ViTs slightly better
+%    Lower balanced accuracy
+%    Higher balanced accuracy for type classification and even more for artist
+%    Artist higherst, which is surprising since it is the most difficult intuitively
+
+\subsection{Fine-tuning} \label{results:ft}
+
 \begin{figure*}
@@
     \caption{Validation accuracy using fine-tuned models.}
+    \label{results:img:ft}
 \end{figure*}
@@
 
+\subsection{Scaling dataset size}
+
 \begin{figure*}
@@
 
+\subsection{Off the shelf learning versus fine-tuning}
+
 \begin{figure}
@@
 \end{figure}
+
+\subsection{Qualitative analysis}
+TODO

### 2022-06-15 writing/tex/draft_introduction.tex
First draft of the results finshed

@@
 The example given in section \ref{vits} can be described as \textit{fine-tuning} (FT) TL.
-Here, a pre-trained model is used as a starting point, and then trained on the target task -- possibly with different hyper-parameters, and often with a new final layer that fits the task.
+Here, a pre-trained model is used as a starting point, and then trained on the target task -- possibly with different hyperparameters, and often with a new final layer that fits the task.
 \citeauthor{yosinski2014transferable} (\citeyear{yosinski2014transferable}) studied transferability properties of CNNs pre-trained on ImageNet, and suggested that FT results in improved performance and generalizability compared to training from scratch.

### 2022-06-15 writing/tex/draft_methods.tex
First draft of the results finshed

@@
 The main focus was on a comparison between CNN architectures pretrained on ImageNet1K when they were either used as OTS feature extractors, or fine-tuned to the new task.
-It was suggested that FT resulted in substantially better performance than training from scratch or OTS TL.
+It was suggested that FT resulted in substantially better performance than either training from scratch or OTS TL.
 
-\subsection{Dataset}
+\subsection{Dataset} \label{methods:dataset}
 
@@
 Swin \citep{liu2021swin}, which was mentioned in section \ref{related_work}, is also selected, as it showed promising performance in \citeauthor{zhou2021convnets}.
-DeiT \citep{touvron2021training}, the distillation based architecture, was also mentioned in section \ref{related_work}.
-The small version of this architecture was used in \citeauthor{matsoukas2021time}, as it is similar in size to ResNet50, to which it was compared.
-Here, however, also a base-sized version is chosen.
+Section \ref{related_work} mentioned DeiT \citep{touvron2021training} too, as the distillation based architecture.
+The small version of this architecture was used in \citeauthor{matsoukas2021time} (\citeyear{matsoukas2021time}), since it is similar in size to ResNet50, to which it was compared.
+Here, however, a base-sized version is chosen.
 Specifically, the version that does not include output for distillation learning.
-% TODO if you do end up using distilation with ConvNext as a teacher: don't forget to change this section!!!
+% TODO if you do end up using distilation with ConvNext as a teacher: don't forget to change this section!!! TODO 2: im not completely happy with the part about DeiT. Doesn't flow super nicely
 

### 2022-06-15 writing/tex/draft_results.tex
First draft of the results finshed

--- 2022-06-08 writing/tex/draft_results.tex
@@
 \section{Results}
+
+This section presents results of the conducted experiments, mostly in the form of tables and line plots.
+Tables show average testing performance over the 5 trials per experiment, with standard deviation ($s$) between brackets.
+A green shaded cell marks best performance, while yellow and red depict second best and worst performance, respectively.
+
+Line plots mostly show average validation accuracy (dashed line is CNN-based, continuous is ViT-based), with \textit{standard error of the mean} ($\pm s \div \sqrt{N}$, where $N=5$) depicted as a shaded region.
+Plots end when an early stop occurred for the first of 5 trials.
+
+% Tell that the appendix shows all other plots as well maybe. Only for the BP paper tho.
+
+\subsection{Off the shelf learning} \label{results:ots}
 
@@
 
+Tables \ref{results:tab:ots_type} through \ref{results:tab:ots_artist} show testing performance of OTS trained models on the 3 classification tasks, while figure \ref{results:img:ots} shows corresponding validation accuracies.
+Performance on the testing set is comparable among all tasks in terms of ranking.
+Other observations are consistent as well, such as ConvNext taking the longest to converge.
+
+Swin is coming out on top in all cases, which is in line with results of \citeauthor{zhou2021convnets} (\citeyear{zhou2021convnets}).
+What slightly deviates from these result, is the observation that ViT performs better than a ResNet-based model (on 2 out of 3 tasks).
+This comparison is, however, not meaningful, as \citeauthor{zhou2021convnets} uses ResNet101 and 152.
+Using these models instead, gives 86.21\% ($\pm$ 0.70\%) accuracy for ResNet101, and 86.19\% ($\pm$ 0.60\%) for ResNet152, which is indeed slightly higher than ViT's 86.06\% shown in table \ref{results:tab:ots_type}.
+
+It is interesting that VGG19 performs worse than ResNet50, since it was the best performing OTS model on all 3 tasks in \citeauthor{sabatelli2018deep} (\citeyear{sabatelli2018deep}).
+The current study uses smaller datasets and fewer classes.
+The difference might be attributable to this, as VGG19 is the largest model (see section \ref{methods:models}), and has the largest final layer -- with 4096 inputs.
+
+When comparing tasks to one another, it can be observed that overall accuracies are lowest for material classification, and highest for artist classification.
+In addition, the difference between balanced and plain accuracy is the largest for material classification (up to 14\% on the testing set), while it is the smallest for artist classification (at most 4.7\%).
+Gini-coefficients in table \ref{methods:datasets} show the same trend, with material classification being the least balanced, and artist classification the most.
+
+As a whole, ViTs are doing well for the OTS experiments.
+Especially Swin and DeiT are ranked high on all testing sets, while ConvNext is the most promising CNN-based architecture.
+Taking the mean accuracy of all ViT- and CNN-based models, reveals that ViTs as a group perform slightly better.
+On type classification, for example, their mean accuracy is 86.5\%, while it is 85.5\% for CNNs.
+
+% This seems counterintuitive, as one would think artist identification is much more challenging than determining if something is a painting, or if it is made out of paper instead of wood.
+An explanation can likely be found in table \ref{methods:datasets}, which shows that material and artist classification sets are the least and most balanced, respectively.
+In addition, tables \ref{results:tab:ots_type} through \ref{results:tab:ots_artist} show that balanced accuracies are up to 14\% lower than plain ones for material classification, while they are much closer for artist classification.
+This suggests much accuracy is lost by misclassifying instances of less occurring classes in favor of more occurring ones.
+Similar finding can be done for FT results in section \ref{results:ft}.
+% Maybe switch this to discussion tho ^
+
+%OTS TL:
+%    Comparisons with Sabatelli paper (ResNet50 vs VGG19)
+%    ViTs slightly better
+%    Lower balanced accuracy
+%    Higher balanced accuracy for type classification and even more for artist
+%    Artist higherst, which is surprising since it is the most difficult intuitively
+
+\subsection{Fine-tuning} \label{results:ft}
+
 \begin{figure*}
@@
 
+Once again, tables \ref{results:tab:ft_type} through \ref{results:tab:ft_artist} show testing performance, while figure \ref{results:img:ft} shows corresponding validation accuracies.
+
+Similar to OTS TL, findings are somewhat consistent across classification tasks, while overall performance is again lowest for material classification, and highest for artist classification.
+Swin is still dominant in terms of testing accuracy, but Conv\-Next is catching up.
+This is especially true for balanced accuracy, where it tops the ranking on 2 of the 3 tasks.
+More general, it looks as if the advantage ViTs seemed to have for OTS TL is gone.
+Mean accuracy of ViTs grouped together is, for example, 91.2\% on type classification, while it is 91.3\% for CNNs.
+
+Different from \citeauthor{matsoukas2021time} (\citeyear{matsoukas2021time}), ResNet50 performs slightly better than DeiT.
+This remains true when using the same `tiny' version of DeiT as that paper used.
+On type classification, for example, it gets 90.44\% ($\pm$ 0.93\%) accuracy, whereas ResNet50 shows 91.78\% in table \ref{results:tab:ft_type}.
+
+FT results also seem to be less favourable for ViTs compared to \citeauthor{zhou2021convnets} (\citeyear{zhou2021convnets}).
+ViT consistently falls behind ResNet50, and even shows the worst accuracy in table \ref{results:tab:ft_type} (90.11\%).
+It should be mentioned that ResNet101 and 152  -- which were used in \citeauthor{zhou2021convnets} --  show performance similar to ResNet50, with type classification accuracies of 91.96\% ($\pm$ 0.59\%) and 91.94\% ($\pm$ 0.25), respectively.
+
+The performance difference between best and worst models is much smaller compared to OTS TL.
+There it was between 7.2\% and 11.8\%, while here it is at most 3.8\%.
+In addition, the worst FT model is often better than the best OTS model.
+The only exception is material classification, where the worst FT model (BeiT) has accuracy 85.74\%, and the best OTS one (Swin) has 85.87\%.
+
+VGG19 has dropped further in the rankings, and now shows the worst balanced accuracy for type classification.
+ResNet50, on the other hand, appears to flourish with FT, often even beating ConvNext in the rankings.
+This is consistent with \citeauthor{sabatelli2018deep} (\citeyear{sabatelli2018deep}), where it was the best FT model overall.
+
+\subsection{Scaling dataset size}
+
 \begin{figure*}
@@
     \caption{Testing accuracy as datasets become smaller.}
+    \label{results:img:scale}
 \end{figure*}
+
+Figure \ref{results:img:scale} shows how testing accuracy decreases when smaller portions of the full dataset are taken (see section \ref{methods:dataset}).
+The x-axes show a logarithmic scale, with the rightmost value being roughly 10\% the size of the leftmost one.
+
+For both OTS TL in figure \ref{results:img:ots_scale}, as FT TL in figure \ref{results:img:ft_scale}, findings done in sections \ref{results:ots} and \ref{results:ft} seem to hold up as dataset sizes are shrunk.
+When going from largest to smallest dataset, the mean accuracy of all models taken together drops with 9.1\% for OTS TL.
+For FT this is 9.2\%, which is very similar.
+
+\subsection{Off the shelf learning versus fine-tuning}
 
@@
 \end{figure}
+
+Figure \ref{results:img:ots_vs_ft_type} shows type classification accuracy as reported in tables \ref{results:tab:ots_type} and \ref{results:tab:ft_type}.
+These are plotted against average epoch duration, as to get a sense of time/accuracy trade-offs.
+Colours make it easy to distinguish ViTs from CNNs, and OTS models from FT ones.
+
+OTS ViTs all take roughly the same time per epoch.
+CNNs are generally faster here, but accuracy also clearly falls behind for all but one (ConvNext).
+For FT, accuracies seem on-par, while CNNs still take much shorter to train.
+One CNN (ResNet50) even is faster than any OTS ViT, while also showing better accuracy.
+In short, FT CNNs have a better time/accuracy trade-off, while it is a FT ViT (Swin) that achieves the best accuracy overall.
+
+\subsection{Qualitative analysis}
+TODO

### 2022-06-15 result_analysis/qualitative/get_swin_attention_layers.py
Looked into attention heads of swin and deit

@@ -54,7 +54,7 @@ with open(args.target + "_prediction.txt", "w") as f:
 
 
 #---CHANGING-ATTENTION-WEIGHTS-TO-VIT-STYLE:---------------------------------------
-def meanAttentionPerLayerTest(att_mat, shift=0):  
+def meanAttentionPerLayer(att_mat, shift=0):  
     # Keeping the dimension in case I want to do something with it later:
     m = np.mean(att_mat, axis=1, keepdims=True)
 
@@ -88,11 +88,11 @@ def meanAttentionPerLayerTest(att_mat, shift=0):
     m = m.reshape(s[0], sqrt_frm, sqrt_frm, s[2], s[3])
 
     # Expanding it such that it is the same size as the final one
-    if repeats > 1:
-        m = m.repeat(repeats=repeats, axis=1)
-        m = m.repeat(repeats=repeats, axis=2)
-        m = m.repeat(repeats=repeats, axis=3)
-        m = m.repeat(repeats=repeats, axis=4)
+    # if repeats > 1:
+    #     m = m.repeat(repeats=repeats, axis=1)     REMOVED THIS BECAUSE
+    #     m = m.repeat(repeats=repeats, axis=2)     I AM NO LONGER DOING
+    #     m = m.repeat(repeats=repeats, axis=3)     ATTENTION ROLLOUT
+    #     m = m.repeat(repeats=repeats, axis=4)     FOR SWIN! GRADCAM INSTEAD
 
     # Change the shape to how it works for ViTs as well
     s = m.shape
@@ -105,7 +105,7 @@ def meanAttentionPerLayerTest(att_mat, shift=0):
 print("Processing layers", flush=True)
 with open(args.target + "_layers.npy", "wb") as f:
     for idx, layer in enumerate(att_weighs):
-        A = meanAttentionPerLayerTest(layer, shift= 3 * int(idx % 2))
+        A = meanAttentionPerLayer(layer, shift= 3 * int(idx % 2))
         np.save(f, A)
         print(".", end="", flush=True)
 

### 2022-06-21 writing/tex/draft_introduction.tex
writing and redoing some plots

@@
 
-More formally, the following research question is proposed: \begin{quote}\textit{To what extent (and in what manners) does utilizing TL allow ViTs to perform on par with CNNs on common art classification problems, when compute resources are limited to a single GPU/TPU?}\end{quote}
+More formally, the following research question is proposed: \begin{quote}\textit{To what extent (and in what manners) does utilizing TL allow ViTs to perform on par with CNNs on common art classification problems, when compute resources are limited to a single GPU?}\end{quote}

### 2022-06-21 writing/tex/draft_methods.tex
writing and redoing some plots

@@
 Specifically, the version that does not include output for distillation learning.
+% TODO TODO TODO TODO TODO TODO TODO TODO YOU FORGOT TO MENTION BEIT!!!
 % TODO if you do end up using distilation with ConvNext as a teacher: don't forget to change this section!!! TODO 2: im not completely happy with the part about DeiT. Doesn't flow super nicely
@@
 
-Finally, all code and data is made available on GitHub\footnote{\url{https://github.com/IndoorAdventurer/ViTTransferLearningForArtClassification}.}.
+Finally, all source code and data is made available on GitHub\footnote{\url{https://github.com/IndoorAdventurer/ViTTransferLearningForArtClassification}.}.

### 2022-06-21 writing/tex/draft_results.tex
Looked into attention heads of swin and deit
writing and redoing some plots

@@
 
-\begin{table}[]
+\begin{table}[tb]
 \resizebox{\columnwidth}{!}{%
@@
 In addition, the difference between balanced and plain accuracy is the largest for material classification (up to 14\% on the testing set), while it is the smallest for artist classification (at most 4.7\%).
-Gini-coefficients in table \ref{methods:datasets} show the same trend, with material classification being the least balanced, and artist classification the most.
+Note that Gini-coefficients in table \ref{methods:datasets} show the same trend, with material classification being the least balanced, and artist classification the most.
 
@@
 ViT consistently falls behind ResNet50, and even shows the worst accuracy in table \ref{results:tab:ft_type} (90.11\%).
-It should be mentioned that ResNet101 and 152  -- which were used in \citeauthor{zhou2021convnets} --  show performance similar to ResNet50, with type classification accuracies of 91.96\% ($\pm$ 0.59\%) and 91.94\% ($\pm$ 0.25), respectively.
+It should be mentioned that ResNet101 and 152  -- which were used alongside ViT in \citeauthor{zhou2021convnets} --  show performance similar to ResNet50, with type classification accuracies of 91.96\% ($\pm$ 0.59\%) and 91.94\% ($\pm$ 0.25), respectively.
 
@@
 One CNN (ResNet50) even is faster than any OTS ViT, while also showing better accuracy.
-In short, FT CNNs have a better time/accuracy trade-off, while it is a FT ViT (Swin) that achieves the best accuracy overall.
+In short, FT CNNs have better time/accuracy trade-offs, while it is a FT ViT (Swin) that achieves the best accuracy overall.
 
 \subsection{Qualitative analysis}
-TODO
+
+\begin{figure*}[tb]
+    \includegraphics[width=\textwidth]{img/layers.png}
+    \caption{Attention layers of successively deeper transformer blocks.}
+    \label{results:img:layers}
+\end{figure*}
+
+\begin{figure*}[tb]
+    \centering
+    \begin{subfigure}{0.49\textwidth}
+    \includegraphics[width=8cm]{img/img005_salience.png}
+    \caption{Dish}
+    \label{results:img:sal:dish}
+    \end{subfigure}
+    \begin{subfigure}{0.49\textwidth}
+    \includegraphics[width=8cm]{img/img011_salience.png}
+    \caption{Picture}
+    \label{results:img:sal:dish}
+    \end{subfigure}
+    \caption{Examples of saliency maps for different models.}
+\end{figure*}

### 2022-06-23 presentation/presentation.tex
working on the presentation

\documentclass{beamer}

% Style:
\usepackage{fontspec}
\setsansfont{Candara}
\usetheme{Rochester}
\definecolor{primarycolor}{RGB}{15,20,40}
\usecolortheme[named=primarycolor]{structure}
\setbeamercolor{normal text}{fg=primarycolor}

\title{\textsc{\bfseries Transfer Learning with\\Vision Transformers}}
\author{\textbf{Vincent} Tonkes}
\institute{\textbf{University} of Groningen}
\date{}

% Also makes it bold and small caps
\newcommand{\createtitle}[1]{\frametitle{\textsc{\bfseries#1}}}

\begin{document}

{
\setbeamercolor*{frametitle}{bg=white}
\frame{
\titlepage
{\tiny Supervised by:}\\\small\textbf{Matthia} Sabatelli\tiny, PhD
}
}

% Leaves them out on slide two and highlights them on slide 3
\newcommand{\toexplain}[1]{
    \textcolor<2>{white}{\textcolor<3>{red}{#1}}
}

\begin{frame}
\createtitle{Research Question}
\begin{center}
\textbf{\large ``\textit{To what extent (and in what manners) does utilizing \toexplain{Transfer Learning} allow \toexplain{Vision Transformers} to perform on par with \toexplain{Convolutional Neural Networks} on common \toexplain{art classification problems}, when compute resources are limited to a single \toexplain{GPU}?}''}
\end{center}
\end{frame}

\end{document}

### 2022-06-26 presentation/presentation.tex
First part of the presentation pretty much done. Now second part: my methods and results

@@
 \setbeamercolor{normal text}{fg=primarycolor}
+\setbeamertemplate{caption}{\raggedright\insertcaption\par}
 
@@
 
+%-------------------------------------------------------------------------------
+% Leaves them out on slide two and highlights them on slide 3
 {
@@
 
-% Leaves them out on slide two and highlights them on slide 3
+%-------------------------------------------------------------------------------
+
+%-------------------------------------------------------------------------------
 \newcommand{\toexplain}[1]{
@@
 }
-
 \begin{frame}
@@
 \end{frame}
+%-------------------------------------------------------------------------------
+
+%-------------------------------------------------------------------------------
+% IT HURTS, but I have to remove this because its too long already D-':
+%\begin{frame}
+%\createtitle{Transfer Learning}
+
+%\begin{columns}
+%\pause
+%\column{0.5\textwidth}
+%{\large Learning = \textbf{Machine} Learning:}
+%\pause
+%\begin{itemize}
+%\item Data that `summarizes' all previous learning experiences;
+%\pause
+%\item Way of updating that data with new experiences;
+%\pause
+%\item Way of making decisions based on that data.
+%\pause
+%\end{itemize}
+
+%\column{0.5\textwidth}
+%But what about Artificial Neural Networks?
+%\begin{figure}
+%\input{img/ann.pdf_tex}
+%\end{figure}
+
+%\end{columns}
+%\end{frame}
+%-------------------------------------------------------------------------------
+
+%-------------------------------------------------------------------------------
+\begin{frame}
+
+\pause
+\createtitle{Transfer Learning}
+\begin{block}{In this context...}
+Learning = \textbf{Machine} Learning
+\end{block}
+
+\pause
+\begin{columns}[t]
+\column{0.5\textwidth}
+\centering
+Task $T_A$\\
+\vspace{0.15cm}
+$f($
+\raisebox{-0.3cm}{\includegraphics[width=0.8cm]{img/cat.png}}
+$) =$ \texttt{cat}
+
+$f($
+\raisebox{-0.3cm}{\includegraphics[width=0.8cm]{img/dog.png}}
+$) =$ \texttt{dog}
+\column{0.5\textwidth}
+\pause
+\centering
+Model $M_A$\\
+\def\svgwidth{0.5\textwidth}
+\input{img/ann.pdf_tex}
+\end{columns}
+
+\begin{columns}
+\column{0.5\textwidth}
+\pause
+\centering
+Task $T_B$\\
+\vspace{0.15cm}
+$f($
+\raisebox{-0.1cm}{\includegraphics[width=0.8cm]{img/garfield.jpeg}}
+$) =$ \texttt{cat}
+
+$f($
+\raisebox{-0.3cm}{\includegraphics[width=0.8cm]{img/odie.png}}
+$) =$ \texttt{dog}
+\column{0.5\textwidth}
+\pause
+\centering
+Can we somehow use $M_A$ to improve performance on $T_B$?
+\end{columns}
+
+\end{frame}
+%-------------------------------------------------------------------------------
+
+%-------------------------------------------------------------------------------
+\begin{frame}
+\createtitle{Transfer Learning}
+
+\begin{definition}
+\textbf{Transfer Learning:} {\footnotesize taking advantage of a model $M_A$ that was trained on task $T_A$ to improve performance on task $T_B$.}
+\end{definition}
+
+\pause
+\begin{definition}
+\textbf{Fine-tuning:} {\footnotesize a type of Transfer Learning where we take $M_A$ as a starting point for training on $T_B$, instead of a newly initialized model.}
+\end{definition}
+
+\pause
+\begin{definition}
+\textbf{Off the shelf learning:} {\footnotesize same as fine-tuning, except that all but the final layer is kept frozen during training.}
+\end{definition}
+
+\end{frame}
+%-------------------------------------------------------------------------------
+
+%-------------------------------------------------------------------------------
+\begin{frame}
+\createtitle{Convolutional Neural Networks\\and Vision Transformers}
+
+\pause
+Convolutional Neural Networks (CNNs)
+\begin{itemize}
+\item Have been around for a long time;
+\item Played a dominant role in Computer Vision the last decade.
+\end{itemize}
+
+\pause
+Vision Transformers (VTs)
+\begin{itemize}
+\item Fairly new (2020);
+\item Have been taking over in Computer Vision;
+\item Based on the Transformer architecture.
+\end{itemize}
+
+\end{frame}
+%-------------------------------------------------------------------------------
+
+%-------------------------------------------------------------------------------
+\begin{frame}
+\createtitle{Architecture}
+
+\begin{columns}
+\column{0.6\textwidth}
+\begin{figure}
+\includegraphics[width=\textwidth]{img/vit_vs_cnn.png}
+\caption{\tiny Top right: CNN, Bottom right: Transformer,
+
+Top + Bottom left: hypothetical fully connected variants.}
+\end{figure}
+
+\column{0.4\textwidth}
+\begin{itemize}
+\pause
+\item CNNs: designed to work with images;
+\pause
+\item Transformer (\alert{Not VTs!}): designed for natural language processing.
+\end{itemize}
+
+\end{columns}
+
+\end{frame}
+%-------------------------------------------------------------------------------
+
+%-------------------------------------------------------------------------------
+\begin{frame}
+\createtitle{From Transformer to Vision Transformer}
+
+Split image up into $16 \times 16$ patches, and treat them as words in a sentence.
+\vspace{0.5cm}
+
+\includegraphics[width=\textwidth]{img/vit_illustration.png}
+
+\pause
+\vspace{0.5cm}
+Transfer Learning to the rescue!
+
+\end{frame}
+%-------------------------------------------------------------------------------
+
+%-------------------------------------------------------------------------------
+\newcommand{\stilltoexplain}[1]{\textcolor{red}{#1}}
+\begin{frame}
+\createtitle{Research Question}
+\begin{center}
+\textbf{\large ``\textit{To what extent (and in what manners) does utilizing Transfer Learning allow Vision Transformers to perform on par with Convolutional Neural Networks \pause on common \stilltoexplain{art classification problems}, when compute resources are limited to a single \stilltoexplain{GPU}?}''}
+\end{center}
+\end{frame}
+%-------------------------------------------------------------------------------
+
+%-------------------------------------------------------------------------------
+\begin{frame}
+\createtitle{Art Classification Problems}
+\includegraphics[width=\textwidth]{img/examples.png}
+
+\vspace{0.5cm}
+\centering
+Taken from ``\textit{The Rijksmuseum Challenge}'' dataset
+\end{frame}
+%-------------------------------------------------------------------------------
+
+%-------------------------------------------------------------------------------
+\begin{frame}
+\createtitle{GPU}
+\begin{columns}
+\column{0.8\textwidth}
+Graphics processing unit
+\begin{itemize}
+\item Does the actual computations.
+\end{itemize}
+\column{0.2\textwidth}
+\includegraphics[width=\textwidth]{img/gpu.png}
+\end{columns}
+\vspace{0.5cm}
+\pause
+\rule{\textwidth}{0.7pt}
+
+\vspace{0.5cm}
+
+Why this constraint? ...
+\pause
+
+\vspace{0.5cm}
+
+... Well, because I am specifically interested in:
+\begin{itemize}
+\item Problems not too demanding for consumer-grade hardware;
+\item Datasets small enough to be labelled by one person.
+\end{itemize}
+\pause
+In other words: focusing on the realm of tinkerers!
+\end{frame}
+%-------------------------------------------------------------------------------
 

### 2022-06-26 presentation/presentation.tex
Slideshow done. First draft at least

--- 2022-06-23 presentation/presentation.tex
@@
 \setbeamercolor{normal text}{fg=primarycolor}
+\setbeamertemplate{caption}{\raggedright\insertcaption\par}
+\setbeamertemplate{bibliography item}[text]
+
+% For tables:
+\usepackage{colortbl}
+\definecolor{bestcol}{RGB}{99, 255, 105}
+\definecolor{secondbestcol}{RGB}{250, 255, 99}
+\definecolor{worstcol}{RGB}{255, 141, 99}
+\usepackage{arydshln}
 
@@
 
+% Make \cite commands tiny
+\newcommand{\tinycite}[1]{{\tiny \cite{#1}}}
+
 \begin{document}
 
+%-------------------------------------------------------------------------------
+% Leaves them out on slide two and highlights them on slide 3
 {
@@
 
-% Leaves them out on slide two and highlights them on slide 3
-\newcommand{\toexplain}[1]{
-    \textcolor<2>{white}{\textcolor<3>{red}{#1}}
+%-------------------------------------------------------------------------------
+
+\input{tex/question.tex}
+\input{tex/answer.tex}
+
+%-------------------------------------------------------------------------------
+{
+\setbeamercolor*{frametitle}{bg=white}
+\begin{frame}
+\centering
+\textsc{\Huge \textbf{That's} all!}
+
+\pause
+{\small except for the questions :-)}
+\end{frame}
 }
+%-------------------------------------------------------------------------------
 
-\begin{frame}
-\createtitle{Research Question}
-\begin{center}
-\textbf{\large ``\textit{To what extent (and in what manners) does utilizing \toexplain{Transfer Learning} allow \toexplain{Vision Transformers} to perform on par with \toexplain{Convolutional Neural Networks} on common \toexplain{art classification problems}, when compute resources are limited to a single \toexplain{GPU}?}''}
-\end{center}
+%-------------------------------------------------------------------------------
+\begin{frame}[t, allowframebreaks]
+\createtitle{References}
+\bibliographystyle{acm}
+{\tiny
+\bibliography{literature}}
 \end{frame}
+%-------------------------------------------------------------------------------
 

### 2022-06-26 presentation/tex/answer.tex
Slideshow done. First draft at least

%-------------------------------------------------------------------------------
\begin{frame}
\createtitle{Methods}
\pause
\begin{itemize}
\item 4 VTs (\texttt{ViT} \tinycite{dosovitskiy2020image}, \texttt{Swin} \tinycite{liu2021swin}, \texttt{DeiT} \tinycite{touvron2021training}, and \texttt{BeiT} \tinycite{bao2022beit}) versus 4 CNNs (\texttt{ResNet} \tinycite{he2016deep}, \texttt{VGG} \tinycite{simonyan2014very}, \texttt{ConvNext} \tinycite{liu2022convnet}, and \texttt{EfficientNetV2} \tinycite{tan2021efficientnetv2});
\begin{itemize}
\item All pre-trained on ImageNet1K;
\end{itemize}
\pause
\item Off-the-shelf and fine-tuning experiments on ...
\pause
\item ... those 3 tasks from the ``\textit{Rijksmuseum Challenge}'';
\begin{itemize}
\pause
\item Problems:
\begin{itemize}
\item Big (112K images);
\item Many classes;
\item Unbalanced;
\end{itemize}
\pause
\item Solutions:
\begin{itemize}
\item Taking top 30 classes;
\item Max of 1K images per class;
\end{itemize}
\pause
\item Bonus:
\begin{itemize}
\item Dataset sizes now between 6.5K and 9K;
\item Randomly generating 5 subsets per task for extra robustness.
\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}
%-------------------------------------------------------------------------------

%-------------------------------------------------------------------------------
\begin{frame}
\createtitle{Off the shelf results}
\centering
{\tiny \textbf{Validation accuracy during training}}
\def\svgwidth{\textwidth}
\input{img/ots_mean_accuracy.pdf_tex}

\vspace{0.3cm}

\resizebox{\columnwidth}{!}{%
\begin{tabular}{lllllll}
\hline
\textbf{Model} & \textbf{Type} &  & \textbf{Material} & & \textbf{Artist} & \\
& \textbf{Accuracy} & \textbf{Bal. accuracy} & \textbf{Accuracy} & \textbf{Bal. accuracy} & \textbf{Accuracy} & \textbf{Bal. accuracy} \\ \hline
\textbf{vit\_b\_16} & 86.06\% $_{\pm 1.06\%}$ & 84.13\% $_{\pm 1.57\%}$ & 81.78\% $_{\pm 0.48\%}$ & 67.38\% $_{\pm 1.37\%}$ & 84.89\% $_{\pm 0.46\%}$ & 81.42\% $_{\pm 0.42\%}$  \\
\textbf{swin\_b} & \cellcolor{bestcol}89.43\% $_{\pm 0.93\%}$ & \cellcolor{bestcol}87.47\% $_{\pm 1.02\%}$ & \cellcolor{bestcol}85.87\% $_{\pm 0.35\%}$ & \cellcolor{bestcol}71.19\% $_{\pm 1.60\%}$ & \cellcolor{bestcol}90.40\% $_{\pm 0.65\%}$ & \cellcolor{bestcol}88.64\% $_{\pm 0.78\%}$  \\
\textbf{beit\_b\_16} & \cellcolor{worstcol}82.26\% $_{\pm 0.72\%}$ & \cellcolor{worstcol}77.75\% $_{\pm 0.27\%}$ & 76.87\% $_{\pm 0.96\%}$ & 60.16\% $_{\pm 1.56\%}$ & 79.70\% $_{\pm 0.69\%}$ & 75.35\% $_{\pm 1.09\%}$  \\
\textbf{deit\_b\_16} & 88.18\% $_{\pm 0.66\%}$ & 85.36\% $_{\pm 0.54\%}$ & 82.80\% $_{\pm 1.12\%}$ & 66.46\% $_{\pm 1.03\%}$ & 88.13\% $_{\pm 0.76\%}$ & 85.62\% $_{\pm 0.87\%}$  \\\hdashline
\textbf{vgg19} & 83.93\% $_{\pm 0.72\%}$ & 83.35\% $_{\pm 0.81\%}$ & 76.87\% $_{\pm 0.44\%}$ & 61.39\% $_{\pm 1.47\%}$ & 82.01\% $_{\pm 0.66\%}$ & 78.10\% $_{\pm 0.77\%}$  \\
\textbf{resnet50} & 85.51\% $_{\pm 0.64\%}$ & 82.33\% $_{\pm 1.85\%}$ & 80.99\% $_{\pm 0.82\%}$ & 65.51\% $_{\pm 0.93\%}$ & 87.71\% $_{\pm 1.06\%}$ & 85.12\% $_{\pm 1.34\%}$  \\
\textbf{eff. netv2\_m} & 83.41\% $_{\pm 0.76\%}$ & 82.05\% $_{\pm 1.25\%}$  & \cellcolor{worstcol}75.96\% $_{\pm 1.24\%}$ & \cellcolor{worstcol}59.15\% $_{\pm 1.24\%}$ & \cellcolor{worstcol}78.62\% $_{\pm 1.07\%}$ & \cellcolor{worstcol}73.92\% $_{\pm 0.96\%}$  \\
\textbf{convnext\_b} & \cellcolor{secondbestcol}89.19\% $_{\pm 0.64\%}$ & \cellcolor{secondbestcol}86.95\% $_{\pm 1.38\%}$ & \cellcolor{secondbestcol}84.14\% $_{\pm 0.92\%}$ & \cellcolor{secondbestcol}69.10\% $_{\pm 1.05\%}$ & \cellcolor{secondbestcol}90.13\% $_{\pm 0.94\%}$ & \cellcolor{secondbestcol}87.84\% $_{\pm 1.07\%}$  \\\hline
\end{tabular}
}
\centering
{\tiny \textbf{Testing performance}}

\end{frame}
%-------------------------------------------------------------------------------

%-------------------------------------------------------------------------------
\begin{frame}
\createtitle{Fine-tuning results}
\centering
{\tiny \textbf{Validation accuracy during training}}
\def\svgwidth{\textwidth}
\input{img/ft_mean_accuracy.pdf_tex}

\vspace{0.3cm}

\resizebox{\columnwidth}{!}{%
\begin{tabular}{lllllll}
\hline
\textbf{Model} & \textbf{Type} &  & \textbf{Material} & & \textbf{Artist} & \\
& \textbf{Accuracy} & \textbf{Bal. accuracy} & \textbf{Accuracy} & \textbf{Bal. accuracy} & \textbf{Accuracy} & \textbf{Bal. accuracy} \\ \hline
\textbf{vit\_b\_16} & \cellcolor{worstcol}90.11\% $_{\pm 0.35\%}$ & 87.40\% $_{\pm 0.29\%}$ & 87.42\% $_{\pm 0.45\%}$ & 73.46\% $_{\pm 1.44\%}$ & 92.05\% $_{\pm 0.44\%}$ & 89.77\% $_{\pm 0.38\%}$  \\
\textbf{swin\_b} & \cellcolor{bestcol}92.17\% $_{\pm 0.98\%}$ & \cellcolor{secondbestcol}89.71\% $_{\pm 1.03\%}$ & \cellcolor{bestcol}89.35\% $_{\pm 0.68\%}$ & 77.16\% $_{\pm 2.98\%}$ & \cellcolor{bestcol}95.05\% $_{\pm 0.47\%}$ & \cellcolor{bestcol}93.94\% $_{\pm 0.81\%}$  \\
\textbf{beit\_b\_16} & 90.81\% $_{\pm 0.41\%}$ & 87.95\% $_{\pm 0.69\%}$ & \cellcolor{worstcol}85.74\% $_{\pm 0.37\%}$ & \cellcolor{worstcol}72.12\% $_{\pm 1.38\%}$ & \cellcolor{worstcol}91.27\% $_{\pm 1.13\%}$ & \cellcolor{worstcol}88.83\% $_{\pm 1.63\%}$  \\
\textbf{deit\_b\_16} & 91.78\% $_{\pm 0.64\%}$ & 89.22\% $_{\pm 0.90\%}$ & 87.85\% $_{\pm 1.12\%}$ & 74.42\% $_{\pm 1.99\%}$ & 93.37\% $_{\pm 1.15\%}$ & 91.67\% $_{\pm 1.55\%}$  \\\hdashline
\textbf{vgg19} & 90.54\% $_{\pm 0.37\%}$ & \cellcolor{worstcol}87.05\% $_{\pm 1.03\%}$ & 85.74\% $_{\pm 1.40\%}$ & 72.43\% $_{\pm 3.03\%}$ & 92.20\% $_{\pm 0.49\%}$ & 90.18\% $_{\pm 0.72\%}$  \\
\textbf{resnet50} & 91.78\% $_{\pm 0.44\%}$ & 88.24\% $_{\pm 0.59\%}$ & 88.69\% $_{\pm 0.99\%}$ & \cellcolor{secondbestcol}77.97\% $_{\pm 2.25\%}$ & \cellcolor{secondbestcol}94.72\% $_{\pm 0.74\%}$ & \cellcolor{secondbestcol}93.41\% $_{\pm 1.05\%}$  \\
\textbf{eff. netv2\_m} & 90.87\% $_{\pm 0.67\%}$ & 88.34\% $_{\pm 1.37\%}$ & 87.55\% $_{\pm 1.15\%}$ & 75.31\% $_{\pm 1.60\%}$ & 92.65\% $_{\pm 0.54\%}$ & 90.84\% $_{\pm 0.51\%}$  \\
\textbf{convnext\_b} & \cellcolor{secondbestcol}92.15\% $_{\pm 0.40\%}$ & \cellcolor{bestcol}89.82\% $_{\pm 1.18\%}$ & \cellcolor{secondbestcol}88.79\% $_{\pm 1.07\%}$ & \cellcolor{bestcol}78.40\% $_{\pm 1.26\%}$ & 94.60\% $_{\pm 0.54\%}$ & 93.13\% $_{\pm 0.61\%}$  \\\hline
\end{tabular}
}
\centering
{\tiny \textbf{Testing performance}}

\end{frame}
%-------------------------------------------------------------------------------

%-------------------------------------------------------------------------------
\begin{frame}
\createtitle{Scaling}
\centering
{\tiny \textbf{Testing accuracy as dataset gets smaller}}
\def\svgwidth{\textwidth}
\input{img/scaling_accuracy.pdf_tex}
\end{frame}
%-------------------------------------------------------------------------------

%-------------------------------------------------------------------------------
\begin{frame}
\createtitle{Scaling}
\centering
{\tiny \textbf{Time/accuracy trade-offs}}\\
\vspace{-1cm}
\def\svgwidth{7cm}
\input{img/ots_vs_ft_type.pdf_tex}
\end{frame}
%-------------------------------------------------------------------------------

%-------------------------------------------------------------------------------
\begin{frame}
\createtitle{Diving deeper ...}
\centering
{\tiny \textbf{Activation for deeper layers of VTs}}\\
\includegraphics[width=\textwidth]{img/layers.png}

\centering
{\tiny \textbf{Saliency maps}}\\

\vspace{-0.3cm}
\begin{columns}
\column{0.5\textwidth}
\includegraphics[width=\textwidth]{img/img011_salience.png}
\column{0.5\textwidth}
\includegraphics[width=\textwidth]{img/img005_salience.png}
\end{columns}
\end{frame}
%-------------------------------------------------------------------------------

### 2022-06-26 presentation/tex/question.tex
Slideshow done. First draft at least

%-------------------------------------------------------------------------------
\newcommand{\toexplain}[1]{
    \textcolor<2>{white}{\textcolor<3>{red}{#1}}
}
\begin{frame}
\createtitle{Research Question}
\begin{center}
\textbf{\large ``\textit{To what extent (and in what manners) does utilizing \toexplain{Transfer Learning} allow \toexplain{Vision Transformers} to perform on par with \toexplain{Convolutional Neural Networks} on common \toexplain{art classification problems}, when compute resources are limited to a single \toexplain{GPU}?}''}
\end{center}
\end{frame}
%-------------------------------------------------------------------------------

%-------------------------------------------------------------------------------
% IT HURTS, but I have to remove this because its too long already D-':
%\begin{frame}
%\createtitle{Transfer Learning}

%\begin{columns}
%\pause
%\column{0.5\textwidth}
%{\large Learning = \textbf{Machine} Learning:}
%\pause
%\begin{itemize}
%\item Data that `summarizes' all previous learning experiences;
%\pause
%\item Way of updating that data with new experiences;
%\pause
%\item Way of making decisions based on that data.
%\pause
%\end{itemize}

%\column{0.5\textwidth}
%But what about Artificial Neural Networks?
%\begin{figure}
%\input{img/ann.pdf_tex}
%\end{figure}

%\end{columns}
%\end{frame}
%-------------------------------------------------------------------------------

%-------------------------------------------------------------------------------
\begin{frame}

\pause
\createtitle{Transfer Learning}
\begin{block}{In this context...}
Learning = \textbf{Machine} Learning
\end{block}

\pause
\begin{columns}[t]
\column{0.5\textwidth}
\centering
Task $T_A$\\
\vspace{0.15cm}
$f($
\raisebox{-0.3cm}{\includegraphics[width=0.8cm]{img/cat.png}}
$) =$ \texttt{cat}

$f($
\raisebox{-0.3cm}{\includegraphics[width=0.8cm]{img/dog.png}}
$) =$ \texttt{dog}
\column{0.5\textwidth}
\pause
\centering
Model $M_A$\\
\def\svgwidth{0.5\textwidth}
\input{img/ann.pdf_tex}
\end{columns}

\begin{columns}
\column{0.5\textwidth}
\pause
\centering
Task $T_B$\\
\vspace{0.15cm}
$f($
\raisebox{-0.1cm}{\includegraphics[width=0.8cm]{img/garfield.jpeg}}
$) =$ \texttt{cat}

$f($
\raisebox{-0.3cm}{\includegraphics[width=0.8cm]{img/odie.png}}
$) =$ \texttt{dog}
\column{0.5\textwidth}
\pause
\centering
Can we somehow use $M_A$ to improve performance on $T_B$?
\end{columns}

\end{frame}
%-------------------------------------------------------------------------------

%-------------------------------------------------------------------------------
\begin{frame}
\createtitle{Transfer Learning}

\begin{definition}
\textbf{Transfer Learning:} {\footnotesize taking advantage of a model $M_A$ that was trained on task $T_A$ to improve performance on task $T_B$.}
\end{definition}

\pause
\begin{definition}
\textbf{Fine-tuning:} {\footnotesize a type of Transfer Learning where we take $M_A$ as a starting point for training on $T_B$, instead of a newly initialized model.}
\end{definition}

\pause
\begin{definition}
\textbf{Off the shelf learning:} {\footnotesize same as fine-tuning, except that all but the final layer is kept frozen during training.}
\end{definition}

\end{frame}
%-------------------------------------------------------------------------------

%-------------------------------------------------------------------------------
\begin{frame}
\createtitle{Convolutional Neural Networks\\and Vision Transformers}

\pause
Convolutional Neural Networks (CNNs) \tinycite{lecun1998gradient}
\begin{itemize}
\item Have been around for a long time;
\item Played a dominant role in Computer Vision the last decade.
\end{itemize}

\pause
Vision Transformers (VTs) \tinycite{dosovitskiy2020image}
\begin{itemize}
\item Fairly new (2020);
\item Have been taking over in Computer Vision;
\item Based on the Transformer architecture \tinycite{vaswani2017attention}.
\end{itemize}

\end{frame}
%-------------------------------------------------------------------------------

%-------------------------------------------------------------------------------
\begin{frame}
\createtitle{Architecture}

\begin{columns}
\column{0.6\textwidth}
\begin{figure}
\includegraphics[width=\textwidth]{img/vit_vs_cnn.png}
\caption{\tiny Top right: CNN, Bottom right: Transformer,

Top + Bottom left: hypothetical fully connected variants.}
\end{figure}

\column{0.4\textwidth}
\begin{itemize}
\pause
\item CNNs: designed to work with images;
\pause
\item Transformer (\alert{Not VTs!}): designed for natural language processing.
\end{itemize}

\end{columns}

\end{frame}
%-------------------------------------------------------------------------------

%-------------------------------------------------------------------------------
\begin{frame}
\createtitle{From Transformer to Vision Transformer}

Split image up into $16 \times 16$ patches, and treat them as words in a sentence.
\vspace{0.5cm}

\includegraphics[width=\textwidth]{img/vit_illustration.png}

\pause
\vspace{0.5cm}
Transfer Learning to the rescue!

\end{frame}
%-------------------------------------------------------------------------------

%-------------------------------------------------------------------------------
\newcommand{\stilltoexplain}[1]{\textcolor{red}{#1}}
\begin{frame}
\createtitle{Research Question}
\begin{center}
\textbf{\large ``\textit{To what extent (and in what manners) does utilizing Transfer Learning allow Vision Transformers to perform on par with Convolutional Neural Networks \pause on common \stilltoexplain{art classification problems}, when compute resources are limited to a single \stilltoexplain{GPU}?}''}
\end{center}
\end{frame}
%-------------------------------------------------------------------------------

%-------------------------------------------------------------------------------
\begin{frame}
\createtitle{Art Classification Problems}
\includegraphics[width=\textwidth]{img/examples.png}

\vspace{0.5cm}
\centering
Taken from ``\textit{The Rijksmuseum Challenge}'' dataset \tinycite{mensink14icmr}.
\end{frame}
%-------------------------------------------------------------------------------

%-------------------------------------------------------------------------------
\begin{frame}
\createtitle{GPU}
\begin{columns}
\column{0.8\textwidth}
Graphics processing unit
\begin{itemize}
\item Does the actual computations.
\end{itemize}
\column{0.2\textwidth}
\includegraphics[width=\textwidth]{img/gpu.png}
\end{columns}
\vspace{0.5cm}
\pause
\rule{\textwidth}{0.7pt}

\vspace{0.5cm}

Why this constraint? ...
\pause

\vspace{0.5cm}

... Well, because I am specifically interested in:
\begin{itemize}
\item Problems not too demanding for consumer-grade hardware;
\item Datasets small enough to be labelled by one person.
\end{itemize}
\pause
In other words: focusing on the realm of tinkerers!
\end{frame}
%-------------------------------------------------------------------------------

### 2022-06-30 writing/tex/draft_introduction.tex
First part of the presentation pretty much done. Now second part: my methods and results
Continued working on writing y paper

@@
 Since the introduction of AlexNet, roughly a decade ago, \textit{Convolutional Neural Networks} (CNNs) have played an important role in \textit{Computer Vision} (CV) \citep{krizhevsky2012imagened}.
-Recently, however, a new type of architecture, called \textit{Vision Transformer} (ViT), gained state of the art performance on common learning benchmarks, including the ImageNet dataset\footnote{At the time of writing TODO is the best performing model, and the top TODO are all Transformer-based: (website)}.
+Recently, however, a new type of architecture, called \textit{Vision Transformer} (VT), gained state of the art performance on common learning benchmarks, including the ImageNet dataset\footnote{At the time of writing TODO is the best performing model, and the top TODO are all Transformer-based: (website)}.
 
-Exciting as this may be, a plain ViT is not likely to become useful in circumstances with limited training data and/or only consumer-grade hardware available for learning.
-This is due to a limitation inherent to the ViT architecture, which will be elaborated on in section \ref{vits}.
-Consequently, the current paper investigates whether utilizing \textit{Transfer Learning} (TL) capabilities of ViTs can help overcome these limitations, and if this allows ViTs to become the preferred architecture in the aforementioned circumstances as well.
+Exciting as this may be, a plain VT is not likely to become useful in circumstances with limited training data and/or only consumer-grade hardware available for learning.
+This is due to a limitation inherent to the VT architecture, which will be elaborated on in section \ref{vits}.
+Consequently, the current paper investigates whether utilizing \textit{Transfer Learning} (TL) capabilities of VTs can help overcome these limitations, and if this allows VTs to become the preferred architecture in the aforementioned circumstances as well.
 
 Before going into more detail, however, some background information is needed.
-The remainder of this section will first briefly describe CNNs and ViTs, and compare them to one another.
-What follows, is an explanation of TL, and why it might help ViTs in circumstances with limited data/compute available.
+The remainder of this section will first briefly describe CNNs and VTs, and compare them to one another.
+What follows, is an explanation of TL, and why it might help VTs in circumstances with limited data/compute available.
 Finally, related work is discussed, and a research question is proposed.
@@
 %    \includegraphics[width=7.7cm]{img/vit_vs_cnn.png}
-%    \caption{Intuition behind CNNs and ViTs.
+%    \caption{Intuition behind CNNs and VTs.
 Note that many details are left out for the purpose of illustration.
@@
 Connections are only allowed such as the accentuated 4, which reduces the total number of connections greatly.
-Moreover, the non-accentuated connections don't add to the number of trainable weights, as the same 4 are shared between all neuron-receptive-field-pairs.\\The bottom half draws a similar comparison between MLPs (left) and ViTs (right).
+Moreover, the non-accentuated connections don't add to the number of trainable weights, as the same 4 are shared between all neuron-receptive-field-pairs.\\The bottom half draws a similar comparison between MLPs (left) and VTs (right).
 A single query vector is combined with all key vectors to get a scalar value for each position (blue boxes).
@@
 \paragraph{Architecture}
-ViTs were first introduced in \citeauthor{dosovitskiy2020image} (\citeyear{dosovitskiy2020image}), and are based on the \textit{encoder stack} of the \textit{Transformer} architecture \citep{vaswani2017attention}.
+VTs were first introduced in \citeauthor{dosovitskiy2020image} (\citeyear{dosovitskiy2020image}), and are based on the \textit{encoder stack} of the \textit{Transformer} architecture \citep{vaswani2017attention}.
 This stack takes a sequence of vectors as input, and is composed of alternating \textit{multi-headed self-attention} layers and position-wise MLP-type layers -- position-wise meaning the MLP can take only a single position (constituent vector) as input at a time, and is shared between all positions.
 
-The self-attention mechanism plays a central role in ViTs, and is worth elaborating on.
+The self-attention mechanism plays a central role in VTs, and is worth elaborating on.
 Each head of a self-attention layer contains 3 learned projection matrices, which are used to produce $key$, $value$, and $query$ vectors for every position in the input sequence.
@@
 The original Transformer architecture was designed to solve problems in \textit{natural language processing} (a field it now dominates).
-To make it work for CV applications, ViTs divide an image into square patches, which get treated as tokens (words).
-Learned embeddings convert these into a sequence of vectors, to which 1D positional encodings are added\footnote{Do mind that ViTs have no built-in knowledge of an image's 2D topology.}.
+To make it work for CV applications, VTs divide an image into square patches, which get treated as tokens (words).
+Learned embeddings convert these into a sequence of vectors, to which 1D positional encodings are added\footnote{Do mind that VTs have no built-in knowledge of an image's 2D topology.}.
 A learned $[class]$ vector gets prepended to the sequence, and its representation at the end of the encoder stack is the final output, which can then be used for classification, segmentation, etc.
@@
 
-%ViTs were first introduced in \citeauthor{dosovitskiy2020image} (\citeyear{dosovitskiy2020image}), and are based on the \textit{Transformer} architecture \citep{vaswani2017attention}.
+%VTs were first introduced in \citeauthor{dosovitskiy2020image} (\citeyear{dosovitskiy2020image}), and are based on the \textit{Transformer} architecture \citep{vaswani2017attention}.
 This type of ANN was originally designed to solve problems in \textit{natural language processing} (a field it now dominates).
 It takes a series of tokens as input, and uses a learned mapping to convert them into a series of vectors, called \textit{token embeddings}.
-These embeddings are then fed through an encoder and decoder, but since the latter is discarded for ViTs, it won't be discussed here.
+These embeddings are then fed through an encoder and decoder, but since the latter is discarded for VTs, it won't be discussed here.
 
@@
 For each it learns a vector representation, called a token embedding.} as input.
-While Transformers are made up of an \textit{encoder} and \textit{decoder}, the latter won't be discussed here, as it is discarded in the ViT architecture.
+While Transformers are made up of an \textit{encoder} and \textit{decoder}, the latter won't be discussed here, as it is discarded in the VT architecture.
 
@@
 
-%The architecture consists of an \textit{encoder}, and \textit{decoder}, but since the latter is discarded in ViTs, it won't be discussed here.
+%The architecture consists of an \textit{encoder}, and \textit{decoder}, but since the latter is discarded in VTs, it won't be discussed here.
 
@@
 
-% While a CNN is restricted to only incorporate information from its surrounding pixels in a subsequent layer, a ViT
+% While a CNN is restricted to only incorporate information from its surrounding pixels in a subsequent layer, a VT
 
-\paragraph{CNNs versus ViTs}
-Where CNNs reduce the number of trainable weights by taking an image's topology into account, ViTs do this by learning themselves what information they should incorporate in successive representations of the input.
+\paragraph{CNNs versus VTs}
+Where CNNs reduce the number of trainable weights by taking an image's topology into account, VTs do this by learning themselves what information they should incorporate in successive representations of the input.
 As such, they are not restricted to a local receptive field, but can potentially include information from all over the previous layer.
 
-Moreover, weight sharing applies as much to ViTs as it does to CNNs.
-In ViTs, each position goes through the same computational pipelines of being multiplied by 3 projection matrices, followed by a weighted sum calculation, followed by being fed through an MLP-type layer\footnote{Neglecting skip connections and normalization, as this is a comparison on an intuitive level.}.
+Moreover, weight sharing applies as much to VTs as it does to CNNs.
+In VTs, each position goes through the same computational pipelines of being multiplied by 3 projection matrices, followed by a weighted sum calculation, followed by being fed through an MLP-type layer\footnote{Neglecting skip connections and normalization, as this is a comparison on an intuitive level.}.
 In CNNs these pipelines take the form of convolution kernels.
 
-\paragraph{A limitation of ViTs}
-While ViTs are promising; the fact that they are not tailored to CV tasks, makes that they lack a so called image specific \textit{Inductive Bias}, which CNNs do have.
-Intuitively this means a ViT first has to learn what an image is, before it can move on to the learning task at hand.
+\paragraph{A limitation of VTs}
+While VTs are promising; the fact that they are not tailored to CV tasks, makes that they lack a so called image specific \textit{Inductive Bias}, which CNNs do have.
+Intuitively this means a VT first has to learn what an image is, before it can move on to the learning task at hand.
 The result is that they require substantially more training data (and time) than CNNs do.
 
-It is for this reason the paper that first introduced ViTs \citep{dosovitskiy2020image} mentions using pre-trained models rather than randomly initialised ones, as it allows general knowledge about images to be carried over to the new task.
+It is for this reason the paper that first introduced VTs \citep{dosovitskiy2020image} mentions using pre-trained models rather than randomly initialised ones, as it allows general knowledge about images to be carried over to the new task.
 This technique falls within the category of TL, which will be covered next.
@@
 % CNNs have image specific inductive bias: the architecture itself is tailored to 2D images, and as such has structural information about how images work built-in.
-This is not the case for ViTs, which lack an image specific inductive bias, which makes them very data hungry.
+This is not the case for VTs, which lack an image specific inductive bias, which makes them very data hungry.
 
@@
 \subsection{Related work and the current study} \label{related_work}
-As mentioned, the paper that first introduced ViTs already demonstrated they benefit from TL \citep{dosovitskiy2020image}, and more has been discovered about ViTs' TL properties since then.
-\citeauthor{matsoukas2021time} (\citeyear{matsoukas2021time}) investigated whether TL allowed ViTs to replace CNNs in Medical Imaging.
+As mentioned, the paper that first introduced VTs already demonstrated they benefit from TL \citep{dosovitskiy2020image}, and more has been discovered about VTs' TL properties since then.
+\citeauthor{matsoukas2021time} (\citeyear{matsoukas2021time}) investigated whether TL allowed VTs to replace CNNs in Medical Imaging.
 This domain is characterized by small datasets, so to overcome the inductive bias problem described earlier, TL is imperative.
-It was found that ViTs benefit more from FT TL than CNNs do, and that ViTs pre-trained on ImageNet perform on par CNNs.
+It was found that VTs benefit more from FT TL than CNNs do, and that VTs pre-trained on ImageNet perform on par CNNs.
 
-What \citeauthor{matsoukas2021time} did not investigate, is OTS performance of ViTs.
-This was, however, done in \citeauthor{zhou2021convnets} (\citeyear{zhou2021convnets}), which found that ViTs pre-trained on ImageNet had worse OTS performance than their CNN counterparts.
-An exception was the \textit{Swin-B} architecture, which was the best performing model on all datasets, and uses \textit{(shifted-)window-based} self-attention methods, to make Transformers better suited for CV tasks \citep{liu2021swin}.
+What \citeauthor{matsoukas2021time} did not investigate, is OTS performance of VTs.
+This was, however, done in \citeauthor{zhou2021convnets} (\citeyear{zhou2021convnets}), which found that VTs pre-trained on ImageNet had worse OTS performance than their CNN counterparts.
+An exception was the \texttt{Swin} architecture, which was the best performing model on all datasets, and uses \textit{(shifted-)window-based} self-attention methods, to make Transformers better suited for CV tasks \citep{liu2021swin}.
 % Not into detail about ots and other techniques
 
-Another ``shortcomming'' of \citeauthor{matsoukas2021time} is that it merely examined whether ViTs can use TL to outperform CNNs in a domain that happens to be characterized by limited-sized datasets, but that it did not actively focus on the extent to which ViTs can be made useful in limited data/compute environments.
-More related to this is \citeauthor{touvron2021training} (\citeyear{touvron2021training}), which proposed transformer-specific distillation-based learning techniques to make ViTs applicable for smaller datasets.
+Another ``shortcomming'' of \citeauthor{matsoukas2021time} is that it merely examined whether VTs can use TL to outperform CNNs in a domain that happens to be characterized by limited-sized datasets, but that it did not actively focus on the extent to which VTs can be made useful in limited data/compute environments.
+More related to this is \citeauthor{touvron2021training} (\citeyear{touvron2021training}), which proposed transformer-specific distillation-based learning techniques to make VTs applicable for smaller datasets.
 This is not necessarily related to TL as defined in this paper, however.
-% matsoukas2021time not specifically focussed on seeing to what degree TL allows ViTs to dominate CNNs in limited compute/data environments
+% matsoukas2021time not specifically focussed on seeing to what degree TL allows VTs to dominate CNNs in limited compute/data environments
 \\\\
-The current study will examine to what extent TL techniques (both FT and OTS) allow ViTs to become the preferred architecture when compute resources are limited, and attempts to find a methodology that is recommended in these circumstances.
+The current study will examine to what extent TL techniques (both FT and OTS) allow VTs to become the preferred architecture when compute resources are limited, and attempts to find a methodology that is recommended in these circumstances.
 It will, however, restrict itself to art classification tasks.
@@
 
+%TODO: talk about datasets small enough to be hand labelled by one person
 %The current study limits itself to the task of identifying which materials were used in artworks.
-For this, it will examine to what extent TL techniques (both fine-tuning and OTS) allow ViTs to become the preferred architecture when compute resources are limited, and attempt to find a methodology that is preferred in circumstances as described.
+For this, it will examine to what extent TL techniques (both fine-tuning and OTS) allow VTs to become the preferred architecture when compute resources are limited, and attempt to find a methodology that is preferred in circumstances as described.
 It will do this by means of several experiments, and both quantitative and qualitative analysis of the results.
 
-More formally, the following research question is proposed: \begin{quote}\textit{To what extent (and in what manners) does utilizing TL allow ViTs to perform on par with CNNs on common art classification problems, when compute resources are limited to a single GPU?}\end{quote}
+More formally, the following research question is proposed: \begin{quote}\textit{To what extent (and in what manners) does utilizing TL allow VTs to perform on par with CNNs on common art classification problems, when compute resources are limited to a single GPU?}\end{quote}

### 2022-06-30 writing/tex/draft_methods.tex
Continued working on writing y paper

@@
 This paper investigated TL properties of CNNs in the art classification domain.
-The main focus was on a comparison between CNN architectures pretrained on ImageNet1K when they were either used as OTS feature extractors, or fine-tuned to the new task.
+The main focus was on a comparison between CNN architectures pre-trained on ImageNet1K when they were either used as OTS feature extractors, or fine-tuned to the new task.
 It was suggested that FT resulted in substantially better performance than either training from scratch or OTS TL.
@@
 
+\begin{figure*}[tb]
+\includegraphics[width=\textwidth]{img/examples.png}
+\caption{Example images from the \textit{Rijksmuseum Challenge}.}
+\label{methods:eg}
+\end{figure*}
+
 Similar to \citeauthor{sabatelli2018deep}, used models are pre-trained on ImageNet1K.
@@
 This dataset consists of a large collection of digitized artworks, together with xml-formatted metadata.
-The classification tasks distilled from it are: (1) \textit{Type classification}, containing labels as `painting', `sculpture', `drawing', etc.; (2) \textit{Material classification}, with labels as `paper', `porcelain', `silver', etc.; and finally, (3) `Artist classification', where the model has to predict who the creator is.
+The classification tasks distilled from it are: (1) \textit{Type classification}, containing labels as `painting', `sculpture', `drawing', etc.; (2) \textit{Material classification}, with labels as `paper', `porcelain', `silver', etc.; and finally, (3) \textit{Artist classification}, where the model has to predict who the creator is.
+Figure \ref{methods:eg} shows labelled examples for all three tasks.
 
@@
 
-For robustness, five subsets are randomly generated per classification task. These are then split up into 80\%, 10\%, 10\% training, validating and testing sets, respectively.
+For robustness, five subsets are randomly generated per classification task.
+These are then split up into 80\%, 10\%, 10\% training, validating and testing sets, such that all experiments can be run five times.
+The average sample overlap between two of the five datasets is shown in table \ref{methods:datasets} as well.
 
-Other than the tasks described above, table \ref{methods:datasets} shows an additional experiment called `Scaling'.
-The goal of this experiment is to see how the findings of this paper hold up as dataset sizes are shrunk further.
+Other than the tasks described above, table \ref{methods:datasets} shows an additional experiment called \textit{Scaling}.
+The goal of this experiment is to see how the findings of this paper hold up as dataset sizes are gradually shrunk.
 It does this by taking the top 15 classes of type classification, and scaling it 4 times by a factor $\sqrt[4]{\frac{1}{10}} \approx 0.56$.
@@
 Steps are taken to ensure the distribution remains the same.
-Finally, 5 datasets are generated per scaling factor, leading to a total of 25 datasets.
+Finally, also here 5 datasets are generated per scaling factor, leading to a total of 25 datasets.
 The same 80\%, 10\%, 10\% split is used for all.
@@
 \subsection{Models} \label{methods:models}
-Eight different neural network architectures are used throughout this paper, of which four are CNN-based and four ViT-based.
+Eight different neural network architectures are used throughout this paper, of which four are CNN-based and four VT-based.
 When multiple versions are available (i.e.
@@
 
-To start off with CNNs, ResNet50 \citep{he2016deep} and VGG19 \citep{simonyan2014very} are chosen, as these were the best performing models in \citeauthor{sabatelli2018deep} for respectively FT and OTS TL.
-In addition, \citeauthor{matsoukas2021time} (\citeyear{matsoukas2021time}) used ResNet50 too, while \citeauthor{zhou2021convnets} (\citeyear{zhou2021convnets}) used ResNet101 and ResNet152.
-To also add some more recent CNNs, ConvNext \citep{liu2022convnet} -- a purely CNN-based model inspired by ViTs' recent success -- and EfficientNetV2 \citep{tan2021efficientnetv2} are included as well.
+To start off with CNNs, \texttt{ResNet50} \citep{he2016deep} and \texttt{VGG19} \citep{simonyan2014very} are chosen, as these were the best performing models in \citeauthor{sabatelli2018deep} for respectively FT and OTS TL.
+In addition, \citeauthor{matsoukas2021time} (\citeyear{matsoukas2021time}) used \texttt{ResNet50} too, while \citeauthor{zhou2021convnets} (\citeyear{zhou2021convnets}) used \texttt{ResNet101} and \texttt{ResNet152}.
+To also add some more recent CNNs, \texttt{ConvNext} \citep{liu2022convnet} -- a purely CNN-based model inspired by VTs' recent success -- and \texttt{EfficientNetV2} \citep{tan2021efficientnetv2} are included as well.
 
-For ViTs, versions with a 16x16 patch size are chosen.
-The first model is the original ViT \citep{dosovitskiy2020image}.
-Swin \citep{liu2021swin}, which was mentioned in section \ref{related_work}, is also selected, as it showed promising performance in \citeauthor{zhou2021convnets}.
-Section \ref{related_work} mentioned DeiT \citep{touvron2021training} too, as the distillation based architecture.
+For VTs, versions with a 16x16 patch size are chosen.
+The first model is the original VT \citep{dosovitskiy2020image}, which will be referred to as \texttt{ViT}.
+Next is \texttt{Swin} \citep{liu2021swin}, since it showed promising performance in \citeauthor{zhou2021convnets}.
+The final two are \texttt{DeiT} and \texttt{BeiT}.
+The small version of \texttt{DeiT} was used in \citeauthor{matsoukas2021time} (\citeyear{matsoukas2021time}), because it is similar in size to \texttt{ResNet50}, to which it was compared.
+In this work, however, also a base-sized version is chosen.
+Specifically, the version that does not include output for distillation learning.
+
+% Section \ref{related_work} mentioned DeiT \citep{touvron2021training} too, as the distillation based architecture.
 The small version of this architecture was used in \citeauthor{matsoukas2021time} (\citeyear{matsoukas2021time}), since it is similar in size to ResNet50, to which it was compared.
@@
 % TODO TODO TODO TODO TODO TODO TODO TODO YOU FORGOT TO MENTION BEIT!!!
-% TODO if you do end up using distilation with ConvNext as a teacher: don't forget to change this section!!! TODO 2: im not completely happy with the part about DeiT. Doesn't flow super nicely
+% TODO 2: im not completely happy with the part about DeiT. Doesn't flow super nicely
 
-All ViT-based models have a bit over 86 million parameters.
-ConvNext is of similar size, with 88.6 million parameters.
-ResNet50 and EfficientNetV2 are smaller, with 25.6 and 13.6 million, respectively, while VGG19 is the largest model, with 143.7 million parameters.
+All VT-based models have a bit over 86 million parameters.
+\texttt{ConvNext} is of similar size, with 88.6 million parameters.
+\texttt{ResNet50} and \texttt{EfficientNetV2} are smaller, with 25.6 and 13.6 million, respectively, while \texttt{VGG19} is the largest model, with 143.7 million parameters.
 It should be mentioned that for both OTS TL and FT, preliminary experiments were performed with `small' and `tiny' models.
@@
 %  - ResNet50 and VGG19 best fine-tuned and ots networks in Sabatelli
-%  - ResNet50 I think also used in medical imaging paper; ResNet101 ots vit paper
+%  - ResNet50 I think also used in medical imaging paper; ResNet101 ots vt paper
 %  * Convnext and EfficientNetV2 to have some more modern ones in the mix (to make comparison more fair)
-% Paragraph/section about ViT models -- explain the idea behind each in 1 sentence maybe (also for CNNs)
+% Paragraph/section about VT models -- explain the idea behind each in 1 sentence maybe (also for CNNs)
 
@@
 For all models, the linear classification layer is replaced as to fit the new task.
-Cross entropy loss uses the raw outputs of this layer (conform PyTorch documentation), together with one-hot encoded prediction targets.
+Cross entropy loss uses the raw outputs of this layer (conform \texttt{PyTorch} documentation), together with one-hot encoded prediction targets.
 An early stop occurs after 10 epochs without improved loss on the validation set, and the model with the lowest loss is benchmarked on the testing set.
 
-For OTS TL, the Adam optimizer is used with standard PyTorch parameters (\texttt{lr}=1e-3, $\beta_1=0.9$, $\beta_2=0.999$), and a batch size of 256.
-For FT, more is needed to achieve good convergence, especially for ViT-based architectures.
+For OTS TL, the Adam optimizer is used with standard \texttt{PyTorch} parameters (\texttt{lr}=1e-3, $\beta_1=0.9$, $\beta_2=0.999$), and a batch size of 256.
+For FT, more is needed to achieve good convergence, especially for VT-based architectures.
 Used hyperparameters, etc.
@@
 
-The PyTorch machine learning framework is used for all experiments, and many pre-trained models are taken from its Torchvision library.
-Exceptions are EfficientNetV2, Swin, DeiT and Beit, which are taken from the Timm library\footnote{\url{https://timm.fast.ai/}.}.
+The \texttt{PyTorch} machine learning framework is used for all experiments, and many pre-trained models are taken from its \texttt{Torchvision} library.
+Exceptions are \texttt{EfficientNetV2}, \texttt{Swin}, \texttt{DeiT} and \texttt{BeiT}, which are taken from the \texttt{Timm} library\footnote{\url{https://timm.fast.ai/}.}.
 

### 2022-06-30 writing/tex/draft_results.tex
Continued working on writing y paper

@@
 This section presents results of the conducted experiments, mostly in the form of tables and line plots.
-Tables show average testing performance over the 5 trials per experiment, with standard deviation ($s$) between brackets.
+Tables show average testing performance over the 5 trials per experiment, with standard deviation ($s$) as subscript.
 A green shaded cell marks best performance, while yellow and red depict second best and worst performance, respectively.
+Finally, both accuracy and balanced accuracy are reported.
+The latter is defined as average recall over all classes.
+Note how this penalises errors on less occurring classes more than plain accuracy would.
 
-Line plots mostly show average validation accuracy (dashed line is CNN-based, continuous is ViT-based), with \textit{standard error of the mean} ($\pm s \div \sqrt{N}$, where $N=5$) depicted as a shaded region.
+Line plots mostly show average validation accuracy (dashed line is CNN-based, continuous is VT-based), with \textit{standard error of the mean} ($\pm s \div \sqrt{N}$, where $N=5$) depicted as a shaded region.
 Plots end when an early stop occurred for the first of 5 trials.
@@
 
-\begin{figure*}
+\begin{figure*}[tb]
     \centering
-    \begin{subfigure}{0.32\textwidth}
-    \def\svgwidth{5.5cm}
-    \input{img/ots_type_acc.pdf_tex}
-    \caption{Type}
-    \label{results:img:ots_type}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}{0.32\textwidth}
-    \def\svgwidth{5.5cm}
-    \input{img/ots_mat_acc.pdf_tex}
-    \caption{Material}
-    \label{results:img:ots_mat}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}{0.32\textwidth}
-    \def\svgwidth{5.5cm}
-    \input{img/ots_artist_acc.pdf_tex}
-    \caption{Artist}
-    \label{results:img:ots_artist}
-    \end{subfigure}
+    \def\svgwidth{\textwidth}
+    \input{img/ots_mean_accuracy.pdf_tex}
     \caption{Validation accuracy using off the shelf models.}
@@
 
-\begin{table}[tb]
-\resizebox{\columnwidth}{!}{%
-\begin{tabular}{lll}
+\begin{table*}[tb]
+\centering
+\resizebox{\textwidth}{!}{%
+\begin{tabular}{lllllll}
 \hline
-\textbf{Model} & \textbf{Accuracy} & \textbf{Balanced accuracy} \\ \hline
-\textbf{vit\_b\_16} & 86.06\% (${\pm 1.06\%}$) & 84.13\% (${\pm 1.57\%}$)  \\
-\textbf{swin\_b} & \cellcolor{bestcol}89.43\% (${\pm 0.93\%}$) & \cellcolor{bestcol}87.47\% (${\pm 1.02\%}$)  \\
-\textbf{beit\_b\_16} & \cellcolor{worstcol}82.26\% (${\pm 0.72\%}$) & \cellcolor{worstcol}77.75\% (${\pm 0.27\%}$)  \\
-\textbf{deit\_b\_16} & 88.18\% (${\pm 0.66\%}$) & 85.36\% (${\pm 0.54\%}$)  \\\hdashline
-\textbf{vgg19} & 83.93\% (${\pm 0.72\%}$) & 83.35\% (${\pm 0.81\%}$)  \\
-\textbf{resnet50} & 85.51\% (${\pm 0.64\%}$) & 82.33\% (${\pm 1.85\%}$)  \\
-\textbf{efficientnetv2\_m} & 83.41\% (${\pm 0.76\%}$) & 82.05\% (${\pm 1.25\%}$)  \\
-\textbf{convnext\_b} & \cellcolor{secondbestcol}89.19\% (${\pm 0.64\%}$) & \cellcolor{secondbestcol}86.95\% (${\pm 1.38\%}$)  \\\hline
+\textbf{Model} & \textbf{Type} &  & \textbf{Material} & & \textbf{Artist} & \\
+& \textbf{Accuracy} & \textbf{Bal. accuracy} & \textbf{Accuracy} & \textbf{Bal. accuracy} & \textbf{Accuracy} & \textbf{Bal. accuracy} \\ \hline
+\textbf{vit\_b\_16} & 86.06\% $_{\pm 1.06\%}$ & 84.13\% $_{\pm 1.57\%}$ & 81.78\% $_{\pm 0.48\%}$ & 67.38\% $_{\pm 1.37\%}$ & 84.89\% $_{\pm 0.46\%}$ & 81.42\% $_{\pm 0.42\%}$  \\
+\textbf{swin\_b} & \cellcolor{bestcol}89.43\% $_{\pm 0.93\%}$ & \cellcolor{bestcol}87.47\% $_{\pm 1.02\%}$ & \cellcolor{bestcol}85.87\% $_{\pm 0.35\%}$ & \cellcolor{bestcol}71.19\% $_{\pm 1.60\%}$ & \cellcolor{bestcol}90.40\% $_{\pm 0.65\%}$ & \cellcolor{bestcol}88.64\% $_{\pm 0.78\%}$  \\
+\textbf{beit\_b\_16} & \cellcolor{worstcol}82.26\% $_{\pm 0.72\%}$ & \cellcolor{worstcol}77.75\% $_{\pm 0.27\%}$ & 76.87\% $_{\pm 0.96\%}$ & 60.16\% $_{\pm 1.56\%}$ & 79.70\% $_{\pm 0.69\%}$ & 75.35\% $_{\pm 1.09\%}$  \\
+\textbf{deit\_b\_16} & 88.18\% $_{\pm 0.66\%}$ & 85.36\% $_{\pm 0.54\%}$ & 82.80\% $_{\pm 1.12\%}$ & 66.46\% $_{\pm 1.03\%}$ & 88.13\% $_{\pm 0.76\%}$ & 85.62\% $_{\pm 0.87\%}$  \\\hdashline
+\textbf{vgg19} & 83.93\% $_{\pm 0.72\%}$ & 83.35\% $_{\pm 0.81\%}$ & 76.87\% $_{\pm 0.44\%}$ & 61.39\% $_{\pm 1.47\%}$ & 82.01\% $_{\pm 0.66\%}$ & 78.10\% $_{\pm 0.77\%}$  \\
+\textbf{resnet50} & 85.51\% $_{\pm 0.64\%}$ & 82.33\% $_{\pm 1.85\%}$ & 80.99\% $_{\pm 0.82\%}$ & 65.51\% $_{\pm 0.93\%}$ & 87.71\% $_{\pm 1.06\%}$ & 85.12\% $_{\pm 1.34\%}$  \\
+\textbf{eff.
+netv2\_m} & 83.41\% $_{\pm 0.76\%}$ & 82.05\% $_{\pm 1.25\%}$  & \cellcolor{worstcol}75.96\% $_{\pm 1.24\%}$ & \cellcolor{worstcol}59.15\% $_{\pm 1.24\%}$ & \cellcolor{worstcol}78.62\% $_{\pm 1.07\%}$ & \cellcolor{worstcol}73.92\% $_{\pm 0.96\%}$  \\
+\textbf{convnext\_b} & \cellcolor{secondbestcol}89.19\% $_{\pm 0.64\%}$ & \cellcolor{secondbestcol}86.95\% $_{\pm 1.38\%}$ & \cellcolor{secondbestcol}84.14\% $_{\pm 0.92\%}$ & \cellcolor{secondbestcol}69.10\% $_{\pm 1.05\%}$ & \cellcolor{secondbestcol}90.13\% $_{\pm 0.94\%}$ & \cellcolor{secondbestcol}87.84\% $_{\pm 1.07\%}$  \\\hline
 \end{tabular}
 }
-\caption{Type classification off the shelf}
-\label{results:tab:ots_type}
-\end{table}
+\caption{Testing performance after off the shelf learning.}
+\label{results:tab:ots}
+\end{table*}
 
-\begin{table}[tb]
-\resizebox{\columnwidth}{!}{%
-\begin{tabular}{lll}
-\hline
-\textbf{Model} & \textbf{Accuracy} & \textbf{Balanced accuracy} \\ \hline
-\textbf{vit\_b\_16} & 81.78\% (${\pm 0.48\%}$) & 67.38\% (${\pm 1.37\%}$)  \\
-\textbf{swin\_b} & \cellcolor{bestcol}85.87\% (${\pm 0.35\%}$) & \cellcolor{bestcol}71.19\% (${\pm 1.60\%}$)  \\
-\textbf{beit\_b\_16} & 76.87\% (${\pm 0.96\%}$) & 60.16\% (${\pm 1.56\%}$)  \\
-\textbf{deit\_b\_16} & 82.80\% (${\pm 1.12\%}$) & 66.46\% (${\pm 1.03\%}$)  \\\hdashline
-\textbf{vgg19} & 76.87\% (${\pm 0.44\%}$) & 61.39\% (${\pm 1.47\%}$)  \\
-\textbf{resnet50} & 80.99\% (${\pm 0.82\%}$) & 65.51\% (${\pm 0.93\%}$)  \\
-\textbf{efficientnetv2\_m} & \cellcolor{worstcol}75.96\% (${\pm 1.24\%}$) & \cellcolor{worstcol}59.15\% (${\pm 1.24\%}$)  \\
-\textbf{convnext\_b} & \cellcolor{secondbestcol}84.14\% (${\pm 0.92\%}$) & \cellcolor{secondbestcol}69.10\% (${\pm 1.05\%}$)  \\\hline
-\end{tabular}
-}
-\caption{Material classification off the shelf}
-\label{results:tab:ots_mat}
-\end{table}
+Table \ref{results:tab:ots} shows testing performance of OTS trained models on the 3 classification tasks, while figure \ref{results:img:ots} shows corresponding validation accuracies.
+Performance on the testing set is comparable among all tasks in terms of ranking.
+Other observations are consistent as well, such as \texttt{ConvNext} taking the longest to converge.
 
-\begin{table}[tb]
-\resizebox{\columnwidth}{!}{%
-\begin{tabular}{lll}
-\hline
-\textbf{Model} & \textbf{Accuracy} & \textbf{Balanced accuracy} \\ \hline
-\textbf{vit\_b\_16} & 84.89\% (${\pm 0.46\%}$) & 81.42\% (${\pm 0.42\%}$)  \\
-\textbf{swin\_b} & \cellcolor{bestcol}90.40\% (${\pm 0.65\%}$) & \cellcolor{bestcol}88.64\% (${\pm 0.78\%}$)  \\
-\textbf{beit\_b\_16} & 79.70\% (${\pm 0.69\%}$) & 75.35\% (${\pm 1.09\%}$)  \\
-\textbf{deit\_b\_16} & 88.13\% (${\pm 0.76\%}$) & 85.62\% (${\pm 0.87\%}$)  \\\hdashline
-\textbf{vgg19} & 82.01\% (${\pm 0.66\%}$) & 78.10\% (${\pm 0.77\%}$)  \\
-\textbf{resnet50} & 87.71\% (${\pm 1.06\%}$) & 85.12\% (${\pm 1.34\%}$)  \\
-\textbf{efficientnetv2\_m} & \cellcolor{worstcol}78.62\% (${\pm 1.07\%}$) & \cellcolor{worstcol}73.92\% (${\pm 0.96\%}$)  \\
-\textbf{convnext\_b} & \cellcolor{secondbestcol}90.13\% (${\pm 0.94\%}$) & \cellcolor{secondbestcol}87.84\% (${\pm 1.07\%}$)  \\\hline
-\end{tabular}
-}
-\caption{Artist classification off the shelf}
-\label{results:tab:ots_artist}
-\end{table}
+\texttt{Swin} is coming out on top in all cases, which is in line with results of \citeauthor{zhou2021convnets} (\citeyear{zhou2021convnets}).
+What slightly deviates from these result, is the observation that \texttt{ViT} performs better than a \texttt{ResNet}-based model (on 2 out of 3 tasks).
+This comparison is, however, not meaningful, as \citeauthor{zhou2021convnets} uses \texttt{ResNet101} and 152.
+Testing these models instead, gives 86.21\% ($\pm$ 0.70\%) accuracy for \texttt{ResNet101}, and 86.19\% ($\pm$ 0.60\%) for \texttt{ResNet152}, which is indeed slightly higher than \texttt{ViT}'s 86.06\% shown in table \ref{results:tab:ots}.
 
-Tables \ref{results:tab:ots_type} through \ref{results:tab:ots_artist} show testing performance of OTS trained models on the 3 classification tasks, while figure \ref{results:img:ots} shows corresponding validation accuracies.
-Performance on the testing set is comparable among all tasks in terms of ranking.
-Other observations are consistent as well, such as ConvNext taking the longest to converge.
-
-Swin is coming out on top in all cases, which is in line with results of \citeauthor{zhou2021convnets} (\citeyear{zhou2021convnets}).
-What slightly deviates from these result, is the observation that ViT performs better than a ResNet-based model (on 2 out of 3 tasks).
-This comparison is, however, not meaningful, as \citeauthor{zhou2021convnets} uses ResNet101 and 152.
-Using these models instead, gives 86.21\% ($\pm$ 0.70\%) accuracy for ResNet101, and 86.19\% ($\pm$ 0.60\%) for ResNet152, which is indeed slightly higher than ViT's 86.06\% shown in table \ref{results:tab:ots_type}.
-
-It is interesting that VGG19 performs worse than ResNet50, since it was the best performing OTS model on all 3 tasks in \citeauthor{sabatelli2018deep} (\citeyear{sabatelli2018deep}).
+It is interesting that \texttt{VGG19} performs worse than \texttt{ResNet50}, since it was the best performing OTS model on all 3 tasks in \citeauthor{sabatelli2018deep} (\citeyear{sabatelli2018deep}).
 The current study uses smaller datasets and fewer classes.
-The difference might be attributable to this, as VGG19 is the largest model (see section \ref{methods:models}), and has the largest final layer -- with 4096 inputs.
+The difference might be attributable to this, as \texttt{VGG19} is the largest model (see section \ref{methods:models}), and has the largest final layer -- with 4096 inputs.
 
@@
 
-As a whole, ViTs are doing well for the OTS experiments.
-Especially Swin and DeiT are ranked high on all testing sets, while ConvNext is the most promising CNN-based architecture.
-Taking the mean accuracy of all ViT- and CNN-based models, reveals that ViTs as a group perform slightly better.
+As a whole, VTs are doing well for the OTS experiments.
+Especially \texttt{Swin} and \texttt{DeiT} are ranked high on all testing sets, while \texttt{ConvNext} is the most promising CNN-based architecture.
+Taking the mean accuracy of all VT- and CNN-based models, reveals that VTs as a group perform slightly better.
 On type classification, for example, their mean accuracy is 86.5\%, while it is 85.5\% for CNNs.
@@
 %    Comparisons with Sabatelli paper (ResNet50 vs VGG19)
-%    ViTs slightly better
+%    VTs slightly better
 %    Lower balanced accuracy
@@
 
-\begin{figure*}
+\begin{figure*}[tb]
     \centering
-    \begin{subfigure}{0.32\textwidth}
-    \def\svgwidth{5.5cm}
-    \input{img/ft_type_acc.pdf_tex}
-    \caption{Type}
-    \label{results:img:ft_type}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}{0.32\textwidth}
-    \def\svgwidth{5.5cm}
-    \input{img/ft_mat_acc.pdf_tex}
-    \caption{Material}
-    \label{results:img:ft_mat}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}{0.32\textwidth}
-    \def\svgwidth{5.5cm}
-    \input{img/ft_artist_acc.pdf_tex}
-    \caption{Artist}
-    \label{results:img:ft_artist}
-    \end{subfigure}
+    \def\svgwidth{\textwidth}
+    \input{img/ft_mean_accuracy.pdf_tex}
     \caption{Validation accuracy using fine-tuned models.}
@@
 
-\begin{table}[tb]
-\resizebox{\columnwidth}{!}{%
-\begin{tabular}{lll}
+\begin{table*}[tb]
+\centering
+\resizebox{\textwidth}{!}{%
+\begin{tabular}{lllllll}
 \hline
-\textbf{Model} & \textbf{Accuracy} & \textbf{Balanced accuracy} \\ \hline
-\textbf{vit\_b\_16} & \cellcolor{worstcol}90.11\% (${\pm 0.35\%}$) & 87.40\% (${\pm 0.29\%}$)  \\
-\textbf{swin\_b} & \cellcolor{bestcol}92.17\% (${\pm 0.98\%}$) & \cellcolor{secondbestcol}89.71\% (${\pm 1.03\%}$)  \\
-\textbf{beit\_b\_16} & 90.81\% (${\pm 0.41\%}$) & 87.95\% (${\pm 0.69\%}$)  \\
-\textbf{deit\_b\_16} & 91.78\% (${\pm 0.64\%}$) & 89.22\% (${\pm 0.90\%}$)  \\\hdashline
-\textbf{vgg19} & 90.54\% (${\pm 0.37\%}$) & \cellcolor{worstcol}87.05\% (${\pm 1.03\%}$)  \\
-\textbf{resnet50} & 91.78\% (${\pm 0.44\%}$) & 88.24\% (${\pm 0.59\%}$)  \\
-\textbf{efficientnetv2\_m} & 90.87\% (${\pm 0.67\%}$) & 88.34\% (${\pm 1.37\%}$)  \\
-\textbf{convnext\_b} & \cellcolor{secondbestcol}92.15\% (${\pm 0.40\%}$) & \cellcolor{bestcol}89.82\% (${\pm 1.18\%}$)  \\\hline
+\textbf{Model} & \textbf{Type} &  & \textbf{Material} & & \textbf{Artist} & \\
+& \textbf{Accuracy} & \textbf{Bal. accuracy} & \textbf{Accuracy} & \textbf{Bal. accuracy} & \textbf{Accuracy} & \textbf{Bal. accuracy} \\ \hline
+\textbf{vit\_b\_16} & \cellcolor{worstcol}90.11\% $_{\pm 0.35\%}$ & 87.40\% $_{\pm 0.29\%}$ & 87.42\% $_{\pm 0.45\%}$ & 73.46\% $_{\pm 1.44\%}$ & 92.05\% $_{\pm 0.44\%}$ & 89.77\% $_{\pm 0.38\%}$  \\
+\textbf{swin\_b} & \cellcolor{bestcol}92.17\% $_{\pm 0.98\%}$ & \cellcolor{secondbestcol}89.71\% $_{\pm 1.03\%}$ & \cellcolor{bestcol}89.35\% $_{\pm 0.68\%}$ & 77.16\% $_{\pm 2.98\%}$ & \cellcolor{bestcol}95.05\% $_{\pm 0.47\%}$ & \cellcolor{bestcol}93.94\% $_{\pm 0.81\%}$  \\
+\textbf{beit\_b\_16} & 90.81\% $_{\pm 0.41\%}$ & 87.95\% $_{\pm 0.69\%}$ & \cellcolor{worstcol}85.74\% $_{\pm 0.37\%}$ & \cellcolor{worstcol}72.12\% $_{\pm 1.38\%}$ & \cellcolor{worstcol}91.27\% $_{\pm 1.13\%}$ & \cellcolor{worstcol}88.83\% $_{\pm 1.63\%}$  \\
+\textbf{deit\_b\_16} & 91.78\% $_{\pm 0.64\%}$ & 89.22\% $_{\pm 0.90\%}$ & 87.85\% $_{\pm 1.12\%}$ & 74.42\% $_{\pm 1.99\%}$ & 93.37\% $_{\pm 1.15\%}$ & 91.67\% $_{\pm 1.55\%}$  \\\hdashline
+\textbf{vgg19} & 90.54\% $_{\pm 0.37\%}$ & \cellcolor{worstcol}87.05\% $_{\pm 1.03\%}$ & 85.74\% $_{\pm 1.40\%}$ & 72.43\% $_{\pm 3.03\%}$ & 92.20\% $_{\pm 0.49\%}$ & 90.18\% $_{\pm 0.72\%}$  \\
+\textbf{resnet50} & 91.78\% $_{\pm 0.44\%}$ & 88.24\% $_{\pm 0.59\%}$ & 88.69\% $_{\pm 0.99\%}$ & \cellcolor{secondbestcol}77.97\% $_{\pm 2.25\%}$ & \cellcolor{secondbestcol}94.72\% $_{\pm 0.74\%}$ & \cellcolor{secondbestcol}93.41\% $_{\pm 1.05\%}$  \\
+\textbf{eff. netv2\_m} & 90.87\% $_{\pm 0.67\%}$ & 88.34\% $_{\pm 1.37\%}$ & 87.55\% $_{\pm 1.15\%}$ & 75.31\% $_{\pm 1.60\%}$ & 92.65\% $_{\pm 0.54\%}$ & 90.84\% $_{\pm 0.51\%}$  \\
+\textbf{convnext\_b} & \cellcolor{secondbestcol}92.15\% $_{\pm 0.40\%}$ & \cellcolor{bestcol}89.82\% $_{\pm 1.18\%}$ & \cellcolor{secondbestcol}88.79\% $_{\pm 1.07\%}$ & \cellcolor{bestcol}78.40\% $_{\pm 1.26\%}$ & 94.60\% $_{\pm 0.54\%}$ & 93.13\% $_{\pm 0.61\%}$  \\\hline
 \end{tabular}
 }
-\caption{Type classification fine-tuned}
-\label{results:tab:ft_type}
-\end{table}
+\caption{Testing performance after fine-tuning.}
+\label{results:tab:ft}
+\end{table*}
 
-\begin{table}[tb]
-\resizebox{\columnwidth}{!}{%
-\begin{tabular}{lll}
-\hline
-\textbf{Model} & \textbf{Accuracy} & \textbf{Balanced accuracy} \\ \hline
-\textbf{vit\_b\_16} & 87.42\% (${\pm 0.45\%}$) & 73.46\% (${\pm 1.44\%}$)  \\
-\textbf{swin\_b} & \cellcolor{bestcol}89.35\% (${\pm 0.68\%}$) & 77.16\% (${\pm 2.98\%}$)  \\
-\textbf{beit\_b\_16} & \cellcolor{worstcol}85.74\% (${\pm 0.37\%}$) & \cellcolor{worstcol}72.12\% (${\pm 1.38\%}$)  \\
-\textbf{deit\_b\_16} & 87.85\% (${\pm 1.12\%}$) & 74.42\% (${\pm 1.99\%}$)  \\\hdashline
-\textbf{vgg19} & 85.74\% (${\pm 1.40\%}$) & 72.43\% (${\pm 3.03\%}$)  \\
-\textbf{resnet50} & 88.69\% (${\pm 0.99\%}$) & \cellcolor{secondbestcol}77.97\% (${\pm 2.25\%}$)  \\
-\textbf{efficientnetv2\_m} & 87.55\% (${\pm 1.15\%}$) & 75.31\% (${\pm 1.60\%}$)  \\
-\textbf{convnext\_b} & \cellcolor{secondbestcol}88.79\% (${\pm 1.07\%}$) & \cellcolor{bestcol}78.40\% (${\pm 1.26\%}$)  \\\hline
-\end{tabular}
-}
-\caption{Material classification fine-tuned}
-\label{results:tab:ft_mat}
-\end{table}
-
-\begin{table}[tb]
-\resizebox{\columnwidth}{!}{%
-\begin{tabular}{lll}
-\hline
-\textbf{Model} & \textbf{Accuracy} & \textbf{Balanced accuracy} \\ \hline
-\textbf{vit\_b\_16} & 92.05\% (${\pm 0.44\%}$) & 89.77\% (${\pm 0.38\%}$)  \\
-\textbf{swin\_b} & \cellcolor{bestcol}95.05\% (${\pm 0.47\%}$) & \cellcolor{bestcol}93.94\% (${\pm 0.81\%}$)  \\
-\textbf{beit\_b\_16} & \cellcolor{worstcol}91.27\% (${\pm 1.13\%}$) & \cellcolor{worstcol}88.83\% (${\pm 1.63\%}$)  \\
-\textbf{deit\_b\_16} & 93.37\% (${\pm 1.15\%}$) & 91.67\% (${\pm 1.55\%}$)  \\\hdashline
-\textbf{vgg19} & 92.20\% (${\pm 0.49\%}$) & 90.18\% (${\pm 0.72\%}$)  \\
-\textbf{resnet50} & \cellcolor{secondbestcol}94.72\% (${\pm 0.74\%}$) & \cellcolor{secondbestcol}93.41\% (${\pm 1.05\%}$)  \\
-\textbf{efficientnetv2\_m} & 92.65\% (${\pm 0.54\%}$) & 90.84\% (${\pm 0.51\%}$)  \\
-\textbf{convnext\_b} & 94.60\% (${\pm 0.54\%}$) & 93.13\% (${\pm 0.61\%}$)  \\\hline
-\end{tabular}
-}
-\caption{Artist classification fine-tuned}
-\label{results:tab:ft_artist}
-\end{table}
-
-Once again, tables \ref{results:tab:ft_type} through \ref{results:tab:ft_artist} show testing performance, while figure \ref{results:img:ft} shows corresponding validation accuracies.
+Once again, table \ref{results:tab:ft} shows testing performance, while figure \ref{results:img:ft} shows corresponding validation accuracies.
 
 Similar to OTS TL, findings are somewhat consistent across classification tasks, while overall performance is again lowest for material classification, and highest for artist classification.
-Swin is still dominant in terms of testing accuracy, but Conv\-Next is catching up.
+\texttt{Swin} is still dominant in terms of testing accuracy, but Conv\-Next is catching up.
 This is especially true for balanced accuracy, where it tops the ranking on 2 of the 3 tasks.
-More general, it looks as if the advantage ViTs seemed to have for OTS TL is gone.
-Mean accuracy of ViTs grouped together is, for example, 91.2\% on type classification, while it is 91.3\% for CNNs.
+More general, it looks as if the advantage VTs seemed to have for OTS TL is gone.
+Mean accuracy of VTs grouped together is, for example, 91.2\% on type classification, while it is 91.3\% for CNNs.
 
-Different from \citeauthor{matsoukas2021time} (\citeyear{matsoukas2021time}), ResNet50 performs slightly better than DeiT.
-This remains true when using the same `tiny' version of DeiT as that paper used.
-On type classification, for example, it gets 90.44\% ($\pm$ 0.93\%) accuracy, whereas ResNet50 shows 91.78\% in table \ref{results:tab:ft_type}.
+Different from \citeauthor{matsoukas2021time} (\citeyear{matsoukas2021time}), \texttt{ResNet50} performs slightly better than \texttt{DeiT}.
+This remains true when using the same `tiny' version of \texttt{DeiT} as that paper used.
+On type classification, for example, it gets 90.44\% ($\pm$ 0.93\%) accuracy, whereas \texttt{ResNet50} shows 91.78\% in table \ref{results:tab:ft}.
 
-FT results also seem to be less favourable for ViTs compared to \citeauthor{zhou2021convnets} (\citeyear{zhou2021convnets}).
-ViT consistently falls behind ResNet50, and even shows the worst accuracy in table \ref{results:tab:ft_type} (90.11\%).
-It should be mentioned that ResNet101 and 152  -- which were used alongside ViT in \citeauthor{zhou2021convnets} --  show performance similar to ResNet50, with type classification accuracies of 91.96\% ($\pm$ 0.59\%) and 91.94\% ($\pm$ 0.25), respectively.
+FT results in this work also seem less favourable for VTs compared to \citeauthor{zhou2021convnets} (\citeyear{zhou2021convnets}).
+\texttt{ViT} consistently falls behind \texttt{ResNet50}, and even shows the worst accuracy in table \ref{results:tab:ft} (90.11\%), while in \citeauthor{zhou2021convnets} it was \texttt{ViT} that outperformed \texttt{ResNet101} and 152.
+It should be mentioned that these two alternatives show performance similar to \texttt{ResNet50}, with type classification accuracies of 91.96\% ($\pm$ 0.59\%) and 91.94\% ($\pm$ 0.25), respectively.
+Lastly, what is consistent with \citeauthor{zhou2021convnets}, is the good performance of \texttt{Swin} mentioned above.
 
@@
 In addition, the worst FT model is often better than the best OTS model.
-The only exception is material classification, where the worst FT model (BeiT) has accuracy 85.74\%, and the best OTS one (Swin) has 85.87\%.
+The only exception is material classification, where the worst FT model (\texttt{BeiT}) has accuracy 85.74\%, and the best OTS one (\texttt{Swin}) has 85.87\%.
 
-VGG19 has dropped further in the rankings, and now shows the worst balanced accuracy for type classification.
-ResNet50, on the other hand, appears to flourish with FT, often even beating ConvNext in the rankings.
+\texttt{VGG19} has dropped further in the rankings, and now shows the worst balanced accuracy for type classification.
+\texttt{ResNet50}, on the other hand, appears to flourish with FT, often even beating Conv\-Next in the rankings.
 This is consistent with \citeauthor{sabatelli2018deep} (\citeyear{sabatelli2018deep}), where it was the best FT model overall.
@@
 
-\begin{figure*}
+\begin{figure*}[tb]
     \centering
-    \begin{subfigure}{0.45\textwidth}
-    \def\svgwidth{7.7cm}
-    \input{img/ots_scale_acc.pdf_tex}
-    \caption{Off the shelf}
-    \label{results:img:ots_scale}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}{0.45\textwidth}
-    \def\svgwidth{7.7cm}
-    \input{img/ft_scale_acc.pdf_tex}
-    \caption{Fine tuning}
-    \label{results:img:ft_scale}
-    \end{subfigure}
-    \caption{Testing accuracy as datasets become smaller.}
+    \def\svgwidth{\textwidth}
+    \input{img/scaling_accuracy.pdf_tex}
+    \caption{Testing accuracy as datasets gradually become smaller.}
     \label{results:img:scale}
 \end{figure*}
+
+%\begin{figure*}
+%    \centering
+%    \begin{subfigure}{0.45\textwidth}
+%    \def\svgwidth{7.7cm}
+%    \input{img/ots_scale_acc.pdf_tex}
+%    \caption{Off the shelf}
+%    \label{results:img:ots_scale}
+%    \end{subfigure}
+%    \hfill
+%    \begin{subfigure}{0.45\textwidth}
+%    \def\svgwidth{7.7cm}
+%    \input{img/ft_scale_acc.pdf_tex}
+%    \caption{Fine tuning}
+%    \label{results:img:ft_scale}
+%    \end{subfigure}
+%    \caption{Testing accuracy as datasets become smaller.}
+%    \label{results:img:scale}
+%\end{figure*}
 
@@
 
-For both OTS TL in figure \ref{results:img:ots_scale}, as FT TL in figure \ref{results:img:ft_scale}, findings done in sections \ref{results:ots} and \ref{results:ft} seem to hold up as dataset sizes are shrunk.
+For both OTS TL as FT TL, findings done in sections \ref{results:ots} and \ref{results:ft} seem to hold up as dataset sizes are shrunk.
 When going from largest to smallest dataset, the mean accuracy of all models taken together drops with 9.1\% for OTS TL.
@@
 
-\begin{figure}
+\begin{figure}[tbh]
     \centering
@@
     \input{img/ots_vs_ft_type.pdf_tex}
-    \caption{Time/accuracy trade-offs for ViTs and CNNs either off the shelf or fine-tuned.}
+    \caption{Time/accuracy trade-offs for VTs and CNNs either off the shelf or fine-tuned.}
     \label{results:img:ots_vs_ft_type}
@@
 
-Figure \ref{results:img:ots_vs_ft_type} shows type classification accuracy as reported in tables \ref{results:tab:ots_type} and \ref{results:tab:ft_type}.
+Figure \ref{results:img:ots_vs_ft_type} shows type classification accuracy as reported in tables \ref{results:tab:ots} and \ref{results:tab:ft}.
 These are plotted against average epoch duration, as to get a sense of time/accuracy trade-offs.
-Colours make it easy to distinguish ViTs from CNNs, and OTS models from FT ones.
+Colours make it easy to distinguish VTs from CNNs, and OTS models from FT ones.
 
-OTS ViTs all take roughly the same time per epoch.
-CNNs are generally faster here, but accuracy also clearly falls behind for all but one (ConvNext).
-For FT, accuracies seem on-par, while CNNs still take much shorter to train.
-One CNN (ResNet50) even is faster than any OTS ViT, while also showing better accuracy.
-In short, FT CNNs have better time/accuracy trade-offs, while it is a FT ViT (Swin) that achieves the best accuracy overall.
+OTS VTs all take roughly the same time per epoch.
+CNNs are generally faster here, but accuracy also clearly falls behind for all but one (\texttt{ConvNext}).
+For FT, VT and CNN accuracies seem on-par, while CNNs still take much shorter to train.
+One CNN (\texttt{ResNet50}) even is faster than any OTS VT, while also showing better accuracy.
+In short, FT CNNs have better time/accuracy trade-offs, while it is a FT VT (\texttt{Swin}) that achieves the best accuracy overall.
 

### 2022-07-11 writing/tex/draft_experiments.tex
Rewriting results section -> Now its an 'Experiments' section with 'Results' subsection. Some small tweaks as well

--- 2022-06-30 writing/tex/draft_results.tex
@@
-\section{Results}
+\section{Experiments}
+This section focuses on the results gotten from the conducted experiments.
+These will first briefly be presented in section \ref{results}.
+After that, section \ref{exp:int} will discuss them in more detail, and see how they compare to previous works, etc.
 
-This section presents results of the conducted experiments, mostly in the form of tables and line plots.
-Tables show average testing performance over the 5 trials per experiment, with standard deviation ($s$) as subscript.
-A green shaded cell marks best performance, while yellow and red depict second best and worst performance, respectively.
-Finally, both accuracy and balanced accuracy are reported.
-The latter is defined as average recall over all classes.
-Note how this penalises errors on less occurring classes more than plain accuracy would.
+\subsection{Results} \label{results}
+Results are presented in the form of tables and line plots. The former of these report final testing performance, whereas the latter mostly visualise validation accuracy during training.
 
-Line plots mostly show average validation accuracy (dashed line is CNN-based, continuous is VT-based), with \textit{standard error of the mean} ($\pm s \div \sqrt{N}$, where $N=5$) depicted as a shaded region.
-Plots end when an early stop occurred for the first of 5 trials.
+To give a more precise description: tables show average testing performance over the 5 trials per experiment, with standard deviation ($s$) as subscript.
+Besides accuracy, also balanced accuracy is reported.
+This measure is defined as the average recall over all classes, such that it penalises errors on less occurring classes more than plain accuracy would.
+Lastly, green shaded cells mark best performance, while yellow and red mark second best and worst performance, respectively.
 
-% Tell that the appendix shows all other plots as well maybe. Only for the BP paper tho.
+For line plots, shaded regions correspond to the \textit{standard error of the mean} ($\pm s \div \sqrt{N}$, where $N=5$).
+In cases where they show average validation accuracy, plots end when an early stop occurred for the first of 5 trials.
+To easily distinguish VTs from CNNs, VTs are plotted with continues lines, and CNNs with dashed ones.
 
-\subsection{Off the shelf learning} \label{results:ots}
+\subsubsection{Off the shelf learning} \label{results:ots}
 
@@
     \input{img/ots_mean_accuracy.pdf_tex}
-    \caption{Validation accuracy using off the shelf models.}
+    \caption{Average validation accuracy when training with an off the shelf learning scheme.
+It can be observed that the \texttt{Swin} VT achieves the highest performance overall, but with the \texttt{ConvNext} CNN being a close second.
+More general, VTs perform on par with CNNs, if not slightly better.}
     \label{results:img:ots}
@@
 }
-\caption{Testing performance after off the shelf learning.}
+\caption{Testing performance after off the shelf learning.
+Results are similar to validation accuracy shown in figure \ref{results:img:ots}, with \texttt{Swin} and \texttt{ConvNext} showing respective best and second best performance in all cases.}
 \label{results:tab:ots}
@@
 Table \ref{results:tab:ots} shows testing performance of OTS trained models on the 3 classification tasks, while figure \ref{results:img:ots} shows corresponding validation accuracies.
-Performance on the testing set is comparable among all tasks in terms of ranking.
-Other observations are consistent as well, such as \texttt{ConvNext} taking the longest to converge.
+Performance on the testing set is comparable among all tasks in terms of rankings.
+Other observations are consistent as well, such as \texttt{ConvNext} taking the most epochs to converge.
 
-\texttt{Swin} is coming out on top in all cases, which is in line with results of \citeauthor{zhou2021convnets} (\citeyear{zhou2021convnets}).
-What slightly deviates from these result, is the observation that \texttt{ViT} performs better than a \texttt{ResNet}-based model (on 2 out of 3 tasks).
-This comparison is, however, not meaningful, as \citeauthor{zhou2021convnets} uses \texttt{ResNet101} and 152.
-Testing these models instead, gives 86.21\% ($\pm$ 0.70\%) accuracy for \texttt{ResNet101}, and 86.19\% ($\pm$ 0.60\%) for \texttt{ResNet152}, which is indeed slightly higher than \texttt{ViT}'s 86.06\% shown in table \ref{results:tab:ots}.
+VTs perform very much on par with CNNs, if not slightly better.
+The best performing model, for instance, is the \texttt{Swin} VT, which shows the highest testing performance without exception.
+Moreover, all VTs except \texttt{BeiT} are positioned relatively high in the rankings.
+This is also reflected in their combined average accuracies being higher than those of CNNs.
+To give \textit{Type classification} as an example: here VTs as a group achieve an 86.5\% accuracy on average, while this is 85.5\% for CNNs.
 
-It is interesting that \texttt{VGG19} performs worse than \texttt{ResNet50}, since it was the best performing OTS model on all 3 tasks in \citeauthor{sabatelli2018deep} (\citeyear{sabatelli2018deep}).
-The current study uses smaller datasets and fewer classes.
-The difference might be attributable to this, as \texttt{VGG19} is the largest model (see section \ref{methods:models}), and has the largest final layer -- with 4096 inputs.
+Finally, when comparing the 3 classification tasks to one another, it can be observed that highest overall accuracies are achieved for \textit{Artist classification}, and lowest ones for \textit{Material classification}.
+For balanced accuracy these differences are more extreme, as this measure falls behind plain accuracy the most for \textit{Material classification}, and the least for \textit{Artist classification}.
 
-When comparing tasks to one another, it can be observed that overall accuracies are lowest for material classification, and highest for artist classification.
-In addition, the difference between balanced and plain accuracy is the largest for material classification (up to 14\% on the testing set), while it is the smallest for artist classification (at most 4.7\%).
-Note that Gini-coefficients in table \ref{methods:datasets} show the same trend, with material classification being the least balanced, and artist classification the most.
+% Results favorable for VTs:
+%    - Best performing model = VT (SWIN)
+%    - 3 of 4 VTs on top ?
+%    - Avg of all VTs better than avg CNNs
 
-As a whole, VTs are doing well for the OTS experiments.
-Especially \texttt{Swin} and \texttt{DeiT} are ranked high on all testing sets, while \texttt{ConvNext} is the most promising CNN-based architecture.
-Taking the mean accuracy of all VT- and CNN-based models, reveals that VTs as a group perform slightly better.
-On type classification, for example, their mean accuracy is 86.5\%, while it is 85.5\% for CNNs.
+% Talk about differences between experiments
+% Talk about balanced accuracy?
 
-% This seems counterintuitive, as one would think artist identification is much more challenging than determining if something is a painting, or if it is made out of paper instead of wood.
-An explanation can likely be found in table \ref{methods:datasets}, which shows that material and artist classification sets are the least and most balanced, respectively.
-In addition, tables \ref{results:tab:ots_type} through \ref{results:tab:ots_artist} show that balanced accuracies are up to 14\% lower than plain ones for material classification, while they are much closer for artist classification.
-This suggests much accuracy is lost by misclassifying instances of less occurring classes in favor of more occurring ones.
-Similar finding can be done for FT results in section \ref{results:ft}.
-% Maybe switch this to discussion tho ^
-
-%OTS TL:
-%    Comparisons with Sabatelli paper (ResNet50 vs VGG19)
-%    VTs slightly better
-%    Lower balanced accuracy
-%    Higher balanced accuracy for type classification and even more for artist
-%    Artist higherst, which is surprising since it is the most difficult intuitively
-
-\subsection{Fine-tuning} \label{results:ft}
+\subsubsection{Fine-tuning} \label{results:ft}
 
@@
     \input{img/ft_mean_accuracy.pdf_tex}
-    \caption{Validation accuracy using fine-tuned models.}
+    \caption{Average validation accuracy when fine-tuning models on the target task.
+Overall, performance is higher compared to off the shelf learning in figure \ref{results:img:ots}, while the differences between models are also smaller here.
+The \texttt{Swin} VT and \texttt{ConvNext} CNN remain the best performing models, while in general, results are less favourable for VTs than they were for off the shelf learning.}
     \label{results:img:ft}
@@
 }
-\caption{Testing performance after fine-tuning.}
+\caption{Testing performance after fine-tuning.
+Results are again similar to validation accuracy in figure \ref{results:img:ft}.
+Compared to table \ref{results:tab:ots}, it is striking that the \texttt{ConvNext} CNN now often takes the lead with respect to balanced accuracy, and that the \texttt{ResNet} CNN often takes second place.}
 \label{results:tab:ft}
@@
 
-Once again, table \ref{results:tab:ft} shows testing performance, while figure \ref{results:img:ft} shows corresponding validation accuracies.
+Once again, table \ref{results:tab:ft} shows testing performance, while figure \ref{results:img:ft} shows validation accuracies.
 
-Similar to OTS TL, findings are somewhat consistent across classification tasks, while overall performance is again lowest for material classification, and highest for artist classification.
-\texttt{Swin} is still dominant in terms of testing accuracy, but Conv\-Next is catching up.
-This is especially true for balanced accuracy, where it tops the ranking on 2 of the 3 tasks.
-More general, it looks as if the advantage VTs seemed to have for OTS TL is gone.
-Mean accuracy of VTs grouped together is, for example, 91.2\% on type classification, while it is 91.3\% for CNNs.
+Results are less favourable for VTs than they were for OTS learning in section \ref{results:ots}, but at the same time, it can still be said that VTs perform on par with CNNs.
 
-Different from \citeauthor{matsoukas2021time} (\citeyear{matsoukas2021time}), \texttt{ResNet50} performs slightly better than \texttt{DeiT}.
-This remains true when using the same `tiny' version of \texttt{DeiT} as that paper used.
-On type classification, for example, it gets 90.44\% ($\pm$ 0.93\%) accuracy, whereas \texttt{ResNet50} shows 91.78\% in table \ref{results:tab:ft}.
+% Results less favourable compared to OTS
+%    - Swin still best, tho more debatable
+%    - ConvNext catching up
+%    - ResNet interesting
+%    - As groups: no differences anymore really
 
-FT results in this work also seem less favourable for VTs compared to \citeauthor{zhou2021convnets} (\citeyear{zhou2021convnets}).
-\texttt{ViT} consistently falls behind \texttt{ResNet50}, and even shows the worst accuracy in table \ref{results:tab:ft} (90.11\%), while in \citeauthor{zhou2021convnets} it was \texttt{ViT} that outperformed \texttt{ResNet101} and 152.
-It should be mentioned that these two alternatives show performance similar to \texttt{ResNet50}, with type classification accuracies of 91.96\% ($\pm$ 0.59\%) and 91.94\% ($\pm$ 0.25), respectively.
-Lastly, what is consistent with \citeauthor{zhou2021convnets}, is the good performance of \texttt{Swin} mentioned above.
+% Overall:
+%    - higher
+%    - closer together
+%    - for all but one worst FT is better than best OTS
+%    - same observations when comparing tasks to one another
 
-The performance difference between best and worst models is much smaller compared to OTS TL.
-There it was between 7.2\% and 11.8\%, while here it is at most 3.8\%.
-In addition, the worst FT model is often better than the best OTS model.
-The only exception is material classification, where the worst FT model (\texttt{BeiT}) has accuracy 85.74\%, and the best OTS one (\texttt{Swin}) has 85.87\%.
+\subsubsection{Scaling}
 
-\texttt{VGG19} has dropped further in the rankings, and now shows the worst balanced accuracy for type classification.
-\texttt{ResNet50}, on the other hand, appears to flourish with FT, often even beating Conv\-Next in the rankings.
-This is consistent with \citeauthor{sabatelli2018deep} (\citeyear{sabatelli2018deep}), where it was the best FT model overall.
+\subsection{Interpretation} \label{exp:int}
+% Mention that we skip scaling here because not needed
 
-\subsection{Scaling dataset size}
+\subsubsection{Off the shelf learning}
 
-\begin{figure*}[tb]
-    \centering
-    \def\svgwidth{\textwidth}
-    \input{img/scaling_accuracy.pdf_tex}
-    \caption{Testing accuracy as datasets gradually become smaller.}
-    \label{results:img:scale}
-\end{figure*}
+\subsubsection{Fine-tuning}
 
-%\begin{figure*}
-%    \centering
-%    \begin{subfigure}{0.45\textwidth}
-%    \def\svgwidth{7.7cm}
-%    \input{img/ots_scale_acc.pdf_tex}
-%    \caption{Off the shelf}
-%    \label{results:img:ots_scale}
-%    \end{subfigure}
-%    \hfill
-%    \begin{subfigure}{0.45\textwidth}
-%    \def\svgwidth{7.7cm}
-%    \input{img/ft_scale_acc.pdf_tex}
-%    \caption{Fine tuning}
-%    \label{results:img:ft_scale}
-%    \end{subfigure}
-%    \caption{Testing accuracy as datasets become smaller.}
-%    \label{results:img:scale}
-%\end{figure*}
-
-Figure \ref{results:img:scale} shows how testing accuracy decreases when smaller portions of the full dataset are taken (see section \ref{methods:dataset}).
-The x-axes show a logarithmic scale, with the rightmost value being roughly 10\% the size of the leftmost one.
-
-For both OTS TL as FT TL, findings done in sections \ref{results:ots} and \ref{results:ft} seem to hold up as dataset sizes are shrunk.
-When going from largest to smallest dataset, the mean accuracy of all models taken together drops with 9.1\% for OTS TL.
-For FT this is 9.2\%, which is very similar.
-
-\subsection{Off the shelf learning versus fine-tuning}
-
-\begin{figure}[tbh]
-    \centering
-    \def\svgwidth{7.7cm}
-    \input{img/ots_vs_ft_type.pdf_tex}
-    \caption{Time/accuracy trade-offs for VTs and CNNs either off the shelf or fine-tuned.}
-    \label{results:img:ots_vs_ft_type}
-\end{figure}
-
-Figure \ref{results:img:ots_vs_ft_type} shows type classification accuracy as reported in tables \ref{results:tab:ots} and \ref{results:tab:ft}.
-These are plotted against average epoch duration, as to get a sense of time/accuracy trade-offs.
-Colours make it easy to distinguish VTs from CNNs, and OTS models from FT ones.
-
-OTS VTs all take roughly the same time per epoch.
-CNNs are generally faster here, but accuracy also clearly falls behind for all but one (\texttt{ConvNext}).
-For FT, VT and CNN accuracies seem on-par, while CNNs still take much shorter to train.
-One CNN (\texttt{ResNet50}) even is faster than any OTS VT, while also showing better accuracy.
-In short, FT CNNs have better time/accuracy trade-offs, while it is a FT VT (\texttt{Swin}) that achieves the best accuracy overall.
-
-\subsection{Qualitative analysis}
-
-\begin{figure*}[tb]
-    \includegraphics[width=\textwidth]{img/layers.png}
-    \caption{Attention layers of successively deeper transformer blocks.}
-    \label{results:img:layers}
-\end{figure*}
-
-\begin{figure*}[tb]
-    \centering
-    \begin{subfigure}{0.49\textwidth}
-    \includegraphics[width=8cm]{img/img005_salience.png}
-    \caption{Dish}
-    \label{results:img:sal:dish}
-    \end{subfigure}
-    \begin{subfigure}{0.49\textwidth}
-    \includegraphics[width=8cm]{img/img011_salience.png}
-    \caption{Picture}
-    \label{results:img:sal:dish}
-    \end{subfigure}
-    \caption{Examples of saliency maps for different models.}
-\end{figure*}
+\subsubsection{General}

### 2022-07-11 writing/tex/draft_introduction.tex
Rewriting results section -> Now its an 'Experiments' section with 'Results' subsection. Some small tweaks as well

@@
 Since the introduction of AlexNet, roughly a decade ago, \textit{Convolutional Neural Networks} (CNNs) have played an important role in \textit{Computer Vision} (CV) \citep{krizhevsky2012imagened}.
-Recently, however, a new type of architecture, called \textit{Vision Transformer} (VT), gained state of the art performance on common learning benchmarks, including the ImageNet dataset\footnote{At the time of writing TODO is the best performing model, and the top TODO are all Transformer-based: (website)}.
+Recently, however, a new type of architecture, called \textit{Vision Transformer} (VT), gained state of the art performance on common learning benchmarks, including the ImageNet dataset\footnote{At the time of writing the VT \texttt{CoCa} is state of the art, according to \url{https://paperswithcode.com/sota/image-classification-on-imagenet}.}.
 

### 2022-07-11 writing/tex/draft_methods.tex
Rewriting results section -> Now its an 'Experiments' section with 'Results' subsection. Some small tweaks as well

@@
 The main focus was on a comparison between CNN architectures pre-trained on ImageNet1K when they were either used as OTS feature extractors, or fine-tuned to the new task.
-It was suggested that FT resulted in substantially better performance than either training from scratch or OTS TL.
+It was suggested that FT resulted in substantially better performance than either training from scratch or using an OTS learning scheme.
 
@@
 
-\begin{figure*}[tb]
+\begin{figure*}[bt]
 \includegraphics[width=\textwidth]{img/examples.png}
-\caption{Example images from the \textit{Rijksmuseum Challenge}.}
+\caption{Example images from the \textit{Rijksmuseum Challenge}. Each row depicts a different task.}
 \label{methods:eg}
@@
 
-Similar to \citeauthor{sabatelli2018deep}, used models are pre-trained on ImageNet1K.
-Moreover, the \textit{Rijksmuseum Challenge} dataset \citep{mensink14icmr} is used to extract the target tasks from.
+Similar to \citeauthor{sabatelli2018deep}, used models are pre-trained on ImageNet1K, while the \textit{Rijksmuseum Challenge} dataset \citep{mensink14icmr} is used for the target tasks.
 This dataset consists of a large collection of digitized artworks, together with xml-formatted metadata.
-The classification tasks distilled from it are: (1) \textit{Type classification}, containing labels as `painting', `sculpture', `drawing', etc.; (2) \textit{Material classification}, with labels as `paper', `porcelain', `silver', etc.; and finally, (3) \textit{Artist classification}, where the model has to predict who the creator is.
+The classification tasks extracted from it are: (1) \textit{Type classification}, containing labels as `painting', `sculpture', `drawing', etc.; (2) \textit{Material classification}, with labels as `paper', `porcelain', `silver', etc.; and finally, (3) \textit{Artist classification}, where the model has to predict who the creator is.
 Figure \ref{methods:eg} shows labelled examples for all three tasks.
@@
 The full dataset contains 112,039 images, and allows for multiple or zero labels per classification task.
-Different from \citeauthor{sabatelli2018deep}, only a fraction of the full set is used per task, as the research question concerns small datasets specifically.
 Samples are excluded from a task if they contain no labels for it.
-Samples containing more than one label are excluded as well, since analysis showed that in all cases this is true for only a small portion of the dataset.
+Samples containing more than one label are excluded as well, since analysis showed that in all cases this is true for only a small portion of the dataset.% Note that this last point is different from \citeauthor{sabatelli2018deep}, which took the first occurring class in these cases.
 
@@
 To counteract this, only the 30 most occurring classes are selected, and a cap of 1000 instances per class is enforced by taking a random sample when this limit is exceeded.
-Table \ref{methods:datasets} shows the result of these balancing operations.
+Table \ref{methods:datasets} shows the result of these balancing and subsampling operations.
 The situation prior is shown between brackets.
 
-For robustness, five subsets are randomly generated per classification task.
-These are then split up into 80\%, 10\%, 10\% training, validating and testing sets, such that all experiments can be run five times.
+It is worth noting that \citeauthor{sabatelli2018deep} does not perform these operations. They are performed here, because the current study is interested in small datasets specifically.
+
+All datasets are split up into 80\%, 10\%, 10\% training, validating and testing sets.
+For robustness, five subsets are randomly generated per classification task, such that each experiment can be run five times.
 The average sample overlap between two of the five datasets is shown in table \ref{methods:datasets} as well.
 
-Other than the tasks described above, table \ref{methods:datasets} shows an additional experiment called \textit{Scaling}.
+Besides the tasks described above, table \ref{methods:datasets} shows an experiment called \textit{Scaling}.
 The goal of this experiment is to see how the findings of this paper hold up as dataset sizes are gradually shrunk.
@@
 In addition, \citeauthor{matsoukas2021time} (\citeyear{matsoukas2021time}) used \texttt{ResNet50} too, while \citeauthor{zhou2021convnets} (\citeyear{zhou2021convnets}) used \texttt{ResNet101} and \texttt{ResNet152}.
-To also add some more recent CNNs, \texttt{ConvNext} \citep{liu2022convnet} -- a purely CNN-based model inspired by VTs' recent success -- and \texttt{EfficientNetV2} \citep{tan2021efficientnetv2} are included as well.
+To also add some more recent CNNs, \texttt{ConvNext} \citep{liu2022convnet} -- a purely CNN-based model inspired by VTs' recent success -- and \texttt{EfficientNetV2} \citep{tan2021efficientnetv2} are included.
 
@@
 The first model is the original VT \citep{dosovitskiy2020image}, which will be referred to as \texttt{ViT}.
-Next is \texttt{Swin} \citep{liu2021swin}, since it showed promising performance in \citeauthor{zhou2021convnets}.
+Next is \texttt{Swin} \citep{liu2021swin}, because it showed promising performance in \citeauthor{zhou2021convnets}.
 The final two are \texttt{DeiT} and \texttt{BeiT}.
@@
 An exception is made for the type classification experiment, as this one is also used to compare OTS TL and FT in terms of time/accuracy trade-offs.
-Preliminary findings suggested that using mixed precision does not affect performance for the aforementioned experiments.
+Preliminary findings suggested that using mixed precision does not affect performance for the aforementioned experiments, however.
 
-The \texttt{PyTorch} machine learning framework is used for all experiments, and many pre-trained models are taken from its \texttt{Torchvision} library.
+The \texttt{PyTorch} machine learning framework is used, and many pre-trained models are taken from its \texttt{Torchvision} library.
 Exceptions are \texttt{EfficientNetV2}, \texttt{Swin}, \texttt{DeiT} and \texttt{BeiT}, which are taken from the \texttt{Timm} library\footnote{\url{https://timm.fast.ai/}.}.
 
-Finally, all source code and data is made available on GitHub\footnote{\url{https://github.com/IndoorAdventurer/ViTTransferLearningForArtClassification}.}.
+Finally, all source code and data will be made available on GitHub\footnote{\url{https://github.com/IndoorAdventurer/ViTTransferLearningForArtClassification}.}.

### 2022-07-11 writing/tex/draft_results.tex
Rewriting results section -> Now its an 'Experiments' section with 'Results' subsection. Some small tweaks as well

@@
 Tables show average testing performance over the 5 trials per experiment, with standard deviation ($s$) as subscript.
-A green shaded cell marks best performance, while yellow and red depict second best and worst performance, respectively.
+A green shaded cell marks best performance, while yellow and red mark second best and worst performance, respectively.
 Finally, both accuracy and balanced accuracy are reported.
-The latter is defined as average recall over all classes.
-Note how this penalises errors on less occurring classes more than plain accuracy would.
+The latter of these is defined as average recall over all classes, such that it penalises errors on less occurring classes more than plain accuracy would.
 
-Line plots mostly show average validation accuracy (dashed line is CNN-based, continuous is VT-based), with \textit{standard error of the mean} ($\pm s \div \sqrt{N}$, where $N=5$) depicted as a shaded region.
+Line plots mostly show average validation accuracy, with \textit{standard error of the mean} ($\pm s \div \sqrt{N}$, where $N=5$) depicted as a shaded region.
 Plots end when an early stop occurred for the first of 5 trials.
+To easily distinguish VTs from CNNs, VTs are plotted with continues lines, and CNNs with dashed ones.
 
@@
 The current study uses smaller datasets and fewer classes.
-The difference might be attributable to this, as \texttt{VGG19} is the largest model (see section \ref{methods:models}), and has the largest final layer -- with 4096 inputs.
+The difference might be attributable to this, as \texttt{VGG19} is the largest model (see section \ref{methods:models}), and has the largest final layer (with 4096 inputs).
 
@@
 
-%\begin{figure*}
-%    \centering
-%    \begin{subfigure}{0.45\textwidth}
-%    \def\svgwidth{7.7cm}
-%    \input{img/ots_scale_acc.pdf_tex}
-%    \caption{Off the shelf}
-%    \label{results:img:ots_scale}
-%    \end{subfigure}
-%    \hfill
-%    \begin{subfigure}{0.45\textwidth}
-%    \def\svgwidth{7.7cm}
-%    \input{img/ft_scale_acc.pdf_tex}
-%    \caption{Fine tuning}
-%    \label{results:img:ft_scale}
-%    \end{subfigure}
-%    \caption{Testing accuracy as datasets become smaller.}
-%    \label{results:img:scale}
-%\end{figure*}
-
 Figure \ref{results:img:scale} shows how testing accuracy decreases when smaller portions of the full dataset are taken (see section \ref{methods:dataset}).
@@
     \includegraphics[width=\textwidth]{img/layers.png}
-    \caption{Attention layers of successively deeper transformer blocks.}
+    \caption{Attention layers of successively deeper transformer blocks.
+For \texttt{DeiT} attention with respect to the class-token is plotted.
+\texttt{Swin}, on the other hand, does not have a class-token, so average overall attention is shown.
+In addition, the window-shifting operations of this architecture are counteracted.}
     \label{results:img:layers}
 \end{figure*}
+
+\paragraph{Activation within VTs} To get more insights into how VTs build their representation of the input, figure \ref{results:img:layers} shows activation of attention layers for deeper and deeper blocks of \texttt{DeiT} and \texttt{Swin}.
+It can be observed that the first layers still retain much of a spatial layout.
+In fact, in the first blocks so much spatial information is kept that the input images can be recognised in the attention maps.
+Later layers show more sporadic activation, which would be expected from an architecture that creates a high-level encoding of the input in a hierarchical fashion.
 
@@
     \caption{Picture}
-    \label{results:img:sal:dish}
+    \label{results:img:sal:picture}
     \end{subfigure}
-    \caption{Examples of saliency maps for different models.}
+    \caption{Example saliency maps for different models either off the shelf (OTS) or fine-tuned (FT).
+GradCam was implemented for \texttt{ConvNext}, and Attention Rollout for \texttt{DeiT}.
+Finally, for \texttt{Swin} a modified version of Attention Rollout was implemented, which simply multiplies activation with gradients for one of the later layers, instead of multiplying average gradients per CNN channel with that channel's activation map.}
+    \label{results:img:sal}
 \end{figure*}
+% TODO find references for the saliency techniques!!!
+
+\paragraph{Saliency}
+Figure \ref{results:img:sal} shows saliency maps for the \texttt{ConvNext} CNN, and the \texttt{DeiT} and \texttt{Swin} VTs.
+It can be observed that for ConvNext representations differ more when comparing OTS TL to FT TL.
+In the case of figure \ref{results:img:sal:picture}, for example, the model learns to focus on the frame to classify the input as a picture after FT, while it was not able to do this after OTS TL.
+In the case of figure \ref{results:img:sal:dish}, something similar can be observed.
+The fine-tuned \texttt{ConvNext} learns to focus more on the center of the object, which seems to make the difference between a wrong and a correct classification.

### 2022-07-12 writing/tex/draft_experiments.tex
Finished the results section and made some progress with the Interpretation/Discussion section

@@
 These will first briefly be presented in section \ref{results}.
-After that, section \ref{exp:int} will discuss them in more detail, and see how they compare to previous works, etc.
+After that, section \ref{exp:int} will discuss them in more detail, and see how they compare to earlier studies, etc.
 
@@
 This is also reflected in their combined average accuracies being higher than those of CNNs.
-To give \textit{Type classification} as an example: here VTs as a group achieve an 86.5\% accuracy on average, while this is 85.5\% for CNNs.
+To give \textit{Type classification} as an example: here VTs as a group achieve an 86.5\% mean accuracy, while this is 85.5\% for CNNs.
 
 Finally, when comparing the 3 classification tasks to one another, it can be observed that highest overall accuracies are achieved for \textit{Artist classification}, and lowest ones for \textit{Material classification}.
-For balanced accuracy these differences are more extreme, as this measure falls behind plain accuracy the most for \textit{Material classification}, and the least for \textit{Artist classification}.
+For balanced accuracy the differences are more extreme, as this measure also falls behind plain accuracy the most for \textit{Material classification}, and the least for \textit{Artist classification}.
 
@@
 
-Results are less favourable for VTs than they were for OTS learning in section \ref{results:ots}, but at the same time, it can still be said that VTs perform on par with CNNs.
+Results are less favourable for VTs than they were for OTS learning in section \ref{results:ots}.
+At the same time, it can still be said that VTs perform on par with CNNs.
+The \texttt{Swin} VT shows the best overall testing accuracy here as well, but the \texttt{ConvNext} CNN is not far behind, and outperforms \texttt{Swin} in terms of balanced accuracy.
+Also noteworthy is the \texttt{ResNet} CNN, which is ranked higher here than it was for OTS learning, and now shows second best testing performance for \textit{Artist classification}.
+
+VTs and CNNs in general are much closer together.
+To again take \textit{Type classification} as an example: the mean testing accuracy of all VTs combined is 91.2\% on this task, while it is 91.3\% for CNNs.
+Recall that section \ref{results:ots} reported a difference of 1\% in favour of VTs here, with an accuracy of 86.5\% for VTs, and 85.5\% for CNNs.
+
+More general, it can be observed that FT leads to substantially better performance than OTS learning, and that models are now much closer together.
+The difference between highest and lowest testing accuracy, for example, is at most 3.8\% after FT, while it was between 7.2\% and 11.8\% after OTS learning.
+In addition, the worst testing accuracy after FT is most often still higher than the best accuracy after OTS learning.
+The only exception is \textit{Material classification}, where the worst FT model (\texttt{BeiT}) has accuracy 85.74\%, and the best OTS one (\texttt{Swin}) has 85.87\%.
+
+Lastly, note that highest overall performances are again achieved on the \textit{Artist classification} task, while the lowest ones are reported for \textit{Material classification}.
+The difference between plain and balanced accuracy is also again the largest for \textit{Material classification}, and lowest for \textit{Artist classification}.
 
@@
 
-\subsection{Interpretation} \label{exp:int}
-% Mention that we skip scaling here because not needed
+\begin{figure*}[tb]
+    \centering
+    \def\svgwidth{\textwidth}
+    \input{img/scaling_accuracy.pdf_tex}
+    \caption{Testing accuracy as datasets gradually become smaller.
+The x-axes show a logarithmic scale, with the smallest value being roughly 10\% the size of largest one.
+Observations done in sections \ref{results:ots} and \ref{results:ft} appear to hold up well as the training set is shrunk.}
+    \label{results:img:scale}
+\end{figure*}
+
+Figure \ref{results:img:scale} shows how testing accuracy decreases when smaller portions of the full dataset are taken (see section \ref{methods:dataset}).
+The x-axes show a logarithmic scale, where each successive value is roughly 56\% the size of its predecessor.
+
+For both OTS learning as FT, findings done in sections \ref{results:ots} and \ref{results:ft} seem to hold up as dataset sizes are shrunk.
+It is not true, for example, that at some point CNNs start to outperform VTs.
+In addition, \texttt{Swin} and \texttt{ConvNext} remain some of the best performing models throughout.
+
+Lastly, when going from largest to smallest dataset, the mean accuracy of all models taken together drops with 9.1\% for OTS learning.
+For FT this is 9.2\%, which is very similar.
+This, then, does not suggest that at some point OTS learning becomes preferred due to overfitting problems for FT.
+% TODO Maybe save this for Interpretation
+
+\subsection{Interpretation} \label{exp:int} % TODO maybe change to Discussion instead
+While most results have now been presented, it is still important to discuss them in more detail, as this leads to a more complete interpretation.
+This section attempts to provide such a discussion, by answering questions one might have at this moment.
+It will first discuss the aforementioned experiments individually, but will conclude with remarks that are more general.
+Topics will include: comparisons with earlier studies, potential shortcomings, and hypotheses that might explain certain observations.
 
@@
 
+\paragraph{How do these results compare to related studies?}
+Of all related work mentioned in section \ref{related_work}, \citeauthor{zhou2021convnets} (\citeyear{zhou2021convnets}) was the only one that examined VTs' OTS learning properties.
+It reported best OTS learning performance for the \texttt{Swin} VT, which is much in line with results in table \ref{results:tab:ots}.
+
+Besides \texttt{Swin}, \citeauthor{zhou2021convnets} also included \texttt{ViT} and \texttt{ResNet101} in its comparisons (among others), and consistently found that \texttt{ResNet101} outperformed \texttt{ViT}.
+The current study, on the other hand, uses \texttt{ResNet50}, which performs worse than \texttt{ViT} on all tasks except \textit{Artist classification}.
+Running the same experiments with \texttt{ResNet101} does, however, lead to similar findings as \citeauthor{zhou2021convnets}.
+On \textit{Type classification}, for example, it achieves a mean testing accuracy of  $86.21\% (\pm 0.70\%)$, which is indeed slightly higher than \texttt{ViT}'s 86.06\% shown in table \ref{results:tab:ots}.
+Mind that these differences are slight compared to the 89.43\% accuracy achieved by \texttt{Swin} here, so replacing \texttt{ResNet50} by \texttt{ResNet101} would not have affected the overall observations that much.
+
+Finally, because \texttt{VGG19} was specifically selected for its good OTS performance in \citeauthor{sabatelli2018deep} (\citeyear{sabatelli2018deep}), it is interesting that this model is outperformed by \texttt{ResNet50} (which that paper also included) in table \ref{results:tab:ots}.
+\texttt{VGG19} is the largest model of all (arguably too large), so this discrepancy might be attributable to the far smaller datasets used in the current study (see section \ref{methods:dataset}).
+Additionally, it might also be attributable to the smaller number of classes used here.
+\texttt{VGG19} has the largest final layer, containing 4096 inputs.
+In \citeauthor{sabatelli2018deep} this might have provided the expressive power needed to differentiate between a few hundred classes after just OTS learning, while here it would cause overfitting before good convergence is reached.
+
+%It stands out that \texttt{VGG19} is outperformed by \texttt{ResNet50} in table \ref{results:tab:ots}, as this model was specifically selected for its good OTS performance in \citeauthor{sabatelli2018deep}.
+
+%It is interesting that \texttt{VGG19} performs worse than \texttt{ResNet50}, since it was the best performing OTS model on all 3 tasks in \citeauthor{sabatelli2018deep} (\citeyear{sabatelli2018deep}).
+The current study uses smaller datasets and fewer classes.
+The difference might be attributable to this, as \texttt{VGG19} is the largest model (see section \ref{methods:models}), and has the largest final layer (with 4096 inputs).
+DIT KAN IK NU WEL EVEN WAT MEER UITLEGGEN NATUURLIJK!!!
+MORE EXPRESSIVE POWER FOR OTS WITH MORE CLASSES THERE
+
+\paragraph{Why do VTs show such good OTS performance?} ANSWER.
+% Also mention that this was unexpected (but in line with blabla);
+% Feature extractor/hierarchical representation building
+% Layers img.
+% Maybe mention that image specific inductive bias isn't a problem for swin?
+
 \subsubsection{Fine-tuning}
 
-\subsubsection{General}
+\paragraph{How do these results compare to related studies?} ANSWER.
+%Different from \citeauthor{matsoukas2021time} (\citeyear{matsoukas2021time}), \texttt{ResNet50} performs slightly better than \texttt{DeiT}.
+This remains true when using the same `tiny' version of \texttt{DeiT} as that paper used.
+On type classification, for example, it gets 90.44\% ($\pm$ 0.93\%) accuracy, whereas \texttt{ResNet50} shows 91.78\% in table \ref{results:tab:ft}.
+
+%FT results in this work also seem less favourable for VTs compared to \citeauthor{zhou2021convnets} (\citeyear{zhou2021convnets}).
+\texttt{ViT} consistently falls behind \texttt{ResNet50}, and even shows the worst accuracy in table \ref{results:tab:ft} (90.11\%), while in \citeauthor{zhou2021convnets} it was \texttt{ViT} that outperformed \texttt{ResNet101} and 152.
+It should be mentioned that these two alternatives show performance similar to \texttt{ResNet50}, with type classification accuracies of 91.96\% ($\pm$ 0.59\%) and 91.94\% ($\pm$ 0.25), respectively.
+Lastly, what is consistent with \citeauthor{zhou2021convnets}, is the good performance of \texttt{Swin} mentioned above.
+
+% Maybe better hyperparameters possible?
+
+%\texttt{VGG19} has dropped further in the rankings, and now shows the worst balanced accuracy for type classification.
+\texttt{ResNet50}, on the other hand, appears to flourish with FT, often even beating Conv\-Next in the rankings.
+This is consistent with \citeauthor{sabatelli2018deep} (\citeyear{sabatelli2018deep}), where it was the best FT model overall.
+
+\paragraph{Do CNNs gain more from FT than VTs?} ANSWER.
+% Show saliency maps here
+
+\subsubsection{Scaling}
+
+\paragraph{What can be concluded from these results?} ANSWER.
+% This, then, does not suggest that at some point OTS learning becomes preferred due to overfitting problems for FT.
+
+\paragraph{Are there any shortcomings?} ANSWER.
+% Same training/testing/validating split used. Was that smart? Defense: 5 runs so that makes up for too small testing sets
+
+\subsubsection{Overall}
+
+\paragraph{Why does \textit{Artist classification} show the best overall performance?} ANSWER.
+% Sabatelli different
+% GINI COEFFICIENTS HYPOTHESIS!!!:
+% This seems counterintuitive, as one would think artist identification is much more challenging than determining if something is a painting, or if it is made out of paper instead of wood.
+An explanation can likely be found in table \ref{methods:datasets}, which shows that material and artist classification sets are the least and most balanced, respectively.
+In addition, tables \ref{results:tab:ots_type} through \ref{results:tab:ots_artist} show that balanced accuracies are up to 14\% lower than plain ones for material classification, while they are much closer for artist classification.
+This suggests much accuracy is lost by misclassifying instances of less occurring classes in favor of more occurring ones.
+Similar finding can be done for FT results in section \ref{results:ft}.
+
+\paragraph{The same hyperparameters are used for all models. Is this fair?}
+% Preliminary studies
+% Not about getting the best overall performance, but seeing how they compare to one another
+% Asterix is that VTs might be able to do better, but then again -> Its also about ease of use
+
+\paragraph{To what extent are the chosen models representative for all VTs and CNNs?}
+
+\paragraph{What can be concluded from the results presented?}

### 2022-07-13 writing/tex/draft_conclusion.tex
Slowly getting there.. Still working on the 'discussion' section. Also started writing the conclusion.

\section{Conclusion}

Seeing which is better, VTs or CNNs, has never been the goal of this paper. The underlying architecture is not important, as one should always pick (or create) the model that fits the task best. What this paper, in fact, was interested in, is whether VTs can become a valuable alternative to CNNs under specific circumstances. So what, then, are these circumstances? Well, they can be described as those in which:
\begin{itemize}
\item small datasets are available: specifically, small enough to have been hand labelled by one person;
\item only consumer grade hardware is available.
\end{itemize}
In other words: this paper was interested in the type of machine learning one could try at home, and wanted to know if VTs can be put to good use there.

The way it tried to put them to use, is by utilising transfer learning. In the domain of art classification, it compared common CNNs and VTs to one another, for both off the shelf learning, where only the last layer of a pre-trained model is retrained on the new task, and fine-tuning, where the whole model is retrained.

The results showed that VTs are very well able to perform on par with CNNs for both off the shelf learning and fine-tuning. The most interesting architecture was the \texttt{Swin} VT, which gave best testing performance in almost all cases. Consequently, this model will be one of the firsts the author considers if he is ever faced with another computer vision task. Provided, of course, that the field does not catch up in ways that draw all models described here obsolete...

% TODO: more blabla

### 2022-07-13 writing/tex/draft_experiments.tex
Slowly getting there.. Still working on the 'discussion' section. Also started writing the conclusion.

@@
 \texttt{VGG19} is the largest model of all (arguably too large), so this discrepancy might be attributable to the far smaller datasets used in the current study (see section \ref{methods:dataset}).
-Additionally, it might also be attributable to the smaller number of classes used here.
-\texttt{VGG19} has the largest final layer, containing 4096 inputs.
-In \citeauthor{sabatelli2018deep} this might have provided the expressive power needed to differentiate between a few hundred classes after just OTS learning, while here it would cause overfitting before good convergence is reached.
+Additionally, it might also be attributable to the smaller number of classes used here: \texttt{VGG19} has the largest final layer, containing 4096 inputs.
+While in \citeauthor{sabatelli2018deep} this might have provided the expressive power needed to differentiate between a few hundred classes after just OTS learning, here it likely causes overfitting before good convergence is reached.
 
-%It stands out that \texttt{VGG19} is outperformed by \texttt{ResNet50} in table \ref{results:tab:ots}, as this model was specifically selected for its good OTS performance in \citeauthor{sabatelli2018deep}.
+\paragraph{Why do VTs show such good performance here?}
 
-%It is interesting that \texttt{VGG19} performs worse than \texttt{ResNet50}, since it was the best performing OTS model on all 3 tasks in \citeauthor{sabatelli2018deep} (\citeyear{sabatelli2018deep}).
-The current study uses smaller datasets and fewer classes.
-The difference might be attributable to this, as \texttt{VGG19} is the largest model (see section \ref{methods:models}), and has the largest final layer (with 4096 inputs).
-DIT KAN IK NU WEL EVEN WAT MEER UITLEGGEN NATUURLIJK!!!
-MORE EXPRESSIVE POWER FOR OTS WITH MORE CLASSES THERE
+\begin{figure*}[tb]
+    \includegraphics[width=\textwidth]{img/layers.png}
+    \caption{Attention layers of successively deeper transformer blocks.
+One can recognise the input image in earlier layers, while for later ones activation seems sporadic.
+For \texttt{DeiT}, attention with respect to the class-token is plotted.
+\texttt{Swin}, on the other hand, does not have a class-token, so average overall attention is used instead.
+In addition, the window-shifting operations of this architecture are counteracted.}
+    \label{results:img:layers}
+\end{figure*}
 
-\paragraph{Why do VTs show such good OTS performance?} ANSWER.
+While results are in line with previous studies, it was still surprising to see that VTs showed such good OTS learning performance in section \ref{results:ots}.
+It suggests that these transformer based architectures are able to give a very complete, high level encoding of the input image.
+For CNNs this is commonly thought to be true, as they are often described as hierarchical feature extractors that build successively higher level feature maps from lower level ones (see section \ref{intro:cnns}).
+If the same is true for VTs, is not well understood at this moment.
+Or at least to the best of the author's knowledge, that is.
+
+A hint of an answer can, however, be found when plotting activation of successively deeper attention layers, as is done in figure \ref{results:img:layers}.
+While earlier layers retain much spatial information, later ones show more sporadic activation patters.
+This is what one might expect from an hierarchical feature extractor: the random pixels lighting up might then, for example, correspond to specific, high level properties of the input.
+Nevertheless, it should be emphasised that no strong conclusions should be drawn from these images, and that more research is required to understand how VTs function.
 % Also mention that this was unexpected (but in line with blabla);
@@
 % Layers img.
-% Maybe mention that image specific inductive bias isn't a problem for swin?
+% Maybe mention that image specific inductive bias isn't a problem for swin? TODO somewhere else!
 
@@
 
-\paragraph{How do these results compare to related studies?} ANSWER.
-%Different from \citeauthor{matsoukas2021time} (\citeyear{matsoukas2021time}), \texttt{ResNet50} performs slightly better than \texttt{DeiT}.
-This remains true when using the same `tiny' version of \texttt{DeiT} as that paper used.
-On type classification, for example, it gets 90.44\% ($\pm$ 0.93\%) accuracy, whereas \texttt{ResNet50} shows 91.78\% in table \ref{results:tab:ft}.
+\paragraph{How do these results compare to related studies?}
+That the \texttt{Swin} VT performs well after FT, is again consistent with \citeauthor{zhou2021convnets} (\citeyear{zhou2021convnets}).
+That paper, however, also shows \texttt{ViT} outperforming the \texttt{ResNet101} and \texttt{ResNet152} CNNs.
+This differs from the results in table \ref{results:tab:ft}, where \texttt{ViT} does not perform well overall, and with 90.11\% even shows the worst testing accuracy on \textit{Type classification}.
+\texttt{ResNet50} achieves 91.78\% here, and testing the same \texttt{ResNet} versions \citeauthor{zhou2021convnets} used, gives 91.96\% ($\pm$ 0.59\%) accuracy for \texttt{ResNet101}, and 91.94\% ($\pm$ 0.25) for \texttt{ResNet152}.
+
+Results are also less favourable for VTs compared to \citeauthor{matsoukas2021time} (\citeyear{matsoukas2021time}).
+In that paper the `tiny' version of \texttt{DeiT} outperforms \texttt{ResNet50}, while in table \ref{results:tab:ft} this same \texttt{ResNet50} performs slightly better than the `base' version of \texttt{DeiT}.
+Going `tiny' also does not help: on \textit{Type classificaiton}, for instance, `tiny' \texttt{DeiT} gets 90.44\% ($\pm$ 0.93\%), which is even worse than `base' \texttt{DeiT}'s 91.78\% in table \ref{results:tab:ots}.
+
+These slight performance differences could be attributable to the different datasets used.
+In that case, it seems, VTs have certain strengths that those other datasets take more advantage of.
+Alternatively, it could also be that hyperparameters and regularisation could still be improved.
+The reason for thinking this, is that preliminary experiments suggested that VTs benefit more from regularisation than CNNs do.
+
+Finally, it was observed that \texttt{ResNet50} appears to flourish with FT, sometimes even beating Conv\-Next in the rankings.
+This is consistent with \citeauthor{sabatelli2018deep} (\citeyear{sabatelli2018deep}), where it was the best FT model overall.
 
@@
 
-\paragraph{Do CNNs gain more from FT than VTs?} ANSWER.
+\paragraph{Do CNNs gain more from FT than VTs?}
 % Show saliency maps here
@@
 
-\paragraph{What can be concluded from these results?} ANSWER.
+\paragraph{What can be concluded from these results?}
 % This, then, does not suggest that at some point OTS learning becomes preferred due to overfitting problems for FT.
 
-\paragraph{Are there any shortcomings?} ANSWER.
+\paragraph{Are there any shortcomings?}
 % Same training/testing/validating split used. Was that smart? Defense: 5 runs so that makes up for too small testing sets
@@
 
-\paragraph{Why does \textit{Artist classification} show the best overall performance?} ANSWER.
+\paragraph{Why does \textit{Artist classification} show the best overall performance?}
 % Sabatelli different

### 2022-07-13 writing/tex/draft_introduction.tex
Slowly getting there.. Still working on the 'discussion' section. Also started writing the conclusion.

@@
 
-\subsection{Convolutional Neural Networks}
+\subsection{Convolutional Neural Networks} \label{intro:cnns}
 The modern CNN architecture is often attributed to \citeauthor{lecun1998gradient} (\citeyear{lecun1998gradient}), which describes how the number of trainable weights per layer can be reduced in manners that are sensible for an image's topology (i.e.

### 2022-07-18 writing/tex/draft_conclusion.tex
Making good progress with writing. Just an intermediate push

@@
 
+% TODO dit moet echt allemaal nog een stuk beter!
+
 Seeing which is better, VTs or CNNs, has never been the goal of this paper.
-The underlying architecture is not important, as one should always pick (or create) the model that fits the task best.
-What this paper, in fact, was interested in, is whether VTs can become a valuable alternative to CNNs under specific circumstances.
-So what, then, are these circumstances?
-Well, they can be described as those in which:
+The underlying architecture is somewhat irrelevant, as one should always pick (or create) the model that fits the task best.
+What this paper, then, was interested in, is whether VTs can become a valuable alternative to CNNs under circumstances in which:
 \begin{itemize}
-\item small datasets are available: specifically, small enough to have been hand labelled by one person;
+\item small datasets are used: specifically, small enough to have been hand labelled by one person;
 \item only consumer grade hardware is available.

### 2022-07-18 writing/tex/draft_experiments.tex
Making good progress with writing. Just an intermediate push

@@
 This, then, does not suggest that at some point OTS learning becomes preferred due to overfitting problems for FT.
-% TODO Maybe save this for Interpretation
 
@@
 This section attempts to provide such a discussion, by answering questions one might have at this moment.
-It will first discuss the aforementioned experiments individually, but will conclude with remarks that are more general.
+It will first discuss some aforementioned experiments individually, but will conclude with remarks that are more general.
 Topics will include: comparisons with earlier studies, potential shortcomings, and hypotheses that might explain certain observations.
@@
 Finally, because \texttt{VGG19} was specifically selected for its good OTS performance in \citeauthor{sabatelli2018deep} (\citeyear{sabatelli2018deep}), it is interesting that this model is outperformed by \texttt{ResNet50} (which that paper also included) in table \ref{results:tab:ots}.
-\texttt{VGG19} is the largest model of all (arguably too large), so this discrepancy might be attributable to the far smaller datasets used in the current study (see section \ref{methods:dataset}).
+\texttt{VGG19} is the largest model of all, so this discrepancy could be an indication that the smaller datasets used in the current study are too small for \texttt{VGG19}.
 Additionally, it might also be attributable to the smaller number of classes used here: \texttt{VGG19} has the largest final layer, containing 4096 inputs.
@@
 
-\paragraph{Why do VTs show such good performance here?}
+\paragraph{Why do VTs perform so well here?}
 
@@
 While earlier layers retain much spatial information, later ones show more sporadic activation patters.
-This is what one might expect from an hierarchical feature extractor: the random pixels lighting up might then, for example, correspond to specific, high level properties of the input.
-Nevertheless, it should be emphasised that no strong conclusions should be drawn from these images, and that more research is required to understand how VTs function.
+This is what one might expect from an hierarchical feature extractor: the random pixels lighting up might, for example, correspond to specific, high level properties of the input.
+Nevertheless, it should be emphasised that no strong conclusions should be drawn from these images, and that more research with respect to explainability is desired.
 % Also mention that this was unexpected (but in line with blabla);
@@
 
-\subsubsection{Fine-tuning}
+\subsubsection{Fine-tuning} \label{exp:int:ft}
 
@@
 That the \texttt{Swin} VT performs well after FT, is again consistent with \citeauthor{zhou2021convnets} (\citeyear{zhou2021convnets}).
-That paper, however, also shows \texttt{ViT} outperforming the \texttt{ResNet101} and \texttt{ResNet152} CNNs.
-This differs from the results in table \ref{results:tab:ft}, where \texttt{ViT} does not perform well overall, and with 90.11\% even shows the worst testing accuracy on \textit{Type classification}.
-\texttt{ResNet50} achieves 91.78\% here, and testing the same \texttt{ResNet} versions \citeauthor{zhou2021convnets} used, gives 91.96\% ($\pm$ 0.59\%) accuracy for \texttt{ResNet101}, and 91.94\% ($\pm$ 0.25) for \texttt{ResNet152}.
+In that paper, however, \texttt{ViT} outperforms the \texttt{ResNet101} and \texttt{ResNet152} CNNs.
+This differs from the results in table \ref{results:tab:ft}, where \texttt{ViT} does not perform well overall, and even shows the worst testing accuracy on \textit{Type classification}, with 90.11\%.
+\texttt{ResNet50} achieves 91.78\% accuracy here.
+Testing the same \texttt{ResNet101} and \texttt{ResNet152} versions that \citeauthor{zhou2021convnets} used does not change much, as this leads to 91.96\% ($\pm$ 0.59\%) and 91.94\% ($\pm$ 0.25) accuracy, respectively.
 
@@
 In that case, it seems, VTs have certain strengths that those other datasets take more advantage of.
-Alternatively, it could also be that hyperparameters and regularisation could still be improved.
+Alternatively, it could also be that hyperparameters and regularisation still leave room for improvement.
 The reason for thinking this, is that preliminary experiments suggested that VTs benefit more from regularisation than CNNs do.
 
-Finally, it was observed that \texttt{ResNet50} appears to flourish with FT, sometimes even beating Conv\-Next in the rankings.
+Finally, it was observed that \texttt{ResNet50} appears to flourish with FT, sometimes even beating \texttt{Conv\-Next} in the rankings.
 This is consistent with \citeauthor{sabatelli2018deep} (\citeyear{sabatelli2018deep}), where it was the best FT model overall.
 
-%FT results in this work also seem less favourable for VTs compared to \citeauthor{zhou2021convnets} (\citeyear{zhou2021convnets}).
-\texttt{ViT} consistently falls behind \texttt{ResNet50}, and even shows the worst accuracy in table \ref{results:tab:ft} (90.11\%), while in \citeauthor{zhou2021convnets} it was \texttt{ViT} that outperformed \texttt{ResNet101} and 152.
-It should be mentioned that these two alternatives show performance similar to \texttt{ResNet50}, with type classification accuracies of 91.96\% ($\pm$ 0.59\%) and 91.94\% ($\pm$ 0.25), respectively.
-Lastly, what is consistent with \citeauthor{zhou2021convnets}, is the good performance of \texttt{Swin} mentioned above.
+\paragraph{Why are results less favourable for VTs after FT than after OTS learning?}
 
-% Maybe better hyperparameters possible?
+\begin{figure*}[tb]
+    \centering
+    \begin{subfigure}{0.49\textwidth}
+    \includegraphics[width=8cm]{img/img011_salience.png}
+    \caption{Picture}
+    \label{results:img:sal:picture}
+    \end{subfigure}
+    \begin{subfigure}{0.49\textwidth}
+    \includegraphics[width=8cm]{img/img005_salience.png}
+    \caption{Dish}
+    \label{results:img:sal:dish}
+    \end{subfigure}
+    \caption{Saliency maps for \texttt{ConvNext}, \texttt{DeiT} and \texttt{Swin} after both off the shelf learning (OTS), and fine-tuning (FT).
+\texttt{ConvNext} appears to benefit more when training the whole model instead of the last layer, because it shows the biggest differences between the OTS learning and FT.
+In case of image \ref{results:img:sal:dish} this even makes the difference between correct and incorrect classification.}
+    \label{results:img:sal}
+\end{figure*}
 
-%\texttt{VGG19} has dropped further in the rankings, and now shows the worst balanced accuracy for type classification.
-\texttt{ResNet50}, on the other hand, appears to flourish with FT, often even beating Conv\-Next in the rankings.
-This is consistent with \citeauthor{sabatelli2018deep} (\citeyear{sabatelli2018deep}), where it was the best FT model overall.
+The FT results in table \ref{results:tab:ft} are all much closer together than the OTS results in table \ref{results:tab:ots}.
+It is as if OTS learning already unlocks much of VTs' potential, while CNNs need an FT learning scheme for this.
 
-\paragraph{Do CNNs gain more from FT than VTs?}
+Supporting `evidence' might be visible in figure \ref{results:img:sal}.
+It plots saliency maps for one CNN (\texttt{ConvNext}), and two VTs (\texttt{DeiT} and \texttt{Swin}).
+It is interesting that \texttt{ConvNext} shows much bigger differences between the OTS and FT plots.
+In figure \ref{results:img:sal:picture}, for example, it learns to just look at the picture frame after FT.
+In figure \ref{results:img:sal:dish} it learns to look more at the center, which allows it to correctly classify the input as a dish instead of a plate\footnote{No idea what the difference is between these classes, but these models do seem to know, so that is good!}.
+% TODO Maybe delete this comment, not very professional
+These bigger differences for \texttt{ConvNext} align well with observations made earlier, as they suggest that there is still much improvement possible after just OTS learning for CNNs.
+
+The images in figure \ref{results:img:sal} are representative examples, and show observation that could have been made from other images too.
+How representative the 3 models are for all 8 is debatable, and was not tested.
+In general, a big disclaimer should be added to these results, because some hand waving is involved.
+Importantly, 3 different salience mapping techniques were used: GradCam \citep{selvaraju2017grad} was implemented for \texttt{ConvNext}, while Attention Rollout \citep{abnar2020quantifying} was implemented for \texttt{DeiT}.
+This already makes the comparison somewhat unfair.
+This becomes even worse when adding \texttt{Swin}, for which a self-invented method was used.
+This method is similar to GradCam, but instead of multiplying average CNN channel gradients with said channel's activation, it directly multiplies activation with corresponding gradients in the output of one of the later transformer blocks\footnote{For the full implementation of all techniques, one is referred to the source code on: \\ \url{https://github.com/IndoorAdventurer/ViTTransferLearningForArtClassification}.}.
+Despite this, saliency maps were included in this thesis, as they are interesting nevertheless.
 % Show saliency maps here
-
-\subsubsection{Scaling}
-
-\paragraph{What can be concluded from these results?}
-% This, then, does not suggest that at some point OTS learning becomes preferred due to overfitting problems for FT.
-
-\paragraph{Are there any shortcomings?}
-% Same training/testing/validating split used. Was that smart? Defense: 5 runs so that makes up for too small testing sets
+% MENTION TECHNIQUES AND SHORTCOMMINGS!!!
 
@@
 \paragraph{Why does \textit{Artist classification} show the best overall performance?}
+This might seem counter-intuitive: one would think it is much easier to determine if something is a painting or a sculpture, than who the creator is.
+However, because only the top 30 classes are used, this is not necessarily true.
+
+Figure \ref{methods:eg} shows example images for all tasks.
+On the one hand, it shows much heterogeneity for the \textit{Artist classification} task, since authors work in very different styles and mediums.
+On the other hand, it also shows difficulties for the other tasks: \textit{Material classification}, which seems to be the most challenging task, shows examples of both big differences within classes (note how the two instances of porcelain are very different), and small differences between classes (note how the porcelain plate is very similar to the faience plate).
+
+Then there is the problem of balance.
+It was mentioned earlier that balanced accuracy falls behind plain accuracy the most for \textit{Material classification}, and the least for \textit{Artist classification}.
+At the same time, table \ref{methods:datasets} shows that it is also \textit{Material classification} which is the least balanced (gini-coefficient of 0.563), while it is \textit{Artist classification} that is the most balanced (coefficient 0.236).
+This suggests that balance is hurting the performance measures, so also plain accuracy, the least for \textit{Artist classification}.
+
 % Sabatelli different
@@
 \paragraph{The same hyperparameters are used for all models. Is this fair?}
+The final learning scheme was chosen after extensive hyperparameter optimization, and was partially inspired by related work.
+It is possible that doing this for each model individually, would have resulted in \textit{slightly} higher performances.
+This paper, however, does not aim to get the best possible performance, but instead wants to see how models compare to one another.
+It is unlikely that individually tuning each model would have changed overall findings.
+
+A small asterisk to be made here, is that VTs might benefit more from better regularisation than CNNs (see section \ref{exp:int:ft}).
+Still, final accuracy should not be the only concern, because ease of use will also determine to what degree VTs are a valuable alternative to CNNs.
+Moreover, the main finding of this paper seems to be that VTs are able to perform on par with CNNs.
+Slightly better performance for VTs would not exactly hurt this finding.
+
 % Preliminary studies
@@
 \paragraph{To what extent are the chosen models representative for all VTs and CNNs?}
+Most models were selected because of their mention in related work.
+Notable exceptions are \texttt{ConvNext} and \texttt{EfficientNetV2}, which were selected because they are equally recent as the used VTs.
+Many of these models are popular/common, but one can't say they are a representative sample.
+The question, of course, is also not how VTs and CNNs as a group compare to one another.
+Instead, the question is if the VT landscape as a whole has something to offer that could be an interesting alternative to common CNNs.
+
+More interesting is the question if, for example, `tiny' or `medium' sized version should have been used instead of `base' versions.
+On \textit{Type classification} `tiny' models were tested as well, and results were reported where appropriate.
+Overall, however, these results did not suggest that using `tiny' versions everywhere have would made much of a difference.
 
 \paragraph{What can be concluded from the results presented?}
+
+\begin{figure}[tbh]
+    \centering
+    \def\svgwidth{7.7cm}
+    \input{img/ots_vs_ft_type.pdf_tex}
+    \caption{Time/accuracy trade-offs for VTs and CNNs either off-the-shelf or fine-tuned.
+Fine-tuned CNNs seem to make the best trade-off here, with one even appearing left above all off-the-shelf VTs.
+The best overall performance is achieved by a fine-tuned VT.}
+    \label{results:img:ots_vs_ft_type}
+\end{figure}

### 2022-07-19 writing/tex/draft_conclusion.tex
Concusion concluded.

@@
 \section{Conclusion}
-
-% TODO dit moet echt allemaal nog een stuk beter!
-
 Seeing which is better, VTs or CNNs, has never been the goal of this paper.
-The underlying architecture is somewhat irrelevant, as one should always pick (or create) the model that fits the task best.
+The underlying architecture is somewhat irrelevant, as one should always pick (or create) the model that fits the task best -- regardless of whether it is a VT, CNN, or something else.
 What this paper, then, was interested in, is whether VTs can become a valuable alternative to CNNs under circumstances in which:
@@
 \end{itemize}
-In other words: this paper was interested in the type of machine learning one could try at home, and wanted to know if VTs can be put to good use there.
+In other words: this paper was interested in the type of machine learning one could try at home, and wanted to know if VTs could be made useful for it.
 
-The way it tried to put them to use, is by utilising transfer learning.
-In the domain of art classification, it compared common CNNs and VTs to one another, for both off the shelf learning, where only the last layer of a pre-trained model is retrained on the new task, and fine-tuning, where the whole model is retrained.
+To that end, it tried to utilise transfer learning methods, as these might help overcome problems that arise from VTs' lack of an inductive bias.
+More concretely, models pre-trained on ImageNet1K were either used as off the shelf feature extractors, or fine-tuned to new tasks.
+These tasks were all taken from the \textit{Rijksmuseum Challenge} dataset \citep{mensink14icmr}, and fall within the domain of art classification.
 
-The results showed that VTs are very well able to perform on par with CNNs for both off the shelf learning and fine-tuning.
-The most interesting architecture was the \texttt{Swin} VT, which gave best testing performance in almost all cases.
-Consequently, this model will be one of the firsts the author considers if he is ever faced with another computer vision task.
-Provided, of course, that the field does not catch up in ways that draw all models described here obsolete...
+The results demonstrated that VTs are very well able to perform on par with CNNs here.
+Especially the \texttt{Swin} VT showed promising results, and might be worth considering for future learning problems.
 
-% TODO: more blabla
+\texttt{ConvNext} was a close second in the rankings.
+It is interesting that this best performing CNN was inspired by VTs, while the best performing VT \texttt{Swin} took inspiration from CNNs.
+It shows the kind of cross-fertilization that moves the field forward, but also hints at something bigger that would be nice to conclude with.
+
+No matter what happens.
+No matter if VTs, CNNs, a hybrid architecture, or even something completely different rains supreme in the coming years; it is great that VTs came about.
+It shakes things up, and forces researchers to look at problems from a different perspective.
+Perhaps that, then, will be VTs' greatest contribution of all.

### 2022-07-19 writing/tex/draft_experiments.tex
Concusion concluded.

@@
 
-VTs and CNNs in general are much closer together.
+VTs and CNNs grouped together are much more evenly matched.
 To again take \textit{Type classification} as an example: the mean testing accuracy of all VTs combined is 91.2\% on this task, while it is 91.3\% for CNNs.
@@
 
-\subsection{Interpretation} \label{exp:int} % TODO maybe change to Discussion instead
+\subsection{Discussion} \label{exp:int} % TODO maybe change to Interpretation instead
 While most results have now been presented, it is still important to discuss them in more detail, as this leads to a more complete interpretation.
 This section attempts to provide such a discussion, by answering questions one might have at this moment.
-It will first discuss some aforementioned experiments individually, but will conclude with remarks that are more general.
+It will first address some of the aforementioned experiments individually, but will conclude with remarks that are more general.
 Topics will include: comparisons with earlier studies, potential shortcomings, and hypotheses that might explain certain observations.
@@
 \paragraph{Why does \textit{Artist classification} show the best overall performance?}
-This might seem counter-intuitive: one would think it is much easier to determine if something is a painting or a sculpture, than who the creator is.
+This might seem counter-intuitive: one would think that it is much easier to determine if something is a painting or a sculpture, than determining who the creator is.
 However, because only the top 30 classes are used, this is not necessarily true.
@@
 Figure \ref{methods:eg} shows example images for all tasks.
-On the one hand, it shows much heterogeneity for the \textit{Artist classification} task, since authors work in very different styles and mediums.
+On the one hand, it shows much heterogeneity for the \textit{Artist classification} task, with artists working in very different styles and mediums.
 On the other hand, it also shows difficulties for the other tasks: \textit{Material classification}, which seems to be the most challenging task, shows examples of both big differences within classes (note how the two instances of porcelain are very different), and small differences between classes (note how the porcelain plate is very similar to the faience plate).
@@
 At the same time, table \ref{methods:datasets} shows that it is also \textit{Material classification} which is the least balanced (gini-coefficient of 0.563), while it is \textit{Artist classification} that is the most balanced (coefficient 0.236).
-This suggests that balance is hurting the performance measures, so also plain accuracy, the least for \textit{Artist classification}.
+This suggests that balance is hurting performance the least for \textit{Artist classification} -- and the most for \textit{Material classification}.
 
@@
 A small asterisk to be made here, is that VTs might benefit more from better regularisation than CNNs (see section \ref{exp:int:ft}).
-Still, final accuracy should not be the only concern, because ease of use will also determine to what degree VTs are a valuable alternative to CNNs.
+Still, final accuracy should not be the only concern, because ease of use (i.e.
+effort needed to find the right hyperparameters) will also determine to what degree VTs are a valuable alternative to CNNs.
 Moreover, the main finding of this paper seems to be that VTs are able to perform on par with CNNs.
@@
 
-More interesting is the question if, for example, `tiny' or `medium' sized version should have been used instead of `base' versions.
-On \textit{Type classification} `tiny' models were tested as well, and results were reported where appropriate.
-Overall, however, these results did not suggest that using `tiny' versions everywhere have would made much of a difference.
+It is also interesting to wonder if, for example, `tiny' or `medium' sized version should have been used instead of `base' versions.
+All available `tiny' models were tested as well, and results have been reported where appropriate (e.g.
+section \ref{exp:int:ft}).
+Overall, however, these results did not suggest that using `tiny' versions would have made much of a difference for the rankings: `tiny' \texttt{Swin} and \texttt{ConvNext}, for example, remained superior under that circumstance.
 
@@
 \end{figure}
+
+After OTS learning, results are slightly better for VTs, while after FT, results are almost the same for CNNs and VTs.
+With that said, it seems, VTs are very much able to perform on par with CNNs on the described art classification problems.
+Especially \texttt{Swin} and \texttt{DeiT} showed promising performance, and together with \texttt{ConvNext}, might be interesting models to consider.
+
+Of course, final accuracy is not the full story, since time/accuracy trade-offs, for example, might also determine which model is preferred.
+Figure \ref{results:img:ots_vs_ft_type} plots \textit{Type classification} accuracies of tables \ref{results:tab:ots} and \ref{results:tab:ft} against training time per epoch.
+Total training time was not chosen instead, as this measure depends much more on selected hyperparameters.
+
+The best overall performing model is a fine-tuned VT (\texttt{Swin}).
+At the same time, it is a CNN that shows the best time/accuracy trade-offs, because it is able to perform better than any OTS trained model, while still taking less time per epoch than any OTS trained VT.
+The particular CNN this regards, is \texttt{ResNet50}.
+Recall that this model only had 25.6 million parameters, while most VTs had around 86 million.
+
+All `tiny' VTs are around the same size as \texttt{ResNet50}, and using these would have changed the situation here completely.
+As such, taking figure \ref{results:img:ots_vs_ft_type} with a grain of salt, would be appropriate.
+
+Still, it seemed fitting to conclude with this figure, as it neatly summarises all the main findings done in this section.

### 2022-07-19 writing/tex/draft_introduction.tex
Concusion concluded.

@@
 
-Another ``shortcomming'' of \citeauthor{matsoukas2021time} is that it merely examined whether VTs can use TL to outperform CNNs in a domain that happens to be characterized by limited-sized datasets, but that it did not actively focus on the extent to which VTs can be made useful in limited data/compute environments.
+Another `shortcomming' of \citeauthor{matsoukas2021time} is that it merely examined whether VTs can use TL to outperform CNNs in a domain that happens to be characterized by limited-sized datasets, but that it did not actively focus on the extent to which VTs can be made useful in limited data/compute environments.
 More related to this is \citeauthor{touvron2021training} (\citeyear{touvron2021training}), which proposed transformer-specific distillation-based learning techniques to make VTs applicable for smaller datasets.

### 2022-07-20 writing/report.tex
working on the presentation
Continued working on writing y paper
Rewriting results section -> Now its an 'Experiments' section with 'Results' subsection. Some small tweaks as well
Finished the results section and made some progress with the Interpretation/Discussion section
Slowly getting there.. Still working on the 'discussion' section. Also started writing the conclusion.
Concusion concluded.
Finished the abstract earlier today. With that, the first draft is done!

@@
 
+%\widowpenalties 1 10000
+\raggedbottom
+
 \title{\vspace{-0.5in}{\bfseries\scshape 
-    Transfer Learning for computationally limited training of Vision Transformers on art classification tasks
+    Transferability of Vision Transformers: The key to success on small art classification datasets?%    Transfer Learning for computationally limited training of Vision Transformers on art classification tasks
     } \\ 
@@
     \normalsize{V. Tonkes, s3617157, v.tonkes@student.rug.nl,} \\ 
-    \normalsize{Supervisor: Dr. M. Sabatelli}
+    \normalsize{Supervisor: M. Sabatelli, PhD}
 }
@@
   \begin{@twocolumnfalse}
-%   \input{./tex/abstract.tex}
+   \input{tex/draft_abstract.tex}
   \end{@twocolumnfalse}
@@
 \input{tex/draft_methods.tex}
-\input{tex/draft_results.tex}
+\input{tex/draft_experiments.tex}
+\input{tex/draft_conclusion.tex}
+
+\section*{Acknowledgements}
+\textit{The author would like to thank {the Center for Information Technology of the University of Groningen} for their support and for providing access to the Peregrine high performance computing cluster.
+He would also like to thank {dr.
+M.
+Sabatelli} for his guidance during this project.}
+
+\clearpage
 
@@
 
-\clearpage
+%\clearpage
 

### 2022-07-20 writing/tex/draft_abstract.tex
Finished the abstract earlier today. With that, the first draft is done!

\begin{abstract} {Convolutional Neural Networks (CNNs) have become the de facto standard in Computer Vision, and played a major role in the advancement of this field. In recent years, however, Vision Transformer based models (VTs) have been outperforming CNNs across the board. While this is exciting, CNNs still have an advantage on small datasets, due to their so-called image specific inductive bias. Recent work suggests that transfer learning methods allow VTs to compete with CNNs on some of these small datasets. This thesis investigates whether that also holds true for art classification problems within the digital humanities. To this end, it compares popular VTs and CNNs in terms of their off-the-shelf and fine-tuning transferability, when going from \textit{ImageNet1K} to target tasks provided by the \textit{Rijksmuseum Challenge} dataset. The results show that VTs posses superior off-the-shelf feature extraction capabilities here, and that in general, transfer learning allows VTs to become an interesting alternative to consider within the digital humanities.\vspace{8ex}}
\end{abstract}

### 2022-07-29 writing/tex/draft_conclusion.tex
Finished the abstract earlier today. With that, the first draft is done!
Made all the changes that I wrote down on the printed out draft. Up next: throwing it through Docs to check grammar, and having it read by one or two friends!

@@
 To that end, it tried to utilise transfer learning methods, as these might help overcome problems that arise from VTs' lack of an inductive bias.
-More concretely, models pre-trained on ImageNet1K were either used as off the shelf feature extractors, or fine-tuned to new tasks.
+More concretely, models pre-trained on ImageNet1K were either used as off-the-shelf feature extractors, or fine-tuned to new tasks.
 These tasks were all taken from the \textit{Rijksmuseum Challenge} dataset \citep{mensink14icmr}, and fall within the domain of art classification.
@@
 \texttt{ConvNext} was a close second in the rankings.
-It is interesting that this best performing CNN was inspired by VTs, while the best performing VT \texttt{Swin} took inspiration from CNNs.
+It is interesting that this best performing CNN was inspired by VTs, while the best performing VT, \texttt{Swin}, took inspiration from CNNs.
 It shows the kind of cross-fertilization that moves the field forward, but also hints at something bigger that would be nice to conclude with.

### 2022-07-29 writing/tex/draft_experiments.tex
Made all the changes that I wrote down on the printed out draft. Up next: throwing it through Docs to check grammar, and having it read by one or two friends!

@@
 
-\subsubsection{Off the shelf learning} \label{results:ots}
+\subsubsection{Off-the-shelf learning} \label{results:ots}
 
@@
     \input{img/ots_mean_accuracy.pdf_tex}
-    \caption{Average validation accuracy when training with an off the shelf learning scheme.
-It can be observed that the \texttt{Swin} VT achieves the highest performance overall, but with the \texttt{ConvNext} CNN being a close second.
+    \caption{Average validation accuracy when training with an off-the-shelf learning scheme.
+It can be observed that the \texttt{Swin} VT achieves the highest performance overall, but that the \texttt{ConvNext} CNN is a close second.
 More general, VTs perform on par with CNNs, if not slightly better.}
@@
 }
-\caption{Testing performance after off the shelf learning.
+\caption{Testing performance after off-the-shelf learning.
 Results are similar to validation accuracy shown in figure \ref{results:img:ots}, with \texttt{Swin} and \texttt{ConvNext} showing respective best and second best performance in all cases.}
@@
     \caption{Average validation accuracy when fine-tuning models on the target task.
-Overall, performance is higher compared to off the shelf learning in figure \ref{results:img:ots}, while the differences between models are also smaller here.
-The \texttt{Swin} VT and \texttt{ConvNext} CNN remain the best performing models, while in general, results are less favourable for VTs than they were for off the shelf learning.}
+Overall, performance is higher compared to off-the-shelf learning in figure \ref{results:img:ots}, while the differences between models are also smaller here.
+The \texttt{Swin} VT and \texttt{ConvNext} CNN remain the best performing models, while in general, results are less favourable for VTs than they were after off-the-shelf learning.}
     \label{results:img:ft}
@@
 Results are again similar to validation accuracy in figure \ref{results:img:ft}.
-Compared to table \ref{results:tab:ots}, it is striking that the \texttt{ConvNext} CNN now often takes the lead with respect to balanced accuracy, and that the \texttt{ResNet} CNN often takes second place.}
+Compared to table \ref{results:tab:ots}, it is striking that the \texttt{ConvNext} CNN now often takes the lead with respect to balanced accuracy, and that the \texttt{ResNet50} CNN often takes second place.}
 \label{results:tab:ft}
@@
 
-Results are less favourable for VTs than they were for OTS learning in section \ref{results:ots}.
+Results are less favourable for VTs than they were after OTS learning in section \ref{results:ots}.
 At the same time, it can still be said that VTs perform on par with CNNs.
 The \texttt{Swin} VT shows the best overall testing accuracy here as well, but the \texttt{ConvNext} CNN is not far behind, and outperforms \texttt{Swin} in terms of balanced accuracy.
-Also noteworthy is the \texttt{ResNet} CNN, which is ranked higher here than it was for OTS learning, and now shows second best testing performance for \textit{Artist classification}.
+Also noteworthy is the \texttt{ResNet50} CNN, which is ranked higher here than it was for OTS learning, and now shows second best testing performance for \textit{Artist classification}.
 
@@
 While most results have now been presented, it is still important to discuss them in more detail, as this leads to a more complete interpretation.
-This section attempts to provide such a discussion, by answering questions one might have at this moment.
+This subsection attempts to provide such a discussion, by answering questions one might have at this moment.
 It will first address some of the aforementioned experiments individually, but will conclude with remarks that are more general.
@@
 
-\subsubsection{Off the shelf learning}
+\subsubsection{Off-the-shelf learning}
 
@@
 This differs from the results in table \ref{results:tab:ft}, where \texttt{ViT} does not perform well overall, and even shows the worst testing accuracy on \textit{Type classification}, with 90.11\%.
-\texttt{ResNet50} achieves 91.78\% accuracy here.
-Testing the same \texttt{ResNet101} and \texttt{ResNet152} versions that \citeauthor{zhou2021convnets} used does not change much, as this leads to 91.96\% ($\pm$ 0.59\%) and 91.94\% ($\pm$ 0.25) accuracy, respectively.
+\texttt{ResNet50} achieves 91.78\% accuracy here, and testing the same \texttt{ResNet101} and \texttt{ResNet152} versions that \citeauthor{zhou2021convnets} used does not change much, as this leads to 91.96\% ($\pm$ 0.59\%) and 91.94\% ($\pm$ 0.25) accuracy, respectively.
 
@@
 In that case, it seems, VTs have certain strengths that those other datasets take more advantage of.
-Alternatively, it could also be that hyperparameters and regularisation still leave room for improvement.
+Alternatively, it could be that hyperparameters and regularisation still leave room for improvement.
 The reason for thinking this, is that preliminary experiments suggested that VTs benefit more from regularisation than CNNs do.
@@
     \end{subfigure}
-    \caption{Saliency maps for \texttt{ConvNext}, \texttt{DeiT} and \texttt{Swin} after both off the shelf learning (OTS), and fine-tuning (FT).
-\texttt{ConvNext} appears to benefit more when training the whole model instead of the last layer, because it shows the biggest differences between the OTS learning and FT.
+    \caption{Saliency maps for \texttt{ConvNext}, \texttt{DeiT} and \texttt{Swin} after both off-the-shelf learning (OTS), and fine-tuning (FT).
+\texttt{ConvNext} appears to benefit more when training the whole model instead of the last layer, because it shows the biggest differences between OTS learning and FT.
 In case of image \ref{results:img:sal:dish} this even makes the difference between correct and incorrect classification.}
@@
 
-The FT results in table \ref{results:tab:ft} are all much closer together than the OTS results in table \ref{results:tab:ots}.
+VTs and CNNs are much closer together after FT in table \ref{results:tab:ft} than after OTS learning in table \ref{results:tab:ots}.
 It is as if OTS learning already unlocks much of VTs' potential, while CNNs need an FT learning scheme for this.
@@
 Importantly, 3 different salience mapping techniques were used: GradCam \citep{selvaraju2017grad} was implemented for \texttt{ConvNext}, while Attention Rollout \citep{abnar2020quantifying} was implemented for \texttt{DeiT}.
-This already makes the comparison somewhat unfair.
-This becomes even worse when adding \texttt{Swin}, for which a self-invented method was used.
-This method is similar to GradCam, but instead of multiplying average CNN channel gradients with said channel's activation, it directly multiplies activation with corresponding gradients in the output of one of the later transformer blocks\footnote{For the full implementation of all techniques, one is referred to the source code on: \\ \url{https://github.com/IndoorAdventurer/ViTTransferLearningForArtClassification}.}.
-Despite this, saliency maps were included in this thesis, as they are interesting nevertheless.
+This already makes the comparison somewhat unfair, and it becomes even worse when \texttt{Swin} is added, for which a self-invented method was used.
+This method is similar to GradCam, but instead of multiplying \textit{average} CNN channel gradients with said channel's activation, it directly multiplies activation with corresponding gradients in the output of one of the later transformer blocks\footnote{For the full implementation of all techniques, one is referred to the source code on: \\ \url{https://github.com/IndoorAdventurer/ViTTransferLearningForArtClassification}.}.
+Despite this, saliency maps were included, as they are interesting nonetheless.
 % Show saliency maps here
@@
 On the one hand, it shows much heterogeneity for the \textit{Artist classification} task, with artists working in very different styles and mediums.
-On the other hand, it also shows difficulties for the other tasks: \textit{Material classification}, which seems to be the most challenging task, shows examples of both big differences within classes (note how the two instances of porcelain are very different), and small differences between classes (note how the porcelain plate is very similar to the faience plate).
+On the other hand, it also shows difficulties for the other tasks: \textit{Material classification}, which seems to be the most challenging task, shows examples of both big differences within classes (note how the two instances of porcelain are very different), and small differences between classes (note how the porcelain plate is similar to the faience plate).
 
@@
 It was mentioned earlier that balanced accuracy falls behind plain accuracy the most for \textit{Material classification}, and the least for \textit{Artist classification}.
-At the same time, table \ref{methods:datasets} shows that it is also \textit{Material classification} which is the least balanced (gini-coefficient of 0.563), while it is \textit{Artist classification} that is the most balanced (coefficient 0.236).
-This suggests that balance is hurting performance the least for \textit{Artist classification} -- and the most for \textit{Material classification}.
+At the same time, table \ref{methods:datasets} shows that it is also \textit{Material classification} which is the least balanced (Gini-coefficient of 0.563), while it is \textit{Artist classification} that is the most balanced (coefficient 0.236).
+This suggests that balance has a positive effect on testing accuracy, from which \textit{Artist classification}, in that case, benefits the most.
 
@@
 Notable exceptions are \texttt{ConvNext} and \texttt{EfficientNetV2}, which were selected because they are equally recent as the used VTs.
-Many of these models are popular/common, but one can't say they are a representative sample.
-The question, of course, is also not how VTs and CNNs as a group compare to one another.
+Many of these models are popular/common, but one can't say that they are a representative sample.
+The question, of course, also is not how VTs and CNNs as a group compare to one another.
 Instead, the question is if the VT landscape as a whole has something to offer that could be an interesting alternative to common CNNs.
@@
 
-Of course, final accuracy is not the full story, since time/accuracy trade-offs, for example, might also determine which model is preferred.
-Figure \ref{results:img:ots_vs_ft_type} plots \textit{Type classification} accuracies of tables \ref{results:tab:ots} and \ref{results:tab:ft} against training time per epoch.
+Of course, final accuracy is not the full story, since time/accuracy trade-offs, for example, can also determine which model is preferred.
+Figure \ref{results:img:ots_vs_ft_type} plots the \textit{Type classification} accuracies of tables \ref{results:tab:ots} and \ref{results:tab:ft} against training time per epoch.
 Total training time was not chosen instead, as this measure depends much more on selected hyperparameters.
@@
 The best overall performing model is a fine-tuned VT (\texttt{Swin}).
-At the same time, it is a CNN that shows the best time/accuracy trade-offs, because it is able to perform better than any OTS trained model, while still taking less time per epoch than any OTS trained VT.
+At the same time, it is perhaps a CNN that shows the best time/accuracy trade-offs, because it is able to perform better than any OTS trained model, while still taking less time per epoch than any OTS trained VT.
 The particular CNN this regards, is \texttt{ResNet50}.
@@
 
-All `tiny' VTs are around the same size as \texttt{ResNet50}, and using these would have changed the situation here completely.
+All `tiny' VTs are around the same size as \texttt{ResNet50}, and using these would have changed the situation completely.
 As such, taking figure \ref{results:img:ots_vs_ft_type} with a grain of salt, would be appropriate.

### 2022-07-29 writing/tex/draft_introduction.tex
Finished the abstract earlier today. With that, the first draft is done!
Made all the changes that I wrote down on the printed out draft. Up next: throwing it through Docs to check grammar, and having it read by one or two friends!

@@
 Since the introduction of AlexNet, roughly a decade ago, \textit{Convolutional Neural Networks} (CNNs) have played an important role in \textit{Computer Vision} (CV) \citep{krizhevsky2012imagened}.
-Recently, however, a new type of architecture, called \textit{Vision Transformer} (VT), gained state of the art performance on common learning benchmarks, including the ImageNet dataset\footnote{At the time of writing the VT \texttt{CoCa} is state of the art, according to \url{https://paperswithcode.com/sota/image-classification-on-imagenet}.}.
+Recently, however, a new type of architecture, called \textit{Vision Transformer} (VT), gained state of the art performance on common learning benchmarks, including the ImageNet dataset \citep{deng2009imagenet}\footnote{At the time of writing, the VT \texttt{CoCa} is state of the art, according to \url{https://paperswithcode.com/sota/image-classification-on-imagenet}.}.
 
@@
 VTs were first introduced in \citeauthor{dosovitskiy2020image} (\citeyear{dosovitskiy2020image}), and are based on the \textit{encoder stack} of the \textit{Transformer} architecture \citep{vaswani2017attention}.
-This stack takes a sequence of vectors as input, and is composed of alternating \textit{multi-headed self-attention} layers and position-wise MLP-type layers -- position-wise meaning the MLP can take only a single position (constituent vector) as input at a time, and is shared between all positions.
+This stack takes a sequence of vectors as input, and is composed of alternating \textit{multi-headed self-attention} layers and position-wise MLP-type layers -- position-wise meaning the MLP's input is the size of a single position (constituent vector), and is shared between all positions.
 
@@
 To make it work for CV applications, VTs divide an image into square patches, which get treated as tokens (words).
-Learned embeddings convert these into a sequence of vectors, to which 1D positional encodings are added\footnote{Do mind that VTs have no built-in knowledge of an image's 2D topology.}.
+Learned embeddings convert these into a sequence of vectors, to which 1D positional embeddings are added\footnote{Do mind that VTs have no built-in knowledge of an image's topology, as the architecture ignores the 2D layout.}.
 A learned $[class]$ vector gets prepended to the sequence, and its representation at the end of the encoder stack is the final output, which can then be used for classification, segmentation, etc.
@@
 Moreover, weight sharing applies as much to VTs as it does to CNNs.
-In VTs, each position goes through the same computational pipelines of being multiplied by 3 projection matrices, followed by a weighted sum calculation, followed by being fed through an MLP-type layer\footnote{Neglecting skip connections and normalization, as this is a comparison on an intuitive level.}.
+In VTs, each position goes through the same computational pipeline of being multiplied by 3 projection matrices, followed by a weighted sum calculation, followed by being fed through an MLP-type layer\footnote{Neglecting skip connections, normalization and multi-headedness, as this is a comparison on an intuitive level.}.
 In CNNs these pipelines take the form of convolution kernels.
@@
 \paragraph{A limitation of VTs}
-While VTs are promising; the fact that they are not tailored to CV tasks, makes that they lack a so called image specific \textit{Inductive Bias}, which CNNs do have.
-Intuitively this means a VT first has to learn what an image is, before it can move on to the learning task at hand.
+While VTs are promising; the fact that they are not specifically designed for CV tasks, makes that they lack a so called image specific \textit{Inductive Bias}, which CNNs do have.
+Intuitively this means a VT first has to learn what an image is, before it can move on to the learning task at hand\footnote{Recall, for example, that the architecture ignores the 2D layout of an image, and hence has to learn about this through training.}.
 The result is that they require substantially more training data (and time) than CNNs do.
@@
 In practice, many TL techniques exist.
-The example given in section \ref{vits} can be described as \textit{fine-tuning} (FT) TL.
+The example given in section \ref{vits} can be described as \textit{fine-tuning} (FT).
 Here, a pre-trained model is used as a starting point, and then trained on the target task -- possibly with different hyperparameters, and often with a new final layer that fits the task.
@@
 
-A different TL technique is called \textit{off-the-shelf} (OTS) TL.
+Another TL technique is called \textit{off-the-shelf} (OTS) learning.
 It is similar to FT, except that all but the final layer is kept frozen during training.
-\citeauthor{sharif2014cnn} (\citeyear{sharif2014cnn}) found that features extracted by a CNN pre-trained on ImageNet were general enough for OTS TL to work.
+\citeauthor{sharif2014cnn} (\citeyear{sharif2014cnn}) found that features extracted by a CNN pre-trained on ImageNet were general enough for OTS learning to work.
 More specifically, these OTS CNNs showed better performance than the (fully trained non-CNN) state of the art at the time.
-Besides reducing computational cost of training, a benefit of OTS TL is that it can prevent overfitting in cases where FT would not \citep{yosinski2014transferable}.
+Besides reducing computational cost of training, a benefit of OTS learning is that it can prevent overfitting in cases where FT would not \citep{yosinski2014transferable}.
 
@@
 This domain is characterized by small datasets, so to overcome the inductive bias problem described earlier, TL is imperative.
-It was found that VTs benefit more from FT TL than CNNs do, and that VTs pre-trained on ImageNet perform on par CNNs.
+It was found that VTs benefit more from FT than CNNs do, and that VTs pre-trained on ImageNet perform on par with CNNs after this type of TL.
 
@@
 
-Another `shortcomming' of \citeauthor{matsoukas2021time} is that it merely examined whether VTs can use TL to outperform CNNs in a domain that happens to be characterized by limited-sized datasets, but that it did not actively focus on the extent to which VTs can be made useful in limited data/compute environments.
+Another `shortcomming' of \citeauthor{matsoukas2021time} is that it merely examined whether VTs can use TL to outperform CNNs in a domain that happens to be characterised by limited-sized datasets, but that it did not actively focus on the extent to which VTs can be made useful in limited data/compute environments.
 More related to this is \citeauthor{touvron2021training} (\citeyear{touvron2021training}), which proposed transformer-specific distillation-based learning techniques to make VTs applicable for smaller datasets.
-This is not necessarily related to TL as defined in this paper, however.
+This is not related to FT or OTS learning, of course.
 % matsoukas2021time not specifically focussed on seeing to what degree TL allows VTs to dominate CNNs in limited compute/data environments
 \\\\
-The current study will examine to what extent TL techniques (both FT and OTS) allow VTs to become the preferred architecture when compute resources are limited, and attempts to find a methodology that is recommended in these circumstances.
-It will, however, restrict itself to art classification tasks.
+The current thesis will examine to what extent TL techniques (both FT and OTS) allow VTs to become the preferred architecture when resources are limited, and attempts to find a methodology that is recommended in these circumstances.
 Several experiments will be performed, of which results will be analysed both quantitatively as well as qualitatively.
 
-%TODO: talk about datasets small enough to be hand labelled by one person
+Limited resources, in this case, primarily refers to the amount of available training data.
+Datasets should be so small that one can reasonable assume they were hand labelled by a single person.
+In addition, it also refers to compute resources, which should be restricted to consumer-grade hardware.
+In practice, of course, the small datasets already ensure that this restriction is met, because these naturally allow such hardware to suffice for learning.
+
+Furthermore, this thesis restricts itself to art classification tasks specifically.
+These tasks fit in well with the described objectives, since the amount of effort needed to create and digitise an art collection often makes for small datasets.
+Moreover, they provide an interesting TL challenge, as knowledge might have to be carried over from natural tot non-natural images.
 %The current study limits itself to the task of identifying which materials were used in artworks.
@@
 
-More formally, the following research question is proposed: \begin{quote}\textit{To what extent (and in what manners) does utilizing TL allow VTs to perform on par with CNNs on common art classification problems, when compute resources are limited to a single GPU?}\end{quote}
+To put all of this more formally, the following research question is proposed: \begin{quote}\textit{To what extent (and in what manners) does utilising TL allow VTs to perform on par with CNNs on common art classification problems, when datasets contain 10,000 images or less?}\end{quote}

### 2022-07-29 writing/tex/draft_methods.tex
Made all the changes that I wrote down on the printed out draft. Up next: throwing it through Docs to check grammar, and having it read by one or two friends!

@@
 The main focus was on a comparison between CNN architectures pre-trained on ImageNet1K when they were either used as OTS feature extractors, or fine-tuned to the new task.
-It was suggested that FT resulted in substantially better performance than either training from scratch or using an OTS learning scheme.
+It was suggested that fine-tuning resulted in substantially better performance than either training from scratch or using an OTS learning scheme.
 
@@
 
-Similar to \citeauthor{sabatelli2018deep}, used models are pre-trained on ImageNet1K, while the \textit{Rijksmuseum Challenge} dataset \citep{mensink14icmr} is used for the target tasks.
+Similar to \citeauthor{sabatelli2018deep}, used models are pre-trained on ImageNet1K, and the \textit{Rijksmuseum Challenge} dataset \citep{mensink14icmr} is used for the target tasks.
 This dataset consists of a large collection of digitized artworks, together with xml-formatted metadata.
-The classification tasks extracted from it are: (1) \textit{Type classification}, containing labels as `painting', `sculpture', `drawing', etc.; (2) \textit{Material classification}, with labels as `paper', `porcelain', `silver', etc.; and finally, (3) \textit{Artist classification}, where the model has to predict who the creator is.
+The classification tasks extracted from it are: (1) \textit{Type classification}, where the model has to distinguish between classes as `painting', `sculpture', `drawing', etc.; (2) \textit{Material classification}, with classes as `paper', `porcelain', `silver', etc.; and finally, (3) \textit{Artist classification}, where the model has to predict who the creator is.
 Figure \ref{methods:eg} shows labelled examples for all three tasks.
@@
 The full dataset contains 112,039 images, and allows for multiple or zero labels per classification task.
-Samples are excluded from a task if they contain no labels for it.
+Naturally, samples are excluded from a task if they contain no labels for it.
 Samples containing more than one label are excluded as well, since analysis showed that in all cases this is true for only a small portion of the dataset.% Note that this last point is different from \citeauthor{sabatelli2018deep}, which took the first occurring class in these cases.
@@
 All datasets are split up into 80\%, 10\%, 10\% training, validating and testing sets.
-For robustness, five subsets are randomly generated per classification task, such that each experiment can be run five times.
-The average sample overlap between two of the five datasets is shown in table \ref{methods:datasets} as well.
+For robustness, five datasets are randomly generated per classification task, such that each experiment can be run five times.
+The average sample overlap between two of the five sets is shown in table \ref{methods:datasets} as well.
 
@@
 The 5 datasets produced in this manner are then respectively 100\%, 56\%, 32\%, 18\% and 10\% the size of the dataset shown in table \ref{methods:datasets}.
-Steps are taken to ensure the distribution remains the same.
+Steps are taken to ensure that the distribution remains the same.
 Finally, also here 5 datasets are generated per scaling factor, leading to a total of 25 datasets.
@@
 
-To start off with CNNs, \texttt{ResNet50} \citep{he2016deep} and \texttt{VGG19} \citep{simonyan2014very} are chosen, as these were the best performing models in \citeauthor{sabatelli2018deep} for respectively FT and OTS TL.
+To start off with CNNs, \texttt{ResNet50} \citep{he2016deep} and \texttt{VGG19} \citep{simonyan2014very} are chosen, as these were the best performing models in \citeauthor{sabatelli2018deep} for respectively FT and OTS learning.
 In addition, \citeauthor{matsoukas2021time} (\citeyear{matsoukas2021time}) used \texttt{ResNet50} too, while \citeauthor{zhou2021convnets} (\citeyear{zhou2021convnets}) used \texttt{ResNet101} and \texttt{ResNet152}.
@@
 The first model is the original VT \citep{dosovitskiy2020image}, which will be referred to as \texttt{ViT}.
-Next is \texttt{Swin} \citep{liu2021swin}, because it showed promising performance in \citeauthor{zhou2021convnets}.
+Next, \texttt{Swin} \citep{liu2021swin} is considered, as it showed promising performance in \citeauthor{zhou2021convnets}.
 The final two are \texttt{DeiT} and \texttt{BeiT}.
 The small version of \texttt{DeiT} was used in \citeauthor{matsoukas2021time} (\citeyear{matsoukas2021time}), because it is similar in size to \texttt{ResNet50}, to which it was compared.
-In this work, however, also a base-sized version is chosen.
+In this thesis, however, also a base-sized version is chosen.
 Specifically, the version that does not include output for distillation learning.
@@
 \texttt{ResNet50} and \texttt{EfficientNetV2} are smaller, with 25.6 and 13.6 million, respectively, while \texttt{VGG19} is the largest model, with 143.7 million parameters.
-It should be mentioned that for both OTS TL and FT, preliminary experiments were performed with `small' and `tiny' models.
-The results of these suggested that similar conclusions could have been drawn if these were used instead.
 
@@
 
-For OTS TL, the Adam optimizer is used with standard \texttt{PyTorch} parameters (\texttt{lr}=1e-3, $\beta_1=0.9$, $\beta_2=0.999$), and a batch size of 256.
+For OTS learning, the Adam optimizer \citep{kingma2014adam} is used with standard \texttt{PyTorch} parameters (\texttt{lr}=1e-3, $\beta_1=0.9$, $\beta_2=0.999$), and a batch size of 256.
 For FT, more is needed to achieve good convergence, especially for VT-based architectures.
@@
 The FT experiments take advantage of the V100's mixed precision acceleration.
-An exception is made for the type classification experiment, as this one is also used to compare OTS TL and FT in terms of time/accuracy trade-offs.
+An exception is made for the type classification experiment, as this one is also used to compare OTS learning and FT in terms of time/accuracy trade-offs.
 Preliminary findings suggested that using mixed precision does not affect performance for the aforementioned experiments, however.
 
-The \texttt{PyTorch} machine learning framework is used, and many pre-trained models are taken from its \texttt{Torchvision} library.
+The \texttt{PyTorch} machine learning framework is used \citep{paszke2019pytorch}, and many pre-trained models are taken from its \texttt{Torchvision} library.
 Exceptions are \texttt{EfficientNetV2}, \texttt{Swin}, \texttt{DeiT} and \texttt{BeiT}, which are taken from the \texttt{Timm} library\footnote{\url{https://timm.fast.ai/}.}.
 
-Finally, all source code and data will be made available on GitHub\footnote{\url{https://github.com/IndoorAdventurer/ViTTransferLearningForArtClassification}.}.
+Finally, all source code and data will be made available on \texttt{GitHub}\footnote{\url{https://github.com/IndoorAdventurer/ViTTransferLearningForArtClassification}.}.

### 2022-07-30 writing/report.tex
Made all the changes that I wrote down on the printed out draft. Up next: throwing it through Docs to check grammar, and having it read by one or two friends!
Made changes suggested by Grammarly

@@
   \begin{@twocolumnfalse}
-   \input{tex/draft_abstract.tex}
+   \input{tex/abstract.tex}
   \end{@twocolumnfalse}
@@
 
-\input{tex/draft_introduction.tex}
-\input{tex/draft_methods.tex}
-\input{tex/draft_experiments.tex}
-\input{tex/draft_conclusion.tex}
+\input{tex/introduction.tex}
+\input{tex/methods.tex}
+\input{tex/experiments.tex}
+\input{tex/conclusion.tex}
 
 \section*{Acknowledgements}
-\textit{The author would like to thank {the Center for Information Technology of the University of Groningen} for their support and for providing access to the Peregrine high performance computing cluster.
+\textit{The author would like to thank {the Center for Information Technology of the University of Groningen} for their support and for providing access to the Peregrine high-performance computing cluster.
 He would also like to thank {dr.

### 2022-07-30 writing/tex/abstract.tex
Made changes suggested by Grammarly

--- 2022-07-20 writing/tex/draft_abstract.tex
@@
 \begin{abstract} {Convolutional Neural Networks (CNNs) have become the de facto standard in Computer Vision, and played a major role in the advancement of this field.
-In recent years, however, Vision Transformer based models (VTs) have been outperforming CNNs across the board.
-While this is exciting, CNNs still have an advantage on small datasets, due to their so-called image specific inductive bias.
+In recent years, however, Vision-Transformer-based models (VTs) have been outperforming CNNs across the board.
+While this is exciting, CNNs still have an advantage on small datasets, due to their so-called image-specific inductive bias.
 Recent work suggests that transfer learning methods allow VTs to compete with CNNs on some of these small datasets.
@@
 To this end, it compares popular VTs and CNNs in terms of their off-the-shelf and fine-tuning transferability, when going from \textit{ImageNet1K} to target tasks provided by the \textit{Rijksmuseum Challenge} dataset.
-The results show that VTs posses superior off-the-shelf feature extraction capabilities here, and that in general, transfer learning allows VTs to become an interesting alternative to consider within the digital humanities.\vspace{8ex}}
+The results show that VTs possess superior off-the-shelf feature extraction capabilities here, and that in general, transfer learning allows VTs to become an interesting alternative to consider within the digital humanities.\vspace{8ex}}
 \end{abstract}

### 2022-07-30 writing/tex/conclusion.tex
Made changes suggested by Grammarly

--- 2022-07-29 writing/tex/draft_conclusion.tex
@@
 \begin{itemize}
-\item small datasets are used: specifically, small enough to have been hand labelled by one person;
-\item only consumer grade hardware is available.
+\item small datasets are used: specifically, small enough to have been hand-labeled by one person;
+\item only consumer-grade hardware is available.
 \end{itemize}
@@
 
-To that end, it tried to utilise transfer learning methods, as these might help overcome problems that arise from VTs' lack of an inductive bias.
+To that end, it tried to utilize transfer learning methods, as these might help overcome problems that arise from VTs' lack of an inductive bias.
 More concretely, models pre-trained on ImageNet1K were either used as off-the-shelf feature extractors, or fine-tuned to new tasks.

### 2022-07-30 writing/tex/experiments.tex
Made changes suggested by Grammarly

--- 2022-07-29 writing/tex/draft_experiments.tex
@@
 \subsection{Results} \label{results}
-Results are presented in the form of tables and line plots. The former of these report final testing performance, whereas the latter mostly visualise validation accuracy during training.
+Results are presented in the form of tables and line plots. The former of these reports final testing performance, whereas the latter mostly visualizes validation accuracy during training.
 
-To give a more precise description: tables show average testing performance over the 5 trials per experiment, with standard deviation ($s$) as subscript.
+To give a more precise description: tables show average testing performance over the 5 trials per experiment, with standard deviation ($s$) as a subscript.
 Besides accuracy, also balanced accuracy is reported.
-This measure is defined as the average recall over all classes, such that it penalises errors on less occurring classes more than plain accuracy would.
+This measure is defined as the average recall over all classes, such that it penalizes errors on less occurring classes more than plain accuracy would.
 Lastly, green shaded cells mark best performance, while yellow and red mark second best and worst performance, respectively.
@@
 In cases where they show average validation accuracy, plots end when an early stop occurred for the first of 5 trials.
-To easily distinguish VTs from CNNs, VTs are plotted with continues lines, and CNNs with dashed ones.
+To easily distinguish VTs from CNNs, VTs are plotted with continuous lines, and CNNs with dashed ones.
 
@@
 \caption{Testing performance after off-the-shelf learning.
-Results are similar to validation accuracy shown in figure \ref{results:img:ots}, with \texttt{Swin} and \texttt{ConvNext} showing respective best and second best performance in all cases.}
+Results are similar to validation accuracy shown in figure \ref{results:img:ots}, with \texttt{Swin} and \texttt{ConvNext} showing the respective best and second best performance in all cases.}
 \label{results:tab:ots}
@@
 
-Table \ref{results:tab:ots} shows testing performance of OTS trained models on the 3 classification tasks, while figure \ref{results:img:ots} shows corresponding validation accuracies.
+Table \ref{results:tab:ots} shows the testing performance of OTS-trained models on the 3 classification tasks, while figure \ref{results:img:ots} shows corresponding validation accuracies.
 Performance on the testing set is comparable among all tasks in terms of rankings.
@@
 VTs perform very much on par with CNNs, if not slightly better.
-The best performing model, for instance, is the \texttt{Swin} VT, which shows the highest testing performance without exception.
+The best-performing model, for instance, is the \texttt{Swin} VT, which shows the highest testing performance without exception.
 Moreover, all VTs except \texttt{BeiT} are positioned relatively high in the rankings.
@@
 
-Finally, when comparing the 3 classification tasks to one another, it can be observed that highest overall accuracies are achieved for \textit{Artist classification}, and lowest ones for \textit{Material classification}.
-For balanced accuracy the differences are more extreme, as this measure also falls behind plain accuracy the most for \textit{Material classification}, and the least for \textit{Artist classification}.
+Finally, when comparing the 3 classification tasks to one another, it can be observed that the highest overall accuracies are achieved for \textit{Artist classification}, and the lowest ones for \textit{Material classification}.
+For balanced accuracy, the differences are more extreme, as this measure also falls behind plain accuracy the most for \textit{Material classification}, and the least for \textit{Artist classification}.
 
@@
 Overall, performance is higher compared to off-the-shelf learning in figure \ref{results:img:ots}, while the differences between models are also smaller here.
-The \texttt{Swin} VT and \texttt{ConvNext} CNN remain the best performing models, while in general, results are less favourable for VTs than they were after off-the-shelf learning.}
+The \texttt{Swin} VT and \texttt{ConvNext} CNN remain the best performing models, while in general, results are less favorable for VTs than they were after off-the-shelf learning.}
     \label{results:img:ft}
@@
 
-Results are less favourable for VTs than they were after OTS learning in section \ref{results:ots}.
+Results are less favorable for VTs than they were after OTS learning in section \ref{results:ots}.
 At the same time, it can still be said that VTs perform on par with CNNs.
 The \texttt{Swin} VT shows the best overall testing accuracy here as well, but the \texttt{ConvNext} CNN is not far behind, and outperforms \texttt{Swin} in terms of balanced accuracy.
-Also noteworthy is the \texttt{ResNet50} CNN, which is ranked higher here than it was for OTS learning, and now shows second best testing performance for \textit{Artist classification}.
+Also noteworthy is the \texttt{ResNet50} CNN, which is ranked higher here than it was for OTS learning, and now shows the second best testing performance for \textit{Artist classification}.
 
-VTs and CNNs grouped together are much more evenly matched.
+VTs and CNNs grouped are much more evenly matched.
 To again take \textit{Type classification} as an example: the mean testing accuracy of all VTs combined is 91.2\% on this task, while it is 91.3\% for CNNs.
-Recall that section \ref{results:ots} reported a difference of 1\% in favour of VTs here, with an accuracy of 86.5\% for VTs, and 85.5\% for CNNs.
+Recall that section \ref{results:ots} reported a difference of 1\% in favor of VTs here, with an accuracy of 86.5\% for VTs, and 85.5\% for CNNs.
 
-More general, it can be observed that FT leads to substantially better performance than OTS learning, and that models are now much closer together.
-The difference between highest and lowest testing accuracy, for example, is at most 3.8\% after FT, while it was between 7.2\% and 11.8\% after OTS learning.
+More generally, it can be observed that FT leads to substantially better performance than OTS learning, and that models are now much closer together.
+The difference between the highest and lowest testing accuracy, for example, is at most 3.8\% after FT, while it was between 7.2\% and 11.8\% after OTS learning.
 In addition, the worst testing accuracy after FT is most often still higher than the best accuracy after OTS learning.
-The only exception is \textit{Material classification}, where the worst FT model (\texttt{BeiT}) has accuracy 85.74\%, and the best OTS one (\texttt{Swin}) has 85.87\%.
+The only exception is \textit{Material classification}, where the worst FT model (\texttt{BeiT}) has an accuracy of 85.74\%, and the best OTS one (\texttt{Swin}) has 85.87\%.
 
-Lastly, note that highest overall performances are again achieved on the \textit{Artist classification} task, while the lowest ones are reported for \textit{Material classification}.
+Lastly, note that the highest overall performances are again achieved on the \textit{Artist classification} task, while the lowest ones are reported for \textit{Material classification}.
 The difference between plain and balanced accuracy is also again the largest for \textit{Material classification}, and lowest for \textit{Artist classification}.
@@
     \caption{Testing accuracy as datasets gradually become smaller.
-The x-axes show a logarithmic scale, with the smallest value being roughly 10\% the size of largest one.
+The x-axes show a logarithmic scale, with the smallest value being roughly 10\% the size of the largest one.
 Observations done in sections \ref{results:ots} and \ref{results:ft} appear to hold up well as the training set is shrunk.}
@@
 It is not true, for example, that at some point CNNs start to outperform VTs.
-In addition, \texttt{Swin} and \texttt{ConvNext} remain some of the best performing models throughout.
+In addition, \texttt{Swin} and \texttt{ConvNext} remain some of the best-performing models throughout.
 
@@
 This subsection attempts to provide such a discussion, by answering questions one might have at this moment.
-It will first address some of the aforementioned experiments individually, but will conclude with remarks that are more general.
+It will first address some of the aforementioned experiments individually, but will conclude with more general remarks.
 Topics will include: comparisons with earlier studies, potential shortcomings, and hypotheses that might explain certain observations.
@@
 Of all related work mentioned in section \ref{related_work}, \citeauthor{zhou2021convnets} (\citeyear{zhou2021convnets}) was the only one that examined VTs' OTS learning properties.
-It reported best OTS learning performance for the \texttt{Swin} VT, which is much in line with results in table \ref{results:tab:ots}.
+It reported the best OTS learning performance for the \texttt{Swin} VT, which is much in line with the results in table \ref{results:tab:ots}.
 
@@
 On \textit{Type classification}, for example, it achieves a mean testing accuracy of  $86.21\% (\pm 0.70\%)$, which is indeed slightly higher than \texttt{ViT}'s 86.06\% shown in table \ref{results:tab:ots}.
-Mind that these differences are slight compared to the 89.43\% accuracy achieved by \texttt{Swin} here, so replacing \texttt{ResNet50} by \texttt{ResNet101} would not have affected the overall observations that much.
+Mind that these differences are slight compared to the 89.43\% accuracy achieved by \texttt{Swin} here, so replacing \texttt{ResNet50} with \texttt{ResNet101} would not have affected the overall observations that much.
 
@@
     \caption{Attention layers of successively deeper transformer blocks.
-One can recognise the input image in earlier layers, while for later ones activation seems sporadic.
+One can recognize the input image in earlier layers, while for later ones, activation seems sporadic.
 For \texttt{DeiT}, attention with respect to the class-token is plotted.
@@
 While results are in line with previous studies, it was still surprising to see that VTs showed such good OTS learning performance in section \ref{results:ots}.
-It suggests that these transformer based architectures are able to give a very complete, high level encoding of the input image.
-For CNNs this is commonly thought to be true, as they are often described as hierarchical feature extractors that build successively higher level feature maps from lower level ones (see section \ref{intro:cnns}).
+It suggests that these transformer-based architectures are able to give a very complete, high-level encoding of the input image.
+For CNNs this is commonly thought to be true, as they are often described as hierarchical feature extractors that build successively higher-level feature maps from lower-level ones (see section \ref{intro:cnns}).
 If the same is true for VTs, is not well understood at this moment.
@@
 A hint of an answer can, however, be found when plotting activation of successively deeper attention layers, as is done in figure \ref{results:img:layers}.
-While earlier layers retain much spatial information, later ones show more sporadic activation patters.
-This is what one might expect from an hierarchical feature extractor: the random pixels lighting up might, for example, correspond to specific, high level properties of the input.
-Nevertheless, it should be emphasised that no strong conclusions should be drawn from these images, and that more research with respect to explainability is desired.
+While earlier layers retain much spatial information, later ones show more sporadic activation patterns.
+This is what one might expect from a hierarchical feature extractor: the random pixels lighting up might, for example, correspond to specific, high-level properties of the input.
+Nevertheless, it should be emphasized that no strong conclusions should be drawn from these images, and that more research with respect to explainability is desired.
 % Also mention that this was unexpected (but in line with blabla);
@@
 
-Results are also less favourable for VTs compared to \citeauthor{matsoukas2021time} (\citeyear{matsoukas2021time}).
-In that paper the `tiny' version of \texttt{DeiT} outperforms \texttt{ResNet50}, while in table \ref{results:tab:ft} this same \texttt{ResNet50} performs slightly better than the `base' version of \texttt{DeiT}.
-Going `tiny' also does not help: on \textit{Type classificaiton}, for instance, `tiny' \texttt{DeiT} gets 90.44\% ($\pm$ 0.93\%), which is even worse than `base' \texttt{DeiT}'s 91.78\% in table \ref{results:tab:ots}.
+Results are also less favorable for VTs compared to \citeauthor{matsoukas2021time} (\citeyear{matsoukas2021time}).
+In that paper, the `tiny' version of \texttt{DeiT} outperforms \texttt{ResNet50}, while in table \ref{results:tab:ft} this same \texttt{ResNet50} performs slightly better than the `base' version of \texttt{DeiT}.
+Going `tiny' also does not help: on \textit{Type classification}, for instance, `tiny' \texttt{DeiT} gets 90.44\% ($\pm$ 0.93\%), which is even worse than `base' \texttt{DeiT}'s 91.78\% in table \ref{results:tab:ots}.
 
@@
 
-\paragraph{Why are results less favourable for VTs after FT than after OTS learning?}
+\paragraph{Why are results less favorable for VTs after FT than after OTS learning?}
 
@@
     \end{subfigure}
-    \caption{Saliency maps for \texttt{ConvNext}, \texttt{DeiT} and \texttt{Swin} after both off-the-shelf learning (OTS), and fine-tuning (FT).
+    \caption{Saliency maps for \texttt{ConvNext}, \texttt{DeiT}, and \texttt{Swin} after both off-the-shelf learning (OTS), and fine-tuning (FT).
 \texttt{ConvNext} appears to benefit more when training the whole model instead of the last layer, because it shows the biggest differences between OTS learning and FT.
-In case of image \ref{results:img:sal:dish} this even makes the difference between correct and incorrect classification.}
+In the case of image \ref{results:img:sal:dish}, this even makes the difference between correct and incorrect classification.}
     \label{results:img:sal}
@@
 
-The images in figure \ref{results:img:sal} are representative examples, and show observation that could have been made from other images too.
+The images in figure \ref{results:img:sal} are representative examples, and show observations that could have been made from other images too.
 How representative the 3 models are for all 8 is debatable, and was not tested.
@@
 The final learning scheme was chosen after extensive hyperparameter optimization, and was partially inspired by related work.
-It is possible that doing this for each model individually, would have resulted in \textit{slightly} higher performances.
+It is possible that doing this for each model individually would have resulted in \textit{slightly} higher performances.
 This paper, however, does not aim to get the best possible performance, but instead wants to see how models compare to one another.
@@
 
-\paragraph{To what extent are the chosen models representative for all VTs and CNNs?}
+\paragraph{To what extent are the chosen models representative of all VTs and CNNs?}
 Most models were selected because of their mention in related work.
@@
 
-It is also interesting to wonder if, for example, `tiny' or `medium' sized version should have been used instead of `base' versions.
+It is also interesting to wonder if, for example, `tiny'- or `medium'-sized version should have been used instead of `base' versions.
 All available `tiny' models were tested as well, and results have been reported where appropriate (e.g.
@@
 The best overall performing model is a fine-tuned VT (\texttt{Swin}).
-At the same time, it is perhaps a CNN that shows the best time/accuracy trade-offs, because it is able to perform better than any OTS trained model, while still taking less time per epoch than any OTS trained VT.
-The particular CNN this regards, is \texttt{ResNet50}.
+At the same time, it is perhaps a CNN that shows the best time/accuracy trade-offs, because it performs better than any OTS-trained model, while still taking less time per epoch than any OTS-trained VT.
+The particular CNN this concerns, is \texttt{ResNet50}.
 Recall that this model only had 25.6 million parameters, while most VTs had around 86 million.
@@
 All `tiny' VTs are around the same size as \texttt{ResNet50}, and using these would have changed the situation completely.
-As such, taking figure \ref{results:img:ots_vs_ft_type} with a grain of salt, would be appropriate.
+As such, taking figure \ref{results:img:ots_vs_ft_type} with a grain of salt would be appropriate.
 

### 2022-07-30 writing/tex/introduction.tex
Made changes suggested by Grammarly

--- 2022-07-29 writing/tex/draft_introduction.tex
@@
 Since the introduction of AlexNet, roughly a decade ago, \textit{Convolutional Neural Networks} (CNNs) have played an important role in \textit{Computer Vision} (CV) \citep{krizhevsky2012imagened}.
-Recently, however, a new type of architecture, called \textit{Vision Transformer} (VT), gained state of the art performance on common learning benchmarks, including the ImageNet dataset \citep{deng2009imagenet}\footnote{At the time of writing, the VT \texttt{CoCa} is state of the art, according to \url{https://paperswithcode.com/sota/image-classification-on-imagenet}.}.
+Recently, however, a new type of architecture, called \textit{Vision Transformer} (VT), gained state-of-the-art performance on common learning benchmarks, including the ImageNet dataset \citep{deng2009imagenet}\footnote{At the time of writing, the VT \texttt{CoCa} is state of the art, according to \url{https://paperswithcode.com/sota/image-classification-on-imagenet}.}.
 
@@
 It is well established (e.g.
-\citeauthor{lecun1998gradient}, \citeyear{lecun1998gradient}; \citeauthor{yosinski2014transferable}, \citeyear{yosinski2014transferable}) that the first layers of a CNN detect low level features, such as edges and corners.
-These are combined in subsequent layers to detect higher order features.
+\citeauthor{lecun1998gradient}, \citeyear{lecun1998gradient}; \citeauthor{yosinski2014transferable}, \citeyear{yosinski2014transferable}) that the first layers of a CNN detect low-level features, such as edges and corners.
+These are combined in subsequent layers to detect higher-order features.
 The final layer should then give a representative summary of the input, such that, for example, an MLP-type layer can extract all the information it needs from it to perform the task at hand.
@@
 \paragraph{A limitation of VTs}
-While VTs are promising; the fact that they are not specifically designed for CV tasks, makes that they lack a so called image specific \textit{Inductive Bias}, which CNNs do have.
+While VTs are promising; the fact that they are not specifically designed for CV tasks makes that they lack a so-called image-specific \textit{Inductive Bias}, which CNNs do have.
 Intuitively this means a VT first has to learn what an image is, before it can move on to the learning task at hand\footnote{Recall, for example, that the architecture ignores the 2D layout of an image, and hence has to learn about this through training.}.
@@
 
-It is for this reason the paper that first introduced VTs \citep{dosovitskiy2020image} mentions using pre-trained models rather than randomly initialised ones, as it allows general knowledge about images to be carried over to the new task.
+It is for this reason the paper that first introduced VTs \citep{dosovitskiy2020image} mentions using pre-trained models rather than randomly initialized ones, as it allows general knowledge about images to be carried over to the new task.
 This technique falls within the category of TL, which will be covered next.
@@
 \citeauthor{yosinski2014transferable} (\citeyear{yosinski2014transferable}) studied transferability properties of CNNs pre-trained on ImageNet, and suggested that FT results in improved performance and generalizability compared to training from scratch.
-Moreover, while transferability decreased as source and target task became more dissimilar, utilizing TL still seemed to be preferred.
+Moreover, while transferability decreased as the source and target task became more dissimilar, utilizing TL still seemed to be preferred.
 
@@
 \citeauthor{sharif2014cnn} (\citeyear{sharif2014cnn}) found that features extracted by a CNN pre-trained on ImageNet were general enough for OTS learning to work.
-More specifically, these OTS CNNs showed better performance than the (fully trained non-CNN) state of the art at the time.
+More specifically, these OTS CNNs showed better performance than the (fully trained non-CNN) state-of-the-art at the time.
 Besides reducing computational cost of training, a benefit of OTS learning is that it can prevent overfitting in cases where FT would not \citep{yosinski2014transferable}.
@@
 
-What \citeauthor{matsoukas2021time} did not investigate, is OTS performance of VTs.
-This was, however, done in \citeauthor{zhou2021convnets} (\citeyear{zhou2021convnets}), which found that VTs pre-trained on ImageNet had worse OTS performance than their CNN counterparts.
+What \citeauthor{matsoukas2021time} did not investigate, is the OTS performance of VTs.
+This was, however, done by \citeauthor{zhou2021convnets} (\citeyear{zhou2021convnets}), which found that VTs pre-trained on ImageNet had worse OTS performance than their CNN counterparts.
 An exception was the \texttt{Swin} architecture, which was the best performing model on all datasets, and uses \textit{(shifted-)window-based} self-attention methods, to make Transformers better suited for CV tasks \citep{liu2021swin}.
@@
 
-Another `shortcomming' of \citeauthor{matsoukas2021time} is that it merely examined whether VTs can use TL to outperform CNNs in a domain that happens to be characterised by limited-sized datasets, but that it did not actively focus on the extent to which VTs can be made useful in limited data/compute environments.
+Another `shortcomming' of \citeauthor{matsoukas2021time} is that it merely examined whether VTs can use TL to outperform CNNs in a domain that happens to be characterized by limited-sized datasets, but that it did not actively focus on the extent to which VTs can be made useful in limited data/compute environments.
 More related to this is \citeauthor{touvron2021training} (\citeyear{touvron2021training}), which proposed transformer-specific distillation-based learning techniques to make VTs applicable for smaller datasets.
@@
 The current thesis will examine to what extent TL techniques (both FT and OTS) allow VTs to become the preferred architecture when resources are limited, and attempts to find a methodology that is recommended in these circumstances.
-Several experiments will be performed, of which results will be analysed both quantitatively as well as qualitatively.
+Several experiments will be performed, of which results will be analyzed both quantitatively as well as qualitatively.
 
 Limited resources, in this case, primarily refers to the amount of available training data.
-Datasets should be so small that one can reasonable assume they were hand labelled by a single person.
+Datasets should be so small that one can reasonably assume they were hand labeled by a single person.
 In addition, it also refers to compute resources, which should be restricted to consumer-grade hardware.
@@
 Furthermore, this thesis restricts itself to art classification tasks specifically.
-These tasks fit in well with the described objectives, since the amount of effort needed to create and digitise an art collection often makes for small datasets.
-Moreover, they provide an interesting TL challenge, as knowledge might have to be carried over from natural tot non-natural images.
+These tasks fit in well with the described objectives, since the amount of effort needed to create and digitize an art collection often makes for small datasets.
+Moreover, they provide an interesting TL challenge, as knowledge might have to be carried over from natural to non-natural images.
 %The current study limits itself to the task of identifying which materials were used in artworks.
@@
 
-To put all of this more formally, the following research question is proposed: \begin{quote}\textit{To what extent (and in what manners) does utilising TL allow VTs to perform on par with CNNs on common art classification problems, when datasets contain 10,000 images or less?}\end{quote}
+To put all of this more formally, the following research question is proposed: \begin{quote}\textit{To what extent (and in what manners) does utilizing TL allow VTs to perform on par with CNNs on common art classification problems, when datasets contain 10,000 images or less?}\end{quote}

### 2022-07-30 writing/tex/methods.tex
Made changes suggested by Grammarly

--- 2022-07-29 writing/tex/draft_methods.tex
@@
 % Experiments build upon Sabatelli et al.
-Much of the experiments and analysis discussed below, builds upon the work done in \citeauthor{sabatelli2018deep} (\citeyear{sabatelli2018deep}).
+Much of the experiments and analysis discussed below, build upon the work done by \citeauthor{sabatelli2018deep} (\citeyear{sabatelli2018deep}).
 This paper investigated TL properties of CNNs in the art classification domain.
@@
 The classification tasks extracted from it are: (1) \textit{Type classification}, where the model has to distinguish between classes as `painting', `sculpture', `drawing', etc.; (2) \textit{Material classification}, with classes as `paper', `porcelain', `silver', etc.; and finally, (3) \textit{Artist classification}, where the model has to predict who the creator is.
-Figure \ref{methods:eg} shows labelled examples for all three tasks.
+Figure \ref{methods:eg} shows labeled examples for all three tasks.
 
@@
 Samples are also unevenly distributed among classes, with Gini-coefficients being as high as 0.98\footnote{The Gini-coefficient is a measure of balance, with 0 meaning all classes have the same number of samples, and 1 meaning all samples belong to a single class.
-A uniform distribution has coefficient $\frac{1}{3}$}.
+A uniform distribution has coefficient $\frac{1}{3}$.}.
 To counteract this, only the 30 most occurring classes are selected, and a cap of 1000 instances per class is enforced by taking a random sample when this limit is exceeded.
@@
 
-All datasets are split up into 80\%, 10\%, 10\% training, validating and testing sets.
+All datasets are split up into 80\%, 10\%, 10\% training, validating, and testing sets.
 For robustness, five datasets are randomly generated per classification task, such that each experiment can be run five times.
@@
 It does this by taking the top 15 classes of type classification, and scaling it 4 times by a factor $\sqrt[4]{\frac{1}{10}} \approx 0.56$.
-The 5 datasets produced in this manner are then respectively 100\%, 56\%, 32\%, 18\% and 10\% the size of the dataset shown in table \ref{methods:datasets}.
+The 5 datasets produced in this manner are then respectively 100\%, 56\%, 32\%, 18\%, and 10\% the size of the dataset shown in table \ref{methods:datasets}.
 Steps are taken to ensure that the distribution remains the same.
@@
 
-For all models, the linear classification layer is replaced as to fit the new task.
-Cross entropy loss uses the raw outputs of this layer (conform \texttt{PyTorch} documentation), together with one-hot encoded prediction targets.
+For all models, the linear classification layer is replaced to fit the new task.
+Cross-entropy loss uses the raw outputs of this layer (conform \texttt{PyTorch} documentation), together with one-hot encoded prediction targets.
 An early stop occurs after 10 epochs without improved loss on the validation set, and the model with the lowest loss is benchmarked on the testing set.
@@
 The Adam learning rate now starts off at 1e-4, and is reduced by a factor 10 after 3 epochs without improvement (i.e.
-patience=2); the batch size is reduced to 32; label smoothing of 0.1 is added to the cross entropy loss, and a dropout layer with p=0.2 is inserted before the classification layer.
+patience=2); the batch size is reduced to 32; label smoothing of 0.1 is added to the cross-entropy loss, and a dropout layer with p=0.2 is inserted before the classification layer.
 Finally, input images are augmented with random horizontal flips and rotations in a $\pm 10 ^\circ$ range.
@@
 All experiments are conducted on a single compute node containing one 32 GB Nvidia V100 GPU.
-These nodes are constituents of the \textit{Peregrine high performance computing cluster}, belonging to the \textit{University of Groningen's Center for Information Technology}.
+These nodes are constituents of the \textit{Peregrine high-performance computing cluster}, belonging to the \textit{University of Groningen's Center for Information Technology}.
 The FT experiments take advantage of the V100's mixed precision acceleration.
@@
 The \texttt{PyTorch} machine learning framework is used \citep{paszke2019pytorch}, and many pre-trained models are taken from its \texttt{Torchvision} library.
-Exceptions are \texttt{EfficientNetV2}, \texttt{Swin}, \texttt{DeiT} and \texttt{BeiT}, which are taken from the \texttt{Timm} library\footnote{\url{https://timm.fast.ai/}.}.
+Exceptions are \texttt{EfficientNetV2}, \texttt{Swin}, \texttt{DeiT}, and \texttt{BeiT}, which are taken from the \texttt{Timm} library\footnote{\url{https://timm.fast.ai/}.}.
 

### 2022-08-07 writing/tex/experiments.tex
Incroporated some changes from Iris' feedback. NOW IT IS DONE. GOING TO HAND IT IN NEXTgit add -A!

@@
 \caption{Testing performance after off-the-shelf learning.
-Results are similar to validation accuracy shown in figure \ref{results:img:ots}, with \texttt{Swin} and \texttt{ConvNext} showing the respective best and second best performance in all cases.}
+Results are similar to validation accuracy shown in Figure \ref{results:img:ots}, with \texttt{Swin} and \texttt{ConvNext} showing the respective best and second best performance in all cases.}
 \label{results:tab:ots}
@@
 
-Table \ref{results:tab:ots} shows the testing performance of OTS-trained models on the 3 classification tasks, while figure \ref{results:img:ots} shows corresponding validation accuracies.
+Table \ref{results:tab:ots} shows the testing performance of OTS-trained models on the 3 classification tasks, while Figure \ref{results:img:ots} shows corresponding validation accuracies.
 Performance on the testing set is comparable among all tasks in terms of rankings.
@@
     \caption{Average validation accuracy when fine-tuning models on the target task.
-Overall, performance is higher compared to off-the-shelf learning in figure \ref{results:img:ots}, while the differences between models are also smaller here.
+Overall, performance is higher compared to off-the-shelf learning in Figure \ref{results:img:ots}, while the differences between models are also smaller here.
 The \texttt{Swin} VT and \texttt{ConvNext} CNN remain the best performing models, while in general, results are less favorable for VTs than they were after off-the-shelf learning.}
@@
 \caption{Testing performance after fine-tuning.
-Results are again similar to validation accuracy in figure \ref{results:img:ft}.
-Compared to table \ref{results:tab:ots}, it is striking that the \texttt{ConvNext} CNN now often takes the lead with respect to balanced accuracy, and that the \texttt{ResNet50} CNN often takes second place.}
+Results are again similar to validation accuracy in Figure \ref{results:img:ft}.
+Compared to Table \ref{results:tab:ots}, it is striking that the \texttt{ConvNext} CNN now often takes the lead with respect to balanced accuracy, and that the \texttt{ResNet50} CNN often takes second place.}
 \label{results:tab:ft}
@@
 
-Once again, table \ref{results:tab:ft} shows testing performance, while figure \ref{results:img:ft} shows validation accuracies.
+Once again, Table \ref{results:tab:ft} shows testing performance, while Figure \ref{results:img:ft} shows validation accuracies.
 
@@
 Of all related work mentioned in section \ref{related_work}, \citeauthor{zhou2021convnets} (\citeyear{zhou2021convnets}) was the only one that examined VTs' OTS learning properties.
-It reported the best OTS learning performance for the \texttt{Swin} VT, which is much in line with the results in table \ref{results:tab:ots}.
+It reported the best OTS learning performance for the \texttt{Swin} VT, which is much in line with the results in Table \ref{results:tab:ots}.
 
@@
 Running the same experiments with \texttt{ResNet101} does, however, lead to similar findings as \citeauthor{zhou2021convnets}.
-On \textit{Type classification}, for example, it achieves a mean testing accuracy of  $86.21\% (\pm 0.70\%)$, which is indeed slightly higher than \texttt{ViT}'s 86.06\% shown in table \ref{results:tab:ots}.
+On \textit{Type classification}, for example, it achieves a mean testing accuracy of  $86.21\% (\pm 0.70\%)$, which is indeed slightly higher than \texttt{ViT}'s 86.06\% shown in Table \ref{results:tab:ots}.
 Mind that these differences are slight compared to the 89.43\% accuracy achieved by \texttt{Swin} here, so replacing \texttt{ResNet50} with \texttt{ResNet101} would not have affected the overall observations that much.
 
-Finally, because \texttt{VGG19} was specifically selected for its good OTS performance in \citeauthor{sabatelli2018deep} (\citeyear{sabatelli2018deep}), it is interesting that this model is outperformed by \texttt{ResNet50} (which that paper also included) in table \ref{results:tab:ots}.
+Finally, because \texttt{VGG19} was specifically selected for its good OTS performance in \citeauthor{sabatelli2018deep} (\citeyear{sabatelli2018deep}), it is interesting that this model is outperformed by \texttt{ResNet50} (which that paper also included) in Table \ref{results:tab:ots}.
 \texttt{VGG19} is the largest model of all, so this discrepancy could be an indication that the smaller datasets used in the current study are too small for \texttt{VGG19}.
@@
 
-A hint of an answer can, however, be found when plotting activation of successively deeper attention layers, as is done in figure \ref{results:img:layers}.
+A hint of an answer can, however, be found when plotting activation of successively deeper attention layers, as is done in Figure \ref{results:img:layers}.
 While earlier layers retain much spatial information, later ones show more sporadic activation patterns.
@@
 In that paper, however, \texttt{ViT} outperforms the \texttt{ResNet101} and \texttt{ResNet152} CNNs.
-This differs from the results in table \ref{results:tab:ft}, where \texttt{ViT} does not perform well overall, and even shows the worst testing accuracy on \textit{Type classification}, with 90.11\%.
+This differs from the results in Table \ref{results:tab:ft}, where \texttt{ViT} does not perform well overall, and even shows the worst testing accuracy on \textit{Type classification}, with 90.11\%.
 \texttt{ResNet50} achieves 91.78\% accuracy here, and testing the same \texttt{ResNet101} and \texttt{ResNet152} versions that \citeauthor{zhou2021convnets} used does not change much, as this leads to 91.96\% ($\pm$ 0.59\%) and 91.94\% ($\pm$ 0.25) accuracy, respectively.
@@
 Results are also less favorable for VTs compared to \citeauthor{matsoukas2021time} (\citeyear{matsoukas2021time}).
-In that paper, the `tiny' version of \texttt{DeiT} outperforms \texttt{ResNet50}, while in table \ref{results:tab:ft} this same \texttt{ResNet50} performs slightly better than the `base' version of \texttt{DeiT}.
-Going `tiny' also does not help: on \textit{Type classification}, for instance, `tiny' \texttt{DeiT} gets 90.44\% ($\pm$ 0.93\%), which is even worse than `base' \texttt{DeiT}'s 91.78\% in table \ref{results:tab:ots}.
+In that paper, the `tiny' version of \texttt{DeiT} outperforms \texttt{ResNet50}, while in Table \ref{results:tab:ft} this same \texttt{ResNet50} performs slightly better than the `base' version of \texttt{DeiT}.
+Going `tiny' also does not help: on \textit{Type classification}, for instance, `tiny' \texttt{DeiT} gets 90.44\% ($\pm$ 0.93\%), which is even worse than `base' \texttt{DeiT}'s 91.78\% in Table \ref{results:tab:ots}.
 
@@
 \texttt{ConvNext} appears to benefit more when training the whole model instead of the last layer, because it shows the biggest differences between OTS learning and FT.
-In the case of image \ref{results:img:sal:dish}, this even makes the difference between correct and incorrect classification.}
+In the case of Figure \ref{results:img:sal:dish}, this even makes the difference between correct and incorrect classification.}
     \label{results:img:sal}
@@
 
-VTs and CNNs are much closer together after FT in table \ref{results:tab:ft} than after OTS learning in table \ref{results:tab:ots}.
+VTs and CNNs are much closer together after FT in Table \ref{results:tab:ft} than after OTS learning in Table \ref{results:tab:ots}.
 It is as if OTS learning already unlocks much of VTs' potential, while CNNs need an FT learning scheme for this.
 
-Supporting `evidence' might be visible in figure \ref{results:img:sal}.
+Supporting `evidence' might be visible in Figure \ref{results:img:sal}.
 It plots saliency maps for one CNN (\texttt{ConvNext}), and two VTs (\texttt{DeiT} and \texttt{Swin}).
 It is interesting that \texttt{ConvNext} shows much bigger differences between the OTS and FT plots.
-In figure \ref{results:img:sal:picture}, for example, it learns to just look at the picture frame after FT.
-In figure \ref{results:img:sal:dish} it learns to look more at the center, which allows it to correctly classify the input as a dish instead of a plate\footnote{No idea what the difference is between these classes, but these models do seem to know, so that is good!}.
+In Figure \ref{results:img:sal:picture}, for example, it learns to just look at the picture frame after FT.
+In Figure \ref{results:img:sal:dish} it learns to look more at the center, which allows it to correctly classify the input as a dish instead of a plate\footnote{No idea what the difference is between these classes, but these models do seem to know, so that is good!}.
 % TODO Maybe delete this comment, not very professional
@@
 
-The images in figure \ref{results:img:sal} are representative examples, and show observations that could have been made from other images too.
+The images in Figure \ref{results:img:sal} are representative examples, and show observations that could have been made from other images too.
 How representative the 3 models are for all 8 is debatable, and was not tested.
@@
 It was mentioned earlier that balanced accuracy falls behind plain accuracy the most for \textit{Material classification}, and the least for \textit{Artist classification}.
-At the same time, table \ref{methods:datasets} shows that it is also \textit{Material classification} which is the least balanced (Gini-coefficient of 0.563), while it is \textit{Artist classification} that is the most balanced (coefficient 0.236).
+At the same time, Table \ref{methods:datasets} shows that it is also \textit{Material classification} which is the least balanced (Gini-coefficient of 0.563), while it is \textit{Artist classification} that is the most balanced (coefficient 0.236).
 This suggests that balance has a positive effect on testing accuracy, from which \textit{Artist classification}, in that case, benefits the most.
@@
 Of course, final accuracy is not the full story, since time/accuracy trade-offs, for example, can also determine which model is preferred.
-Figure \ref{results:img:ots_vs_ft_type} plots the \textit{Type classification} accuracies of tables \ref{results:tab:ots} and \ref{results:tab:ft} against training time per epoch.
+Figure \ref{results:img:ots_vs_ft_type} plots the \textit{Type classification} accuracies of Tables \ref{results:tab:ots} and \ref{results:tab:ft} against training time per epoch.
 Total training time was not chosen instead, as this measure depends much more on selected hyperparameters.
@@
 All `tiny' VTs are around the same size as \texttt{ResNet50}, and using these would have changed the situation completely.
-As such, taking figure \ref{results:img:ots_vs_ft_type} with a grain of salt would be appropriate.
+As such, taking Figure \ref{results:img:ots_vs_ft_type} with a grain of salt would be appropriate.
 

### 2022-08-07 writing/tex/introduction.tex
Incroporated some changes from Iris' feedback. NOW IT IS DONE. GOING TO HAND IT IN NEXTgit add -A!

@@
 CNNs are largely invariant to shifts and small distortions of the input image.
-More important, is that they can be seen as automatic feature extractors.
+More important is that they can be seen as automatic feature extractors.
 It is well established (e.g.
@@
 
-Another `shortcomming' of \citeauthor{matsoukas2021time} is that it merely examined whether VTs can use TL to outperform CNNs in a domain that happens to be characterized by limited-sized datasets, but that it did not actively focus on the extent to which VTs can be made useful in limited data/compute environments.
+Another `shortcoming' of \citeauthor{matsoukas2021time} is that it merely examined whether VTs can use TL to outperform CNNs in a domain that happens to be characterized by limited-sized datasets, but that it did not actively focus on the extent to which VTs can be made useful in limited data/compute environments.
 More related to this is \citeauthor{touvron2021training} (\citeyear{touvron2021training}), which proposed transformer-specific distillation-based learning techniques to make VTs applicable for smaller datasets.

### 2022-08-07 writing/tex/methods.tex
Incroporated some changes from Iris' feedback. NOW IT IS DONE. GOING TO HAND IT IN NEXTgit add -A!

@@
 For robustness, five datasets are randomly generated per classification task, such that each experiment can be run five times.
-The average sample overlap between two of the five sets is shown in table \ref{methods:datasets} as well.
+The average sample overlap between two of the five sets is shown in Table \ref{methods:datasets} as well.
 
-Besides the tasks described above, table \ref{methods:datasets} shows an experiment called \textit{Scaling}.
+Besides the tasks described above, Table \ref{methods:datasets} shows an experiment called \textit{Scaling}.
 The goal of this experiment is to see how the findings of this paper hold up as dataset sizes are gradually shrunk.
 It does this by taking the top 15 classes of type classification, and scaling it 4 times by a factor $\sqrt[4]{\frac{1}{10}} \approx 0.56$.
-The 5 datasets produced in this manner are then respectively 100\%, 56\%, 32\%, 18\%, and 10\% the size of the dataset shown in table \ref{methods:datasets}.
+The 5 datasets produced in this manner are then respectively 100\%, 56\%, 32\%, 18\%, and 10\% the size of the dataset shown in Table \ref{methods:datasets}.
 Steps are taken to ensure that the distribution remains the same.
@@
 In this thesis, however, also a base-sized version is chosen.
-Specifically, the version that does not include output for distillation learning.
+This version does not include output for distillation learning, so no advantage is taken of that.
+
+%Specifically, the version that does not include output for distillation learning.
 
@@
 For FT, more is needed to achieve good convergence, especially for VT-based architectures.
-Used hyperparameters, etc.
-are partially inspired by \citeauthor{matsoukas2021time} (\citeyear{matsoukas2021time}) and \citeauthor{zhou2021convnets} (\citeyear{zhou2021convnets}).
+The learning scheme used here is partially inspired by \citeauthor{matsoukas2021time} (\citeyear{matsoukas2021time}) and \citeauthor{zhou2021convnets} (\citeyear{zhou2021convnets}).
 The Adam learning rate now starts off at 1e-4, and is reduced by a factor 10 after 3 epochs without improvement (i.e.

### 2022-08-09 writing/tex/conclusion.tex
Incorporated feedback from Dr. Sabatelli. Now its time to REALLY submit itgit add -A! :-D

@@
 Seeing which is better, VTs or CNNs, has never been the goal of this paper.
-The underlying architecture is somewhat irrelevant, as one should always pick (or create) the model that fits the task best -- regardless of whether it is a VT, CNN, or something else.
+The underlying architecture is somewhat irrelevant, as one should always pick (or create) the model that fits the task best -- regardless of whether it is a VT, or CNN.
 What this paper, then, was interested in, is whether VTs can become a valuable alternative to CNNs under circumstances in which:

### 2022-08-09 writing/tex/experiments.tex
Incorporated feedback from Dr. Sabatelli. Now its time to REALLY submit itgit add -A! :-D

@@
 \section{Experiments}
-This section focuses on the results gotten from the conducted experiments.
+This section focuses on the results obtained from the conducted experiments.
 These will first briefly be presented in section \ref{results}.
-After that, section \ref{exp:int} will discuss them in more detail, and see how they compare to earlier studies, etc.
+After that, section \ref{exp:int} will discuss them in more detail, and see how they compare to earlier studies.
 
@@
 In Figure \ref{results:img:sal:picture}, for example, it learns to just look at the picture frame after FT.
-In Figure \ref{results:img:sal:dish} it learns to look more at the center, which allows it to correctly classify the input as a dish instead of a plate\footnote{No idea what the difference is between these classes, but these models do seem to know, so that is good!}.
-% TODO Maybe delete this comment, not very professional
+In Figure \ref{results:img:sal:dish} it learns to look more at the center, which allows it to correctly classify the input as a dish instead of a plate.
 These bigger differences for \texttt{ConvNext} align well with observations made earlier, as they suggest that there is still much improvement possible after just OTS learning for CNNs.

### 2022-08-09 writing/tex/introduction.tex
Incorporated feedback from Dr. Sabatelli. Now its time to REALLY submit itgit add -A! :-D

@@
 These are combined in subsequent layers to detect higher-order features.
-The final layer should then give a representative summary of the input, such that, for example, an MLP-type layer can extract all the information it needs from it to perform the task at hand.
+The final layer should then give a representative summary of the input, such that, for example, a Multilayer Perceptron-type (MLP-type) layer can extract all the information it needs from it to perform the task at hand.
 
@@
 The self-attention mechanism plays a central role in VTs, and is worth elaborating on.
-Each head of a self-attention layer contains 3 learned projection matrices, which are used to produce $key$, $value$, and $query$ vectors for every position in the input sequence.
+Each head of a self-attention layer contains three learned projection matrices, which are used to produce $key$, $value$, and $query$ vectors for every position in the input sequence.
 The output for a single position is a weighted sum of all $value$ vectors.
-Weights are obtained by matching this position's $query$ with all $key$ vectors\footnote{In principle by taking softmax over all query-key inner products.}.
+Weights are obtained by matching this position's $query$ with all $key$ vectors\footnote{In principle by taking the softmax over all query-key inner products.}.
 As such, an attention head decides itself how much of each position should be incorporated in subsequent representations.
 
-The original Transformer architecture was designed to solve problems in \textit{natural language processing} (a field it now dominates).
+The original Transformer architecture was designed to solve problems in \textit{natural language processing} -- a field it now dominates \citep{wolf2020transformers}.
 To make it work for CV applications, VTs divide an image into square patches, which get treated as tokens (words).
@@
 \paragraph{CNNs versus VTs}
-Where CNNs reduce the number of trainable weights by taking an image's topology into account, VTs do this by learning themselves what information they should incorporate in successive representations of the input.
+While CNNs reduce the number of trainable weights by taking an image's topology into account, VTs do this by learning themselves what information they should incorporate in successive representations of the input.
 As such, they are not restricted to a local receptive field, but can potentially include information from all over the previous layer.

### 2022-08-09 writing/tex/methods.tex
Incorporated feedback from Dr. Sabatelli. Now its time to REALLY submit itgit add -A! :-D

@@
 \subsection{Models} \label{methods:models}
-Eight different neural network architectures are used throughout this paper, of which four are CNN-based and four VT-based.
+Eight different neural network architectures are used, of which four are CNN-based and four VT-based.
 When multiple versions are available (i.e.
